---
title: "Scholarships - ACEX"
author: "PAM Lab - UWF"
date: "3/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
setwd("G:/My Drive/#1_Jay's Space/Scholarships-YOY")
library(xfun)
```

```{r eval=FALSE, include=FALSE}
# need to select variables and created matching ID with ACEX and data (filter SIF=="Stu_DEMO_DATA_SOURCE")
library(readr)
library(dplyr)
COURSE_STUDENT_INSTRUCTION_SPRING21 <- read_csv("G:/Shared drives/HMCSE-PAM Lab/DATA 202101/202101/COURSE_STUDENT_INSTRUCTION_SPRING21.csv") %>% filter(Course_DEMO_TIME_FRAME>=201505) %>% 
   select(STU_ID,Course_DEMO_TIME_FRAME,contains("term"),Course_CrsTitle, contains("combined"),Course_CrsPrefix,Course_CrsSuffix, contains("number"), contains("award"),contains("credit"))
write.csv(COURSE_STUDENT_INSTRUCTION_SPRING21, "mycrs_acex.csv")

STU_ENROLLMENT_SPRING21 <- read_csv("G:/Shared drives/HMCSE-PAM Lab/DATA 202101/202101/STU_ENROLLMENT_SPRING21.csv") %>%  filter(Stu_DEMO_DATA_SOURCE=="SIF" & Stu_DEMO_TIME_FRAME>= 201505) %>% 
  select(STU_ID, contains("demo"),contains("major"),Stu_CurrentTermLoad,contains("GPA"), contains("inst"), contains("total"),contains("credit"), contains("semester"),contains("coll"),contains("depart")) %>% arrange(Stu_DEMO_TIME_FRAME)
write.csv(STU_ENROLLMENT_SPRING21, "myenc_acex.csv")


STUDENT_DEGREE_SPRING21 <- read_csv("G:/Shared drives/HMCSE-PAM Lab/DATA 202101/202101/STUDENT_DEGREE_SPRING21.csv") %>%  select(STU_ID,  contains("data"), Deg_TermGranted, contains("desc"), contains("major"), contains("banner"))
write.csv(STUDENT_DEGREE_SPRING21, "mydeg_acex.csv")

CSE_MATCH_IDS_1_ <- read_csv("C:/Users/jsparks3/Downloads/CSE_MATCH_IDS (1).csv") # FTIC2015 to 2019

library(readxl)
ACEX_Data_by_Cohort2_22_21 <- read_excel("G:/Shared drives/HMCSE-PAM Lab/ACEX/ACEX Data by Cohort2-22-21.xlsx")
ACEX_Data_by_Cohort2_22_211 <- ACEX_Data_by_Cohort2_22_21[!is.na(ACEX_Data_by_Cohort2_22_21$`Aid Year`),]
#colSums(is.na(ACEX_Data_by_Cohort2_22_211))

Final_id_cohort <- Final_acex_enc_DF %>% select(uwfid, cohort_year) %>% mutate(Cohort=ifelse(cohort_year=="17-18", "FTIC2017",
                                                                                             ifelse(cohort_year=="18-19","FTIC2018",
                                                                                                    ifelse(cohort_year=="19-20", "FTIC2019","FTIC2020")))) %>% select(uwfid, Cohort)
Final_id_cohort_nodup <- Final_id_cohort[!duplicated(Final_id_cohort),] #1709
Final_acex_enc_DFtwo <- merge( Final_acex_enc_DF, Final_id_cohort_nodup, by=c("uwfid"), all.x = T)

Final_acex_enc_DF3 <- Final_acex_enc_DFtwo[!is.na(Final_acex_enc_DFtwo$Cohort),] %>% group_by(uwfid) %>% arrange(Index)
# export data
write.csv(Final_acex_enc_DF3, "V1_Final_acex_enc_DF.csv") 
```

## 1. Questions:

  1. How many FTIC with AXEC:
  
    a. maintained it and stayed at UWF?
    b. maintained it and left UWF?
    c. lost it and stayed at UWF?
    d. lost it and left UWF?
    
  2. What are the proportions of A and B+C+D?
  
  3. What factors best predict an FTIC with AXEC being in A as opposed to B, C, or D?


## 2. Requirements to Remain Eligible for Financial Aid

  * GPA: 3.00 cumulative UWF GPA
  * Full-time enrollment of 12+ credit hours (not summer)
  * At least 30 credit hours / academic year
  * Not be reinstated
  
 [Freshman Scholoarships-UWF](https://uwf.edu/admissions/undergraduate/cost-and-financial-aid/awards-and-scholarships/freshman/)

## 3. Factors should be Considered
- Prior Hours
- Gender
- CRS+GPA: maintained requirements
- Impact of switched majors: scholarship and extra hours (needs info)

## 4. Cleaning Data
- Total of 3,645 FTIC ACEX cases 
- Total of 1,935 unique FTICs in the ACEX data (FT17 to FT20)
- 555 FTIC ACEX (2020) were removed for lack of enrollment information (enrollments data ends Fall 2020)
- 36 UWFID with ACEX but not in CSE data were removed
- Total of 591 FTICs were removed from ACEX data 
- Total of 1,344 FTICs were used for this document
- For degree information, the most recent (Fall 2020) was used 
- For dual-degrees, the first alphabetically was used
- ID = 970522062 is FTIC 2017, but was mislabeled 2019; this was corrected


```{r read data, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(readr)
library(dplyr)
library(readxl)
#used dataname
myenc_acex <- read_csv("myenc_acex.csv", 
    col_types = cols(X1 = col_skip())) %>% filter(Stu_DEMO_TIME_FRAME== 201708 | Stu_DEMO_TIME_FRAME ==201808 | Stu_DEMO_TIME_FRAME ==201908 |Stu_DEMO_TIME_FRAME==202008)

#used dataname
mycrs_acex <- read_csv("mycrs_acex.csv", 
    col_types = cols(X1 = col_skip()))


#used dataname
mydeg_acex <- read_csv("mydeg_acex.csv", 
    col_types = cols(X1 = col_skip()))

# Id match
CSE_MATCH_IDS_1_ <- read_csv("G:/My Drive/#1_Jay's Space/Scholarships-YOY/CSE_MATCH_IDS (1).csv") # FTIC2015 to 2019


#used dataname
mydeg_acex <- read_csv("mydeg_acex.csv", 
    col_types = cols(X1 = col_skip()))

# usded dataname


acex_cleandata <- read.csv("G:/My Drive/#1_Jay's Space/Scholarships-YOY/acex_cleandata.csv")

# used final data

Final_acex_enc_DF <- read_csv("V1_Final_acex_enc_DF.csv") %>% select(-1)


#length(unique(Final_acex_enc_DF$STU_ID))
#xtabs(~Final_acex_enc_DF$cohort_year+ Final_acex_enc_DF$Num_aidYear)
```


```{r eval=FALSE, include=FALSE}
head(acex_cleandata[,-1])
```

### Numbers of FTIC by Aid Years (before removed NA) 

```{r message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(scales)
acex_cleandata <- read.csv("G:/My Drive/#1_Jay's Space/Scholarships-YOY/acex_cleandata.csv")
#length(unique(acex_cleandata$uwfid))


library(plotly)
cleannames_ACEX1_fig1 <- acex_cleandata %>% group_by(cohort_year, Num_aidYear ) %>% dplyr::summarise(count=n()) %>% 
  plot_ly( x=~cohort_year, y=~ count, color=~ Num_aidYear, type="bar" ) %>% 
  layout( xaxis=list(title="COHORT YEAR"), yaxis=list(title="Counts by Aid Year"), barmode="stack")

cleannames_ACEX1_fig1

```

### ACEX Offered and Paid Amounts by FTIC (before removed NA)


```{r}
Summary_acex <- acex_cleandata %>% group_by(cohort_year, aid_year) %>% dplyr::summarise(Count=n(), SumOffered=sum(acex_offered), SumPaid=sum(acex_paid), .groups='drop') %>%  mutate(Diff_YOY= SumPaid -lag(SumPaid, default = SumPaid[1])) %>%  mutate(Diff_YOY1= Count -lag(Count, default = Count[1]))


SumOffered <- dollar(Summary_acex$SumOffered)
Summary_acex$SumOffered <- SumOffered
SumPaid <- dollar(Summary_acex$SumPaid)
Summary_acex$SumPaid <- SumPaid
Diff_YOY <- dollar(Summary_acex$Diff_YOY)
Summary_acex$Diff_YOY <- Diff_YOY

#%>%  arrange(-Count)
knitr::kable(Summary_acex, caption = " Numbers of Cohort vs. Sum of Aid Amounts (before NA)")
# loosing by Tier
```


### ACEX FTIC by Tiers (before removed NA)

```{r summary by tier, echo=FALSE, message=FALSE, warning=FALSE}

Summary_acex_tiers <- acex_cleandata %>% group_by(cohort_year,tier,Num_aidYear) %>% dplyr::summarise(Count=n(),SumOffered=sum(acex_offered), SumPaid=sum(acex_paid)) %>%  mutate(Diff_YOY= SumPaid -lag(SumPaid, default = SumPaid[1])) %>%  mutate(Diff_YOY1= Count -lag(Count, default = Count[1]))

SumOffered <- dollar(Summary_acex_tiers$SumOffered)
Summary_acex_tiers$SumOffered <- SumOffered
SumPaid <- dollar(Summary_acex_tiers$SumPaid)
Summary_acex_tiers$SumPaid <- SumPaid
Diff_YOY <- dollar(Summary_acex_tiers$Diff_YOY)
Summary_acex_tiers$Diff_YOY <- Diff_YOY

#Summary table
library(DT)
summary_acex_tiers_DT <- datatable(Summary_acex_tiers, extensions = "Buttons",caption = "ACEX-FTIC Trends by Tiers",
          filter = "top",
          options = list(dom="Blfrtip",buttons=c("copy","csv","excel","pdf","print"), lengthMenu=list(c(10,25,50,-1), c(10,25,50,"All")),pageLength=25) )
summary_acex_tiers_DT
```



```{r cleaned data, eval=FALSE, include=FALSE}
library(rvest)
library(tidyverse)
library(janitor)
colSums(is.na(cleannames_ACEX1))
cleannames_ACEX3 <- cleannames_ACEX1[, c(9:14)] %>% replace(is.na(.),0)
cleannames_ACEX4 <- cleannames_ACEX1[, c(1:8)]
cleannames_ACEX1 <- cbind(cleannames_ACEX4,cleannames_ACEX3) %>% group_by( uwfid ) %>%  mutate(Num_aidYear=row_number())
AidYear <- rep("AidYear")
cleannames_ACEX1$Num_aidYear <- as.factor(cleannames_ACEX1$Num_aidYear)
cleannames_ACEX1$Num_aidYear <- paste(AidYear,cleannames_ACEX1$Num_aidYear)
#write.csv(cleannames_ACEX1,"acex_cleandata.csv")
```


### No information in dataset STU_IDs


```{r echo=FALSE, message=FALSE, warning=FALSE}

acex_cleandata_STU_ID <- merge(acex_cleandata, CSE_MATCH_IDS_1_, by.x= "uwfid", by.y= "UNIV_ROW_ID", all.x = T ) %>% group_by(uwfid) %>% arrange(aid_year)%>%  mutate(Index=row_number())
#colSums(is.na(acex_cleandata_STU_ID))
cohort_size <- xtabs(~acex_cleandata_STU_ID$cohort_year+ acex_cleandata_STU_ID$Index)
acex_cleandata_STU_ID_17to19 <- acex_cleandata_STU_ID %>% filter(cohort_year != "20-21")
NAs <- acex_cleandata_STU_ID_17to19[which(is.na(acex_cleandata_STU_ID_17to19$STU_ID)),] %>% group_by(uwfid) %>% arrange(aid_year) %>% select(-Num_aidYear)
Missing <- xtabs(~NAs$cohort_year+NAs$aid_year)

acex_cleandata_STU_ID_17to19_nona <- acex_cleandata_STU_ID_17to19[!is.na(acex_cleandata_STU_ID_17to19$STU_ID),] #3,042
acex_id_only <- acex_cleandata_STU_ID_17to19_nona %>% select(uwfid, STU_ID)
all_id_acex_only <- acex_id_only[!duplicated(acex_id_only$uwfid),] #,1344

knitr::kable(Missing)
#NAs
```

### Total FTIC and aid years after removed NAs

```{r}
total_sum <- xtabs(~Final_acex_enc_DF$cohort_year + Final_acex_enc_DF$aid_year)
knitr::kable(total_sum, caption = "Numbers of ACEX Cohort vs. Aid Year")
```

```{r ACEX, eval=FALSE, include=FALSE}
CSE_MATCH_IDS_1_[CSE_MATCH_IDS_1_$STU_ID=="DC8FE3C332A7D51614C81674186529DAA8A35825","UNIV_ROW_ID"]
acex <- acex_cleandata[acex_cleandata$uwfid== "970398543", ]
acex
```

### Example: Lost ACEX but Maintaned at UWF

```{r merging two data set, eval=FALSE, include=FALSE}

myenc_acex_term_index <- myenc_acex %>% group_by(STU_ID) %>% arrange(Stu_DEMO_TIME_FRAME) %>%  mutate(Index=row_number())

# all about enc 3,642
acex_cleandata_STU_ID_enc <- merge(all_id_acex_only, myenc_acex_term_index, by=c("STU_ID"), all.x = T) %>% group_by(STU_ID) %>% arrange(Index) 

# all about acex 3,042
enc_acex_all_STU_ID <- merge(all_id_acex_only, acex_cleandata_STU_ID, by=c("uwfid","STU_ID"), all.x = T ) %>% group_by(STU_ID) %>% arrange(Index)

enc_acex_all_STU_ID2 <-merge(acex_cleandata_STU_ID_enc,enc_acex_all_STU_ID, by=c("uwfid","STU_ID","Index"), all.x = T)

ex1 <- enc_acex_all_STU_ID2[enc_acex_all_STU_ID2$STU_ID=="DC8FE3C332A7D51614C81674186529DAA8A35825",]
```

```{r}
library(DT)
EX1 <- Final_acex_enc_DF[Final_acex_enc_DF$STU_ID=="DC8FE3C332A7D51614C81674186529DAA8A35825",]
#EX1
EX_table <- datatable(EX1)
EX_table
```

```{r with degree info, eval=FALSE, include=FALSE}

deg_acex <- merge(all_id_acex_only, mydeg_acex, by="STU_ID", all.x = T) %>% group_by(STU_ID) %>% filter(Deg_TermGranted==max(Deg_TermGranted))
deg_acex1 <- deg_acex[!duplicated(deg_acex$STU_ID),]

enc_acex_all_STU_ID3 <- merge(enc_acex_all_STU_ID2, deg_acex1, by=c("STU_ID","uwfid"), all.x = T) %>% group_by(uwfid) %>% arrange(Index)
write.csv(enc_acex_all_STU_ID3, "Final_acex_enc_DF.csv")
```

```{r used Final_acex_enc_DF, eval=FALSE, include=FALSE}
ex_deg <- Final_acex_enc_DF[Final_acex_enc_DF$STU_ID=="DC8FE3C332A7D51614C81674186529DAA8A35825",]

ex_deg
```

### Example: Lost ACEX and left at UWF

```{r}
library(DT)
EX2 <- Final_acex_enc_DF[Final_acex_enc_DF$STU_ID=="74AAECD2F0EC58EA82CDE143B404AF0041E24C6E",]
#EX2
EX2_table <- datatable(EX2)
EX2_table
```
### Example: Remained ACEX and staying at UWF

```{r}
EX3 <- Final_acex_enc_DF[Final_acex_enc_DF$STU_ID=="3B6593E215D1346FFB4DC10580A270C8C86DA559",]
#EX3
EX3_table <- datatable(EX3)
EX3_table
```

### Example: Remained ACEX and graduated 

```{r}

EX4_a <- Final_acex_enc_DF[Final_acex_enc_DF$STU_ID=="111D28829BE5F0630859445BC0BF02028FD59654",]
#EX4_a
EX4_a_table <- datatable(EX4_a)
EX4_a_table
```


### Number of UWF graduates in cohort (Fall 2020)

```{r echo=FALSE, message=FALSE, warning=FALSE}

# method 1 (had fill cohort_year 625 NAs that did not created when it merged)
Max_index <- Final_acex_enc_DF %>%   group_by(uwfid) %>%  top_n(1, Index) # choose max index by uwfid
Max_index$B_deg <- ifelse(Max_index$Deg_Desc=="Bachelor",1,0)
Max_index[which(is.na(Max_index$B_deg)), "B_deg"] <- 0 # for nas
Max_index <- Max_index %>% mutate(leftUWF=ifelse(B_deg ==1, "Bachelor" ,
                                                 ifelse((B_deg==0 & Stu_DEMO_TIME_FRAME != 202008), "Dropped", "Stayed")))
table_degree <-Max_index %>% group_by(Cohort,B_deg ) %>% dplyr::summarise(Count=n())
knitr::kable(table_degree)
```

### What is the proportion that maintained ACEX? (FTIC 2017)

  - Total 1,718 ACEX in FTIC 2017 
  - 516 unique FTIC 2017 earned ACEX in their first year
  - 151 FTIC earned bachelor degree by Fall 2020
  - 23 of 151 were in graduate programs

```{r cleaningFTIC2017, eval=FALSE, include=FALSE}
#create B-deg
FTIC2017_YOY <- Final_acex_enc_DF %>% filter(Cohort=="FTIC2017") %>% mutate(B_deg=ifelse(Deg_Desc=="Bachelor","1","0"))
FTIC2017_YOY[which(is.na(FTIC2017_YOY$B_deg)), "B_deg"] <- "0"
#create earned hours by year
FTIC2017_YOY_up <- FTIC2017_YOY %>% group_by(uwfid) %>% mutate(EarnedHrs_Year = Stu_TotalInstHours - lag(Stu_TotalInstHours, default = Stu_TotalInstHours[1])) %>%group_by(uwfid) %>%  mutate(TGP_Year=Stu_TotalInstGradePoints- lag(Stu_TotalInstGradePoints, default = Stu_TotalInstGradePoints[1])) %>% mutate(GPA_Year = round(TGP_Year/EarnedHrs_Year,digits = 4))
#colSums(is.na(FTIC2017_YOY_up))
negative_values <- FTIC2017_YOY_up %>% filter(EarnedHrs_Year<0) %>% select(uwfid, Index, Stu_DEMO_TIME_FRAME) %>%  mutate(Mater_UWF=rep("Master_UWF"))
#length(unique(negative_values$uwfid)) #23 in graduate level
#omit master enc info #1,718 master 23, so 1,695
Final_FTIC2017 <- anti_join(FTIC2017_YOY_up,negative_values, by=c("uwfid","Index","Stu_DEMO_TIME_FRAME"),all.x=T ) 

ACEX_FTIC2017 <- Final_FTIC2017 %>% group_by(uwfid) %>% mutate(GPACRD_Year= ifelse( (GPA_Year>=3.0000 & EarnedHrs_Year >= 30.0000), "On_Track",
                                                                                 ifelse((GPA_Year>=3.0000 & EarnedHrs_Year < 30.0000),"On_GPA",
                                                                                 ifelse((GPA_Year < 3.0000 & EarnedHrs_Year >= 30.0000), "On_CRD30",
                                                                                        ifelse((GPA_Year < 3.0000 & EarnedHrs_Year < 30.0000), "Off_Track","NA")))))

ACEX_FTIC2017[which(is.na(ACEX_FTIC2017$Num_aidYear)), "Num_aidYear"] <- "No_ACEX"
ACEX_FTIC2017[which(is.na(ACEX_FTIC2017$GPACRD_Year)), "GPACRD_Year"] <- "ACEX_1st"
Final_ACEX_FTIC2017 <- ACEX_FTIC2017
write.csv(Final_ACEX_FTIC2017, "ACEX_FTIC2017.csv")
```

### Numbers of On-Track FTIC by Terms

- On-Track: earned >=30 credits and kept GPA >= 3.00 per year
- Off-Track: did not earn >=30 credits and below GPA 3.00 per year
- On-CRD20: earned >=30 credits but GPA below 3.00 per year
- On-GPA: kept GPA >= 3.00 but earned less than 30 credits per year

```{r eval=FALSE, message=FALSE, warning=FALSE,  echo=FALSE, include=FALSE}
library(dplyr)
ACEX_FTIC2017 <- read.csv("G:/My Drive/#1_Jay's Space/Scholarships-YOY/ACEX_FTIC2017.csv", stringsAsFactors=TRUE) %>% select(-1)
strange <- ACEX_FTIC2017[ACEX_FTIC2017$GPACRD_Year=="ACEX_1st" & ACEX_FTIC2017$Stu_DEMO_TIME_FRAME=="202008",]
ACEX_FTIC2017[ACEX_FTIC2017$uwfid=="970518096" & ACEX_FTIC2017$Index=="3", "GPACRD_Year"] <- "Off_Track"
ACEX_FTIC2017[ACEX_FTIC2017$uwfid=="970527857" & ACEX_FTIC2017$Index=="3", "GPACRD_Year"] <- "Off_Track"
ACEX_FTIC2017[ACEX_FTIC2017$uwfid=="970525139" & ACEX_FTIC2017$Index=="4", "GPACRD_Year"] <- "Off_Track"
ACEX_FTIC2017 <- ACEX_FTIC2017 %>% mutate(GPACRD_Year1=ifelse(GPACRD_Year =="On_Track", "On_Track", "Off_Track")) 
#970525139
write.csv(ACEX_FTIC2017, "ACEX_FTIC2017_03_05_21.csv")

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
ACEX_FTIC2017_03_05_21 <- read.csv("G:/My Drive/#1_Jay's Space/Scholarships-YOY/ACEX_FTIC2017_03_05_21.csv", stringsAsFactors=TRUE) %>% select(-1)
Summary_track <- ACEX_FTIC2017_03_05_21 %>% group_by(GPACRD_Year,Stu_DEMO_TIME_FRAME) %>% dplyr::summarise(Count=n()) %>% 
  tidyr::pivot_wider(names_from = Stu_DEMO_TIME_FRAME, values_from=Count) %>% replace(is.na(.),0)
knitr::kable(Summary_track, caption = "CRD+GPA Trend for FTIC 2017")

```

* 215 out of 516 were on-track in the 1st year
* 183 were on-track in the 2nd year
* 149 were on-track in the 3rd year (Fall 2020)


### Proportion of On-Track FTIC by Year

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary_ontrack <- ACEX_FTIC2017_03_05_21 %>% group_by(GPACRD_Year1, Stu_DEMO_TIME_FRAME) %>% dplyr::summarise(Count=n()) %>% 
  tidyr::pivot_wider(names_from = GPACRD_Year1, values_from= Count) %>% replace(is.na(.),0) %>% mutate(Not_Fall2020=516-(Off_Track+On_Track)) %>% 
  mutate(Prop_OnTrack= round(On_Track/516, digits = 4))
knitr::kable( summary_ontrack, caption = "Proportion of On-track per Year (FTIC 2017)")
```
* 41.67% of 516 were on-track in the 1st year (by summer 2018)
* 35.47% were on-track in the 2nd year (by summer 2019)
* 28.88% (149) were on-track in the 3rd year (by summer 2020)
* 29.26% (151) were earned bachelor degree by Fall 2020
  
  
### Proportion of On-track FTIC by Aid-year 

```{r echo=FALSE, message=FALSE, warning=FALSE}
ACEX_ontrack <- ACEX_FTIC2017_03_05_21 %>%filter(Num_aidYear != "AidYear 1")  %>%  group_by(Num_aidYear, GPACRD_Year1) %>% dplyr::summarise(Count=n()) %>% 
  tidyr::pivot_wider(names_from = GPACRD_Year1, values_from=Count) %>% replace(is.na(.),0) %>% 
mutate(SumCountAidFT=(Off_Track+On_Track),  Prop_OnTrackAid= round(On_Track/SumCountAidFT, digits = 4))
knitr::kable(ACEX_ontrack)

```
* 60.12% (208) of the ACEX were on-track in their 2nd year
* 58.21% (163) were on-track in their 3rd year
* 59.41% (120) were on-track in their 4th year
* 15.95% (56) had lost their ACEX, but were on-track (needs more detail) 

