---
title: "Admitted 2021 FTIC - Prediction APR"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

final_admission <- readRDS("./final_admissionFTIC_model.rds")

library(readr)
all_vari_PREDICTION <- read_csv("all_vari_PREDICTION.csv") # clean the final dataset and complete this report


```

## Prediction of APR for 2021 FIC

There are 2969 FTIC admitted and should predict who will be APR or non-APR

* Factors are considered

    + APR by program
    + APR by high school names
    + GENDER (female and male)
    + HS GPA by tenths
    
* Data for this analysis from:

    + FTIC 2015 to 2019, total 5,736 data point (removed NA)
    + Data partition: 70% (train) and 30% (test)
    + Accuracy: 80.15%(train) and 78.6% (test)
    
* Issues

    + Replacement values are based on average APR by program  or average APR by HS GPA (tenths)
    + 1,872 FTIC with missing HS NAMEs were replaced 

## Included  resutls table

```{r}
library(DT)
results_table <- datatable(all_vari_PREDICTION, extensions = "Buttons",caption = "Probability of APR",
          filter = "top",
          options = list(dom="Blfrtip",buttons=c("copy","csv","excel","pdf","print"), lengthMenu=list(c(10,25,50,-1), c(10,25,50,"All")),pageLength=25))

results_table
```


You can download the table by clicking menu

```{r }



```

