---
title: "FTIC's Dashboard"
output: 
  flexdashboard::flex_dashboard:
    
    theme: cerulean
    orientation: rows
    vertical_layout: fill
    social: ["twitter", "facebook","menu"]
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(knitr)
library(DT)
library(rpivotTable)
library(ggplot2)
library(plotly)
library(dplyr)
library(openintro)
library(highcharter)
library(readr)
```
```{css, echo=FALSE}
.rpivottable{overflow-x: scroll; }

```


```{r}
myFTIC <- FTIC_ALL_APR_CRDV1 <- read_csv("FTIC_ALL_APR_CRDV1.csv",  col_types = cols(X1 = col_skip())) %>%  group_by(STU_ID)
#%>% filter(NumFallTerms== max(NumFallTerms))
myFTIC2015 <- myFTIC %>% filter(Cohort==2015& Stu_DEMO_TIME_FRAME.y==max(Stu_DEMO_TIME_FRAME.y))
myFTIC2016 <- myFTIC %>% filter(Cohort==2016& Stu_DEMO_TIME_FRAME.y==max(Stu_DEMO_TIME_FRAME.y))
myFTIC2017 <- myFTIC %>% filter(Cohort==2017& Stu_DEMO_TIME_FRAME.y==max(Stu_DEMO_TIME_FRAME.y))
myFTIC2018 <- myFTIC %>% filter(Cohort==2018& Stu_DEMO_TIME_FRAME.y==max(Stu_DEMO_TIME_FRAME.y))
myFTIC2019 <- myFTIC %>% filter(Cohort==2019& Stu_DEMO_TIME_FRAME.y==max(Stu_DEMO_TIME_FRAME.y))

```
```{r}
mycolors <- c("blue","#FFC125","darkgreen","darkorange" )
```

Interactive Data Visualization
======================================================================

Row
----------------------------------------------------------------------

### FTIC GPA and Credit Hours (CRD) Analysis

```{r}
valueBox(paste("FTIC size Trends"), color="orange")
```

### Total Numbers of FTIC with GPA APR in UWF Fall 2019

```{r}
myFTICsize <- myFTIC %>% group_by(STU_ID) %>%  filter( NumFallTerms == min(NumFallTerms) )
valueBox(length(unique(myFTICsize$STU_ID)), icon="Total-GPA-FTICs", caption = "FT13-19")
```

### FTIC-2013

```{r}
valueBox(sum(myFTICsize$Cohort==2013), caption = "FT13")
```

### FTIC-2014

```{r}
valueBox(sum(myFTICsize$Cohort==2014), caption = "FT14")
```


### FTIC-2015

```{r}
valueBox(sum(myFTICsize$Cohort==2015), caption = "FT15")
```

### FTIC-2016

```{r}
valueBox(sum(myFTICsize$Cohort==2016), caption = "FT16")
```

### FTIC-2017

```{r}
valueBox(sum(myFTICsize$Cohort==2017), caption = "FT17")
```

### FTIC-2018

```{r}
valueBox(sum(myFTICsize$Cohort==2018), caption = "FT18")
```

### FTIC-2019

```{r}
valueBox(sum(myFTICsize$Cohort==2019), caption = "FT19")
```

### FTIC-2020

```{r}
valueBox(sum(myFTICsize$Cohort==2020), caption = "FT20")
```


Row
-----------------------------------------------------------------------

### GPA by FTIC

```{r}
valueBox(paste("APRGPA Y1 Trends"), color="lightgreen")
```

### Total Numbers of FTIC with GPA APR in UWF 

```{r}
myFTICGPAAPR <- myFTIC %>% group_by(STU_ID) %>%  filter(GPAAPRind=="GPAAPR" & NumFallTerms==2 )
valueBox(length(myFTICGPAAPR$STU_ID), icon="Total-GPACRD-FTICs", caption = "FT13-19")
```


### FTIC - 2013

```{r}
valueBox(sum(myFTICGPAAPR$Cohort==2013),  caption  = "FT13")
```

### CASSH

```{r}
valueBox(sum(myFTICGPAAPR$Cohort==2014),  caption  = "FT14")
```

### COB

```{r}
valueBox(sum(myFTICGPAAPR$Cohort==2015),  caption  = "FT15")

```

### UKCOH

```{r}
valueBox(sum(myFTICGPAAPR$Cohort==2016),  caption  ="FT16")

```

### CEPS

```{r}
valueBox(sum(myFTICGPAAPR$Cohort==2017), caption  ="FT17")

```

### No college designated

```{r}
valueBox(sum(myFTICGPAAPR$Cohort==2018), caption  ="FT18")

```

### FTIC - 2019

```{r}
valueBox(sum(myFTICGPAAPR$Cohort==2019), caption  = "FT19")

```

Row
-----------------------------------------------------------------------

### GPA + CRD y1 by FTIC

```{r}
valueBox(paste("GPACRD Y1 Trends "), color="lightblue")
```

### Total Numbers of FTIC with GPA APR in UWF 

```{r}
myFTICGPACRDAPR <- myFTIC %>% group_by(STU_ID) %>%  filter(CRDGPAAPR=="GPACRDAPR" & NumFallTerms==2)
valueBox(length(myFTICGPACRDAPR$STU_ID), icon="Total-GPACRD-FTICs", caption = "FT13-19")
```


### FTIC - 2013

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2013),  caption  = "FT13")
```

### CASSH

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2014),  caption  = "FT14")
```

### COB

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2015),  caption  = "FT15")

```

### UKCOH

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2016),  caption  ="FT16")

```

### CEPS

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2017), caption  ="FT17")

```

### No college designated

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2018), caption  ="FT18")

```

### FTIC - 2019

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2019), caption  = "FT19")

```

Row
-----------------------------------------------------------------------

### GPA + CRD Y2 by FTIC

```{r}
valueBox(paste("GPACRD Y2 Trends "), color="lightblue")
```

### Total Numbers of FTIC with GPA APR in UWF 

```{r}
myFTICGPACRDAPR <- myFTIC %>% group_by(STU_ID) %>%  filter(CRDGPAAPR=="GPACRDAPR" & NumFallTerms == 3)
valueBox(length(myFTICGPACRDAPR$STU_ID), icon="Total-GPACRD-FTICs", caption = "FT13-19")
```


### FTIC - 2013

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2013),  caption  = "FT13")
```

### CASSH

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2014),  caption  = "FT14")
```

### COB

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2015),  caption  = "FT15")

```

### UKCOH

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2016),  caption  ="FT16")

```

### CEPS

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2017), caption  ="FT17")

```

### No college designated

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2018), caption  ="FT18")

```

### FTIC - 2019

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2019), caption  = "FT19")

```

Row
-----------------------------------------------------------------------

### GPA + CRD Y3 by FTIC

```{r}
valueBox(paste("GPACRD Y3 Trends "), color="lightblue")
```

### Total Numbers of FTIC with GPA APR in UWF 

```{r}
myFTICGPACRDAPR <- myFTIC %>% group_by(STU_ID) %>%  filter(CRDGPAAPR=="GPACRDAPR" & NumFallTerms ==4)
valueBox(length(myFTICGPACRDAPR$STU_ID), icon="Total-GPACRD-FTICs", caption = "FT2013-19")
```


### FTIC - 2013

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2013),  caption  = "FT13")
```

### CASSH

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2014),  caption  = "FT14")
```

### COB

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2015),  caption  = "FT15")

```

### UKCOH

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2016),  caption  ="FT16")

```

### CEPS

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2017), caption  ="FT17")

```

### No college designated

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2018), caption  ="FT18")

```

### FTIC - 2019

```{r}
valueBox(sum(myFTICGPACRDAPR$Cohort==2019), caption  = "FT19")

```

Row
-----------------------------------------------------------------------

### GPA + CRD Graduation by FTIC

```{r}
valueBox(paste("Graduation On-Time"), color="blue")
```

### Total Numbers of FTIC with GPA APR in UWF 

```{r}
library(readr)
library(dplyr)
FTIC_ALL_APR_CRD_DEG_firstV1 <- read_csv("FTIC_ALL_APR_CRD_DEG_firstV1.csv")
myFTICdegOntime <- FTIC_ALL_APR_CRD_DEG_firstV1[!is.na(FTIC_ALL_APR_CRD_DEG_firstV1$Deg_TermGranted),] %>%           select(Cohort,Stu_AdmissionTermCode,Stu_College,Stu_Department,Deg_College,Deg_Department,Deg_TermGranted)  %>%
  mutate(GraduationTime = ifelse((Cohort==2013 & Deg_TermGranted <= 201705)|(Cohort==2014 & Deg_TermGranted <=201805)|(Cohort==2015 & Deg_TermGranted <=201905)|(Cohort==2016 & Deg_TermGranted <=202005)|(Cohort==2017 & Deg_TermGranted <=202105)|(Cohort==2018 & Deg_TermGranted <=202205), "OnTime","OverTime" )) 
colnames(myFTICdegOntime) <- c("Cohort","AdimissionTerm","Entry_college","Entry_depart","Deg_college","Deg_depart","Deg_term","GraduationTime") 

valueBox(sum(myFTICdegOntime$GraduationTime =="OnTime"), icon="Total-degree-FTICs", caption = "FTIC 2013-19")
```


### FTIC - 2013

```{r}
FTIC_ALL_APR_CRD_DEG_firstV1 <- read_csv("FTIC_ALL_APR_CRD_DEG_firstV1.csv")
myFTICdeg <- FTIC_ALL_APR_CRD_DEG_firstV1[!is.na(FTIC_ALL_APR_CRD_DEG_firstV1$Deg_TermGranted), ] 
valueBox(sum(myFTICdeg$Cohort==2013 & myFTICdeg$Deg_TermGranted <= 201705 ),  caption  = "FT13")
```

### CASSH

```{r}
valueBox(sum(myFTICdeg$Cohort==2014 & myFTICdeg$Deg_TermGranted <= 201805),  caption  = "FT14")
```

### COB

```{r}
valueBox(sum(myFTICdeg$Cohort==2015 & myFTICdeg$Deg_TermGranted <= 201905),  caption  = "FT15")

```

### UKCOH

```{r}
valueBox(sum(myFTICdeg$Cohort==2016 & myFTICdeg$Deg_TermGranted <= 202005),  caption  ="FT16")

```

### CEPS

```{r}
valueBox(sum(myFTICdeg$Cohort==2017 & myFTICdeg$Deg_TermGranted <= 202105), caption  ="FT17")

```

### FTIC - 2018

```{r}
valueBox(sum(myFTICdeg$Cohort==2018 & myFTICdeg$Deg_TermGranted <= 202205), caption  ="FT18")

```

### FTIC - 2019

```{r}
valueBox(sum(myFTICdeg$Cohort==2019 & myFTICdeg$Deg_TermGranted <= 202305 ), caption  = "FT19")

```

Row
---------------------------------------------------------------------


### Top B-Degree On-time Departments (Fall 2019 (since FT2013~ ))

```{r}
pie1 <- myFTICdegOntime[myFTICdegOntime$GraduationTime=="OnTime",] %>% 
  group_by(Deg_depart) %>% 
  summarise(count=n()) %>% 
  filter(count>1) %>% 
  plot_ly(labels = ~Deg_depart,
          values = ~count,
          marker = list(colors=mycolors)) %>% 
  add_pie(hole = 0.25) %>% 
  layout(xaxis = list(zeroline = F,
                      showline = F,
                      showticklables = F,
                      showgrid = F),
         yaxis = list(zeroline = F,
                      showline = F,
                      showticklables = F,
                      showgrid = F))
pie1  
```


### Top B-Degree Over-time Departments (Fall 2019 (since FT2013~ ))

```{r}
mycolors2 <- c("darkyellow","lightred","lightgreen","lightblue" )
pie2 <- myFTICdegOntime[myFTICdegOntime$GraduationTime == "OverTime",] %>% 
  group_by(Deg_depart) %>% 
  summarise(count=n()) %>% 
  filter(count>1) %>% 
  plot_ly(labels = ~Deg_depart,
          values = ~count,
          marker = list(colors=mycolors2)
          ) %>% 
  add_pie(hole = 0.25) %>% 
  layout(xaxis = list(zeroline = F,
                      showline = F,
                      showticklables = F,
                      showgrid = F),
         yaxis = list(zeroline = F,
                      showline = F,
                      showticklables = F,
                      showgrid = F))
pie2  
```


Overview - No Bachelor Degree (FT 2013~ )
========================================================================

Row
-------------------------------------------------------------------------

### **FT2013 - Mean of Univ.Credit Hours**

```{r}
myFTIC2015 <- myFTIC %>% filter(Cohort==2013  & Stu_DEMO_TIME_FRAME.y==201908 )#& NumFallTerms >= 4
gauge(round(mean(myFTIC2015$Stu_TotalUniversityHours), digits = 2), min=0, max=150, gaugeSectors(success= c(120,150), warning = c(120,120),danger=c(0,120), colors = c("green","yellow","red")))
```


### **FT2014 - Mean of Univ.Credit Hours**

```{r}
myFTIC2015 <- myFTIC %>% filter(Cohort==2014 & Stu_DEMO_TIME_FRAME.y==201908)
gauge(round(mean(myFTIC2015$Stu_TotalUniversityHours), digits = 2), min=0, max=150, gaugeSectors(success= c(120,150), warning = c(120,120),danger=c(0,120), colors = c("green","yellow","red")))
```


### **FT2015 - Mean of Univ.Credit Hours**

```{r}
myFTIC2015 <- myFTIC %>% filter(Cohort==2015  & Stu_DEMO_TIME_FRAME.y==201908)
gauge(round(mean(myFTIC2015$Stu_TotalUniversityHours), digits = 2), min=0, max=150, gaugeSectors(success= c(120,150), warning = c(120,120),danger=c(0,120), colors = c("green","yellow","red")))
```


### **FT2016 - Mean of Univ.Credit Hours**

```{r}
myFTIC2016 <- myFTIC %>% filter(Cohort==2016  & Stu_DEMO_TIME_FRAME.y==201908)
gauge(round(mean(myFTIC2016$Stu_TotalUniversityHours), digits = 2), min=0, max=150, gaugeSectors(success= c(90,150), warning = c(90,90),danger=c(0,90), colors = c("green","yellow","red")))
```

### **FT2017 - Mean of Univ.Credit Hours**

```{r}

myFTIC2017 <- myFTIC %>% filter(Cohort==2017& Stu_DEMO_TIME_FRAME.y==201908)
gauge(round(mean(myFTIC2017$Stu_TotalUniversityHours), digits = 2), min=0, max=150, gaugeSectors(success= c(60,150), warning = c(60,60),danger=c(0,60), colors = c("green","yellow","red")))
```

### **FT2018 - Mean of Univ.Credit Hours**

```{r}
myFTIC2018 <- myFTIC %>% filter(Cohort==2018 & Stu_DEMO_TIME_FRAME.y==201908 )
gauge(round(mean(myFTIC2018$Stu_TotalUniversityHours), digits = 2), min=0, max=150, gaugeSectors(success= c(30,150), warning = c(30,30),danger=c(0,30), colors = c("green","yellow","red")))
```

### **FT2019 - Mean of Prior Univ. Hours**

```{r}
myFTIC2019 <- myFTIC %>% filter(Cohort==2019 & Stu_DEMO_TIME_FRAME.y==201908)
gauge(round(mean(myFTIC2019$Stu_TotalUniversityHours), digits = 2), min=0, max=150, gaugeSectors(success= c(0,150), warning = c(0,0),danger=c(0,0), colors = c("green","yellow","red")))
```


```{r}
FTIC_DEG <- read.csv("FTIC_ALL_APR_CRD_DEG.csv") %>% 
  filter(Cohort <= 2015)  %>%  
  group_by(STU_ID) %>%  
  arrange(NumFallTerms) %>% 
  #select(Stu_TotalUniversityHours, GPAAPRUWF,Deg_TermGranted,NumFallTerms,Stu_TotalInstGradePoints,Cohort) %>% 
  mutate(BDegreeEarnd=ifelse(Deg_TermGranted>0,1,0))   %>% 
  filter(NumFallTerms<=4) %>% 
  mutate(ClssificationYear=ifelse(NumFallTerms==1,"1stFall",ifelse(NumFallTerms==2,"2ndFall",ifelse(NumFallTerms==3,"3rdFall","4thFall")))) %>% 
mutate(GraduationTime = ifelse((Cohort==2013 & Deg_TermGranted <= 201705)|(Cohort==2014 & Deg_TermGranted <=201805)|(Cohort==2015 & Deg_TermGranted <=201905), "OnTime","OverTime" )) 
                                                              
FTIC_DEG[which(is.na(FTIC_DEG$GraduationTime)),"GraduationTime"] <- "Nodegree"
FTIC_DEG_lg <- FTIC_DEG %>% filter(GraduationTime=="OnTime"| GraduationTime=="Nodegree")
FTIC_DEG_lg$GraduationTime <- as.factor(FTIC_DEG_lg$GraduationTime)
FTIC_DEG_lg$Cohort <- as.factor(FTIC_DEG_lg$Cohort)

#p4 <- ggplot(FTIC_DEG[FTIC_DEG$BDegreeEarnd==0,], aes(x=Stu_TotalInstHours,y=Stu_TotalUniversityHours, color=Cohort))+geom_smooth()
#ggplotly(p4)

```


Column { data-with=550}
----------------------------------------------------------------------------------

### Probability of earning Bachelor Degree On-time by Academic Years Univ.credit Hours
```{r}
library(ggiraph)
library(ggiraphExtra)
library(plyr)
FTIC_DEG_lg$ClssificationYear <- as.factor(FTIC_DEG_lg$ClssificationYear)
fit2 <- glm(GraduationTime~Stu_TotalUniversityHours+ClssificationYear, data=FTIC_DEG_lg, family = "binomial")
pglm <- ggPredict(fit2, interactive=TRUE, colorn = 100, jitter = F)
pglm
pglm1 <- predict(fit2, FTIC_DEG_lg, type = "response")
pred1 <- ifelse(pglm1>0.5,1,0)
tab1 <- table(Predicted = pred1, Actural= FTIC_DEG_lg$GraduationTime)
Acc <- sum(diag(tab1))/sum(tab1)

```


### Notes

* This analysis considered FT 2013 to 2016 who earned a Bachelor degree or not within four years
* This analysis using Logistic Regression and the accuracy is `r Acc`






Top Courses Table with Subtotal
=================================================================

column
------------------------------------------------------------------

### Top Courses (Pass/Fail) for FTIC (2013 ~ 2019) by College

```{r}
#filter for FTIC crs 2013-2019(merge with FTIC_ALL_APR_CRDV1 ID)
#detach('package:dplyr')
#library(dplyr)
DWF_FTIC <- read_csv("FTIC_crs_all_deg.csv") %>% filter(Course_CrsNumber < 5000 ) %>% filter(Course_CrsGradeAwarded !="X" | Course_CrsGradeAwarded !="XX")  %>% 
  mutate(PASS = ifelse(Course_CrsGradeAwarded=="A"|Course_CrsGradeAwarded=="A-"|Course_CrsGradeAwarded=="B+"|Course_CrsGradeAwarded=="B"|
                                                                                                               Course_CrsGradeAwarded=="B-"|Course_CrsGradeAwarded=="C+"|                                                                         Course_CrsGradeAwarded=="C"|Course_CrsGradeAwarded=="C-"|
                                                                                                             Course_CrsGradeAwarded=="P"|Course_CrsGradeAwarded=="S","Pass","Fail")) %>% 
  select(Course_StuProgramCIPDesc,Course_CrsPrefix,Course_CrsNumber,Course_CrsGradeAwarded,Course_CrsCombined,Cohort="Cohort.x",Course_CrsTitle,Course_CrsSectionNumber,PASS,Course_StuCollege,Course_CrsCollege,Course_CrsDepartment)

rpivotTable(DWF_FTIC, aggregatorName = "Count as Fraction of Rows",
            rows = c("Course_CrsCollege"),
            cols = c("PASS","Cohort"),
            subtotals = T, rendererName = "Table With Subtotal")


```


Charts
===================================================================
Row 
-------------------------------------------------------------------

### Retention by Cohort (Fall 2019)

```{r}
detach("package:plyr")
library(dplyr)
p1 <- myFTIC %>% filter(Stu_DEMO_TIME_FRAME.y==201908) %>% 
  group_by(Cohort) %>% 
  summarise(count=n()) %>% 
  plot_ly(x=~Cohort, y=~count, color = rainbow(7), type="bar") %>% 
  layout(xaxis=list(title="Cohort"), yaxis=list(title='Counts in Fall 2019'))

  p1
```


### University hours by Fall Terms

```{r}
myFTIC$Cohort <- as.factor(myFTIC$Cohort)
p3 <- ggplot(myFTIC[myFTIC$Stu_DEMO_TIME_FRAME.y==201908,], mapping = aes(x=NumFallTerms, y= Stu_TotalUniversityHours))+
  geom_jitter(mapping = aes(color=Cohort))+
  geom_smooth()+
  annotate("rect", xmin=4.5 , xmax =7.9 , ymin=0 , ymax=225 , color="darkred", alpha=0.25)+
  annotate("text", x=6, y=220, label="Over Time")+
  ggtitle("Total Univ.Hours for Each Falls by Cohort (Fall 2019)")
ggplotly(p3)
```


Row
----------------------------------------------------------------------

### Trends of Graduation Numbers with Bachelor Degree for each Term

```{r}
library(forecast)
deg201901 <- read.csv("C:/Users/jsparks3/Downloads/drive-download-20200901T135512Z-001/STUDENT_DEGREE_SPRING20.csv")
C <- deg201901 %>% filter(Deg_Desc=="Bachelor")
 tbs <- table(C$Deg_TermGranted)
tbst <- t(tbs)
mytbst <- as.data.frame(tbst)
myts <- mytbst$Freq # numbers of degree
 forecast(myts) %>% hchart()

```

### University Credit Hours VS. UWF Credit Hours by Cohort (Fall 2019)

```{r}
A <-  myFTIC %>% group_by(STU_ID) %>% filter(NumFallTerms==max(NumFallTerms))
myhchart <- hchart(A[A$Stu_DEMO_TIME_FRAME.y==201908,], "scatter",
                   hcaes(x=Stu_TotalInstHours, y=Stu_TotalUniversityHours , group=Cohort))
myhchart
```





 Data Table
============================================================================


```{r}
library(readr)
library(dplyr)
library(DT)
FTIC_ALL_APR_CRD_DEG_firstDemo <- read_csv("FTIC_ALL_APR_CRD_DEG_firstV1.csv", 
    col_types = cols(X1 = col_skip(), STU_ID = col_skip(), 
        Stu_DEMO_TIME_FRAME = col_number(), 
        Stu_LastInstitutionDesc = col_skip())) %>% select(Cohort,Stu_AdmissionTermCode,Stu_College,Stu_Department,Deg_College,Deg_Department,Deg_TermGranted) 
FTIC_ALL_APR_CRD_DEG_firstDemo  <- FTIC_ALL_APR_CRD_DEG_firstDemo[!is.na(FTIC_ALL_APR_CRD_DEG_firstDemo$Deg_TermGranted),] %>% 
  mutate(GraduationTime = ifelse((Cohort==2013 & Deg_TermGranted <= 201705)|(Cohort==2014 & Deg_TermGranted <=201805)|(Cohort==2015 & Deg_TermGranted <=201905)|(Cohort==2016 & Deg_TermGranted <=202005)|(Cohort==2017 & Deg_TermGranted <=202105)|(Cohort==2018 & Deg_TermGranted <=202205), "OnTime","OverTime" )) 

colnames(FTIC_ALL_APR_CRD_DEG_firstDemo) <- c("Cohort","AdimissionTerm","Entry_college","Entry_depart","Deg_college","Deg_depart","Deg_term","GraduationTime")


datatable(FTIC_ALL_APR_CRD_DEG_firstDemo, extensions = "Buttons",caption = "Graduation Time Data",
          filter = "top",
          options = list(dom="Blfrtip",buttons=c("copy","csv","excel","pdf","print"), lengthMenu=list(c(10,25,50,-1), c(10,25,50,"All")),pageLength=25))



```

 Pivot Table
========================================================================

column 
------------------------------------------------------------------------
### Graduation On-time(<= 4) & Over-time(> 4) 


```{r}

FTIC_ALL_APR_CRD_DEG_firstV1 <- read_csv("FTIC_ALL_APR_CRD_DEG_firstV1.csv")
myFTICdegOnOvertime <- FTIC_ALL_APR_CRD_DEG_firstV1 %>%           select(Cohort,Stu_AdmissionTermCode,Stu_College,Stu_Department,Deg_College,Deg_Department,Deg_TermGranted,Deg_Term)  %>%
  mutate(GraduationTime = ifelse((Cohort==2013 & Deg_TermGranted <= 201705)|(Cohort==2014 & Deg_TermGranted <=201805)|(Cohort==2015 & Deg_TermGranted <=201905)|(Cohort==2016 & Deg_TermGranted <=202005)|(Cohort==2017 & Deg_TermGranted <=202105)|(Cohort==2018 & Deg_TermGranted <=202205), "OnTime","OverTime" )) 
colnames(myFTICdegOnOvertime) <- c("Cohort","AdimissionTerm","Entry_college","Entry_depart","Deg_college","Deg_depart","Deg_TermGranted","Deg_term","GraduationTime") 
myFTICdegOnOvertime[which(is.na(myFTICdegOnOvertime$GraduationTime)),"GraduationTime"] <- "Nodegree"


rpivotTable(myFTICdegOnOvertime, aggregatorName = "Count as Fraction of Columns",
            rows = c("Deg_college","GraduationTime"),
            
            cols = c("Cohort"),
           
            subtotals = T, rendererName = "Table With Subtotal")
           # vals = "Freq",
          #  rendererName = "Table")

```

Column 
----------------------------------------------------------------------------

### Table for APR and CRD
```{r}
library(data.table)
dtapr <- with(FTIC_ALL_APR_CRDV1[FTIC_ALL_APR_CRDV1$NumFallTerms>1,], table(NumFallTerms,Cohort,GPAAPRind))
dtaprdf <- data.frame(as.data.table(dtapr))
Cohortsize <- c(1316, 1386, 1329, 1285, 1075, 1095, 1076)
Cohort <- c(2013, 2014, 2015, 2016, 2017, 2018, 2019)
CS <- data.frame(Cohortsize,Cohort)
dtaprdf1 <- merge(CS,dtaprdf, by="Cohort",all.y=T) 
dtaprdf1$APRRate <- round(dtaprdf1$N/dtaprdf1$Cohortsize, digits = 4)
dtaprdf1[dtaprdf1$NumFallTerms==2,"NumFallTerms"] <- "2ndFall"
dtaprdf1[dtaprdf1$NumFallTerms==3,"NumFallTerms"] <- "3rdFall"
dtaprdf1[dtaprdf1$NumFallTerms==4,"NumFallTerms"] <- "4thFall"
dtaprdf1[dtaprdf1$NumFallTerms==5,"NumFallTerms"] <- "5thFall"
dtaprdf1[dtaprdf1$NumFallTerms==6,"NumFallTerms"] <- "6thFall"
dtaprdf1[dtaprdf1$NumFallTerms==7,"NumFallTerms"] <- "7thFall"
names(dtaprdf1) <- c("Cohort","Number_FTICSize","Number_Falls","APR","Number_APR","APRRate")
library(tidyverse)

dtaprdf2 <- dtaprdf1 %>% group_by(Cohort,Number_Falls,APR,Number_APR) %>% summarise(APRRate=round(Number_APR/Number_FTICSize,digits = 4))
datatable(dtaprdf2, extensions = "Buttons",caption = "APR Rate Data",
          filter = "top",
          options = list(dom="Blfrtip",buttons=c("copy","csv","excel","pdf","print"), lengthMenu=list(c(10,25,50,-1), c(10,25,50,"All")),pageLength=25))
```


 Summary - FTIC Overview in 2019 Fall Semester {data-orientation=columns}
========================================================================

Column
-----------------------------------------------------------------------


### Number of Enter FTIC size for 2019  

```{r}
valueBox(sum(myFTIC$Cohort==2019), caption = "Cohort size for 2019")
```

### Number of B-dgree Completition (On/Over-Time)

```{r}
valueBox(sum(myFTICdegOntime$Deg_term==201908))
```

### Number of APR Year 1 for FTIC 2018

```{r}
valueBox(count(FTIC_ALL_APR_CRDV1[(FTIC_ALL_APR_CRDV1$Cohort==2018 & FTIC_ALL_APR_CRDV1$NumFallTerms==2 & FTIC_ALL_APR_CRDV1$GPAAPRind=="GPAAPR"),"Cohort"]))

```

Column
-----------------------------------------------------------------------

```{r}
myFTICGPAAPR2018 <- myFTICGPAAPR %>% group_by(Cohort)  %>% summarise(meanGPA = mean(GPAAPRUWF))

```

Report

* This is a report based on total `r count(myFTIC[myFTIC$NumFallTerms==1, "Cohort"])` FTIC (2013 to 2019).

* The average of GPA year1 for FTIC 2018 is `r myFTICGPAAPR2018[myFTICGPAAPR2018$Cohort==2018,"meanGPA"]`.

* This report was created on `r format(Sys.Date(), format = "%B %d, %Y")`.

 Report 
========================================================================


