---
title: "Total University Hours-Change of Program:Biology"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
```


```{css, echo=FALSE}
.rpivottable{overflow-x: scroll; }
.DataTable{"scrollX": true}
.DataTable{"scrollY": true}

```

```{css, echo=FALSE}
.scroll-100 {
  max-height: 100px;
  overflow-y: auto;
  background-color: inherit;
}
```

Impact on Final University Hours among Changed Program - BIO
=============================================================



Column {.tabset .tabset-fade }
-----------------------------------------------------------------------

### What is the average of Final University Hours?


**Data Description**  

  + Entered biology from 2014 to 2016 FTIC
  + Earned bachelor degree
  + Total 305 FTIC 
  + Removed 4 FTIC had not information at their 1st year
  
**Classification**
  + Classification at the time when the changes happened
  
**Changed Index**
  + Stayed Program: stayed in entry program
  + Changed Program: changed program within biology
  + Changed department: changed department within HMCSE
  + Chnaged College: changed college
  

```{r}
LastTermInfo_all <- read.csv("G:/My Drive/#1_Jay's Space/Graduation-chaningMajorsApp/LastTermInfo_all.csv")


```


### Summary of Final University Hours (FUH)

```{r}
### summary on time
library(dplyr)
library(gtsummary)
LastTermInfo1_on <- LastTermInfo_all %>% filter(On_Time =="Yes")
univsummary <- tbl_summary(LastTermInfo_all[,c("On_Time","FinalUnivHrs","ChangeInd")], by=ChangeInd, 
                            statistic = list(all_continuous() ~ "{mean} ({sd})",
                                             all_categorical()~ "{n} /{N} ({p}%)"),
                                       digits = all_continuous() ~2,
                            #label = On_Time ~ "Graduation Time",
                            missing_text = "(Missing)")

univsummary
```
### Average FUH by Graudation Time

```{r message=FALSE, warning=FALSE}

unisummary2 <- LastTermInfo_all %>% group_by(On_Time, ChangeInd) %>% dplyr::summarise( MeanFinalUnivHrs=mean(FinalUnivHrs)) %>% tidyr::pivot_wider(names_from = "ChangeInd", values_from="MeanFinalUnivHrs") %>% 
  replace(is.na(.),0) 
knitr::kable( unisummary2)
```
### summary of Linear Regression

```{r}
#outliers
outliers <- boxplot(LastTermInfo_all$FinalUnivHrs, plot = FALSE)$out
nooutlieres <- LastTermInfo_all
nooutlieres <- nooutlieres[-which(nooutlieres$FinalUnivHrs %in% outliers),]


UniiHrs_lm <- aov(FinalUnivHrs ~  On_Time+FCCol_Class+ChangeInd + On_Time*Deg_Department, data = LastTermInfo_all )
summary(UniiHrs_lm)

UniiHrs_aov <- aov(FinalUnivHrs ~  On_Time+FCCol_Class+ChangeInd  , data = LastTermInfo_all )
summary(UniiHrs_aov)


UniiHrs <-lm(aov(FinalUnivHrs ~  On_Time+FCCol_Class+ChangeInd  , data = LastTermInfo_all ))
summary(UniiHrs)

```



### Tukey Test by Changed Index

```{r}
tukey <- TukeyHSD(UniiHrs_aov, "FCCol_Class")
tukey


```

### Box Plot for FUH by Changed Index

```{r}
library(ggplot2)
#mean(LastTermInfo_all$FinalUnivHrs)
ggplot(data=nooutlieres, aes( x= ChangeInd, y= FinalUnivHrs)) +
  geom_jitter(aes(color=ChangeInd)) +
  geom_boxplot(aes(color=ChangeInd), alpha=0.4)+
  theme()+
  labs(title = "Boxplot for FUH by Changed Groups", subtitle = "Entered BIO FTIC 2014 to 2016 (mean FUH: 136)")+
  geom_hline(yintercept = 135.982, linetype="dashed", color="blue")
```
### Tukey plot

```{r}
 plot(tukey, las=0, col="blue")
```

### Summary



