---
title: "R repeate plot"
output: html_notebook
---
This is repeated plot function practice
https://www.diegobarneche.com/2014-12-11-ufsc/lessons/40-repeating/index.html

### simplifying plots with plyr
```{r}
setwd("C:/Users/jsparks3/Documents/Data_IR")
FTIC_Fall_semester_all_CRS <- read.csv("~/Data_IR/FTIC_Fall_semester_all_CRS.csv")
colnames(FTIC_Fall_semester_all_CRS)
by_cohort_HSGPA_variables <- select(FTIC_Fall_semester_all_CRS, 2,4,5,8,12,20,27)
colnames(by_cohort_HSGPA_variables)
by_cohort_HSGPA_variables[1:5,]

#remove duplicated
by_cohort_HSGPA_variables_nodup <-  by_cohort_HSGPA_variables[!duplicated(by_cohort_HSGPA_variables$STU_ID),]
by_cohort_HSGPA_variables_nodup[1:20,]
by_cohort_HSGPA_variables_nodup[which(is.na(by_cohort_HSGPA_variables_nodup$Stu_GPAHighSchool)), "Stu_GPAHighSchool"] <- 0
by_cohort_HSGPA_variables_nodup[1:20,]

#return !=zero values
by_cohort_HSGPA_variables_nodupna <- by_cohort_HSGPA_variables_nodup[by_cohort_HSGPA_variables_nodup$Stu_GPAHighSchool !=0,]
by_cohort_HSGPA_variables_nodupna[20:30,]
```
### using HS GPA and first term total grade points

```{r}

by_cohrot_GPAHS <- by_cohort_HSGPA_variables_nodupna %>% group_by(Cohort.x) %>% summarise(meanHSGPA=mean(Stu_GPAHighSchool), count=n());by_cohrot_GPAHS


```

### CSE GPA HS data need to import to get FTIC 2018:2020 GPAHS information and ist term GPA

```{r}
CSEdata_copy <- CSEdata
CSEdata_copy[which(is.na(CSEdata_copy$GPA_HIGHSCHOOL)),]
CSEdata_copy[which(CSEdata_copy$GPA_HIGHSCHOOL<=3.3 & CSEdata$APPLICANT_TIER ==3),] # no zeros
CSEdata_copy[which(CSEdata$COUNTY=="Non-Florida" & CSEdata$HIGH_SCHOOL_NAME=="Non-Florida High School"),] # no zeros
CSEdata_copy[which(is.na(CSEdata$ACT_PROPORTION)),"ACT_PROPORTION"] <- 0 # no zeros
CSEdata_copy[which(is.na(CSEdata$SAT_PROPORTION)),"SAT_PROPORTION"] <- 0 # no zeros
CSEdata_copy[which(is.na(CSEdata_copy$GPA_HIGHSCHOOL)), "GPA_HIGHSCHOOL"] <- 0
CSEdata_copy[CSEdata_copy$GPA_HIGHSCHOOL==0, "GPA_HIGHSCHOOL"] <- NA

CSEdata_copy[which(is.na(CSEdata_copy$GPA_HIGHSCHOOL)),]


```
```{r}

CSEdata_copy$Standard_test_pop <-  pmax(CSEdata$ACT_PROPORTION, CSEdata$SAT_PROPORTION, na.rm = T)
CSEdata_copy[1:20, 45:57]
CSEdata_copy[which(is.na(CSEdata_copy$Standard_test_pop)),"Standard_test_pop"]
summary(CSEdata_copy)

```
```{r}
CSEdata_copy$Cohort <- stringr::str_extract(CSEdata_copy$COHORT_YEAR, "^.{4}")
CSEdata_copy[1:10,]

```
```{r}
xtabs(~CSEdata_copy$Cohort +CSEdata_copy$APPLICANT_TIER +CSEdata_copy$APR)
```



```{r}
CSEdata_copy_HSGPA <- CSEdata_copy %>% group_by(Cohort) %>% summarise(meanHSGPA=mean(GPA_HIGHSCHOOL, na.rm = T), meanGPAT1=mean(FIRST_FALL_GPA, na.rm = T), count=n());CSEdata_copy_HSGPA
```

```{r}

HSGPA_by_cohort <- function(ETHNICITY,CSEdata_copy,base_cohort=2015){
  dsub <- CSEdata_copy[CSEdata_copy$ETHNICITY=="White", c("Cohort","GPA_HIGHSCHOOL")]
  dsub$GPA_rel <-  dsub$GPA_HIGHSCHOOL/dsub$GPA_HIGHSCHOOL[dsub$Cohort == base_cohort]
  dsub
}

plot()



```





```{r}
CRSCSEData <- read.csv("~/Data_IR/CRSCSEData.csv")
unique(CRSCSEData$GRADE_AWARDED)#"F"  "B-" "C-" "C+" "W"  "D"  "B+" "C"  "A"  "A-" "I*" "B"  "S"  "D+" "NF" "XX" "I"  "U"  "P"  "NG" NA 
CRSCSEData_copy <- CRSCSEData
CRSCSEData_copy$GradePoints <- CRSCSEData_copy$GRADE_AWARDED
CRSCSEData_copy[which(is.na(CRSCSEData_copy$GradePoints)), "GradePoints"] <- 99
CRSCSEData_copy[CRSCSEData_copy$GradePoints=="A", "GradePoints"] <- 4.0
CRSCSEData_copy[CRSCSEData_copy$GradePoints=="A-", "GradePoints"] <- 3.70
CRSCSEData_copy[CRSCSEData_copy$GradePoints=="B+", "GradePoints"] <- 3.30
CRSCSEData_copy[CRSCSEData_copy$GradePoints=="B", "GradePoints"] <- 3.00
CRSCSEData_copy[CRSCSEData_copy$GradePoints=="B-", "GradePoints"] <- 2.70
CRSCSEData_copy[CRSCSEData_copy$GradePoints=="C", "GradePoints"] <- 2.00
CRSCSEData_copy[CRSCSEData_copy$GradePoints=="C+", "GradePoints"] <- 2.30
CRSCSEData_copy[CRSCSEData_copy$GradePoints=="C-", "GradePoints"] <- 1.70
CRSCSEData_copy[CRSCSEData_copy$GradePoints=="D", "GradePoints"] <- 1.00
CRSCSEData_copy[CRSCSEData_copy$GradePoints=="D+", "GradePoints"] <- 1.30
CRSCSEData_copy[CRSCSEData_copy$GradePoints=="F", "GradePoints"] <- 0.00
CRSCSEData_copy[CRSCSEData_copy$GradePoints=="NF", "GradePoints"] <- 0.00
CRSCSEData_copy[CRSCSEData_copy$GradePoints=="U", "GradePoints"] <- 0.00
CRSCSEData_copy[CRSCSEData_copy$GradePoints=="NG", "GradePoints"] <- 0.00

attach(CRSCSEData_copy)
my_no_grade <- c("S","P","XX","I","I*","W")
CRSCSEData_copy$GradePoints <- sapply(CRSCSEData_copy[,8], function(x) replace(x, x %in% c("S","P","XX","I","I*","W"), 99))
 CRSCSEData_copy$GradePoints <- as.numeric(CRSCSEData_copy$GradePoints)
  CRSCSEData_copy$STU_SECTN_CRED <- as.numeric(CRSCSEData_copy$STU_SECTN_CRED)

# TGP calculation
CRSCSEData_copy_GPAGrades <- CRSCSEData_copy %>% filter(GradePoints<99) # filter out NA
CRSCSEData_copy_GPAGrades$TGP_each <-  CRSCSEData_copy_GPAGrades$GradePoints*CRSCSEData_copy_GPAGrades$STU_SECTN_CRED
CRSCSEData_copy_GPAGrades.TGP_all <- CRSCSEData_copy_GPAGrades %>%group_by(UNIV_ROW_ID)   %>%  summarise(TGP_all=sum(TGP_each),TGP_credit=sum(STU_SECTN_CRED) ,GPAY1=TGP_all/TGP_credit,.groups="drop");CRSCSEData_copy_GPAGrades.TGP_all
#checkc
CSEdata[CSEdata$UNIV_ROW_ID==970295713,]
CRSCSEData_copy_GPAGrades.TGP_all[CRSCSEData_copy_GPAGrades.TGP_all$UNIV_ROW_ID==970295713,]
CRSCSEData[CRSCSEData$UNIV_ROW_ID==970295713,]
sum(is.na(CRSCSEData_copy_GPAGrades.TGP_all$TGP_all))

```
# get HS GPA and TGP 
## find the difference btw GPA_ENTERING_SECOND_FALL and GPAY1

```{r}
CSEdata_copy_HSGPA_variables <- CSEdata_copy[CSEdata_copy$Cohort<= 2019,] %>% select(UNIV_ROW_ID,GPA_HIGHSCHOOL,Cohort,GPA_ENTERING_SECOND_FALL,APR,ATTENDED_SECOND_FALL_TERM,Standard_test_pop );head(CSEdata_copy_HSGPA_variables)

CSE_GPA_HS_TGP <- merge(CSEdata_copy_HSGPA_variables,CRSCSEData_copy_GPAGrades.TGP_all, by="UNIV_ROW_ID", all.x=T)
CSE_GPA_HS_TGP[which(is.na(CSE_GPA_HS_TGP$GPA_HIGHSCHOOL)),]
CSE_GPA_HS_TGP_nona <- CSE_GPA_HS_TGP[complete.cases(CSE_GPA_HS_TGP),]
CSE_GPA_HS_TGP_nona[CSE_GPA_HS_TGP_nona$UNIV_ROW_ID==970295713,]
CSE_GPA_HS_TGP_nona$diffGPA <- round(CSE_GPA_HS_TGP_nona$GPA_ENTERING_SECOND_FALL-CSE_GPA_HS_TGP_nona$GPAY1, digits = 4) # there is difference APR GPA and raw GPA Y1
CSE_GPA_HS_TGP_nona[CSE_GPA_HS_TGP_nona$GPA_ENTERING_SECOND_FALL<= 2.0,]

```
```{r}
CSE_GPA_HS_TGP_nona[CSE_GPA_HS_TGP_nona$GPA_ENTERING_SECOND_FALL<= 2.0 & CSE_GPA_HS_TGP_nona$GPAY1>=2.0 & CSE_GPA_HS_TGP_nona$ATTENDED_SECOND_FALL_TERM=="Yes",]#970401032 #970541362 grade I is included in this case and  970513055 early graduated




```
# check the diff GPAY1 and APRGPA

```{r}
CSE_GPA_HS_TGP_nona[CSE_GPA_HS_TGP_nona$diffGPA != 0,]
```
```{r}
CES_GPA <- CSE_GPA_HS_TGP_nona %>% group_by(Cohort) %>% summarise(count=n());CES_GPA
CSE_GPA_HS_TGP_nona$APR <- as.factor(CSE_GPA_HS_TGP_nona$APR)
summary(CSE_GPA_HS_TGP_nona[CSE_GPA_HS_TGP_nona$GPA_ENTERING_SECOND_FALL>0,])
```


### mean values for Credits ==28(median) , APR & find the growth GPA rate base on Cohort == mean(2015) & CSE_GPA_HS_TGP_nona$APR=="Yes" 
```{r}
mean_f <- CSE_GPA_HS_TGP_nona[CSE_GPA_HS_TGP_nona$TGP_credit == 28 & CSE_GPA_HS_TGP_nona$APR=="Yes" ,] %>%  group_by(Cohort)%>%  summarise(meanHSGPA=mean(GPA_HIGHSCHOOL), meanTGP=mean(TGP_all), meanGPAY1=mean(GPAY1), meanCredit=mean(TGP_credit),meanGPAY1APR=mean(GPA_ENTERING_SECOND_FALL),meanTEST=mean(Standard_test_pop),count=n(), .groups="drop");mean_f

# find growth
mean_f_growth <- mean_f %>% summarise(growthHSGPA=round(mean_f$meanHSGPA/3.613256-1, digits = 6), growthTGP=round(mean_f$meanTGP/84.38023-1, digits = 6),growthGPAAPRY1=round((mean_f$meanGPAY1APR/3.041989)-1, digits = 6),growthTEST1=round(mean_f$meanTEST/0.5898980-1, digits = 6) ,growthGPAY1=round(mean_f$meanGPAY1/3.013580-1, digits = 6));mean_f_growth
# using GPA year1(rawdata)
```

```{r}
hist(CSE_GPA_HS_TGP_nona$GPA_HIGHSCHOOL)
hist(CSE_GPA_HS_TGP_nona$TGP_all)
hist(CSE_GPA_HS_TGP_nona$GPAY1)



```
### Normalixation function
```{r}
glimpse(CSE_GPA_HS_TGP_nona)
CSE_GPA_HS_TGP_nona <- CSE_GPA_HS_TGP_nona[c(1,2,3,5,6,4,7,8,9,10)]
  head(CSE_GPA_HS_TGP_nona)


range_normal <- function(x){
  (x-min(x))/(max(x)-min(x))
}

# apply
HS_GPA_normal <- as.data.frame(lapply(CSE_GPA_HS_TGP_nona[2:4], range_normal))
names(HS_GPA_normal) <- c("Normal_HSGPA","Normal_GPAY1APR","NormalTGPY1")
head(HS_GPA_normal)

```
# using 2015 HS GPA ==3.582532	
```{r}



```





```{r}

based_2015GPA <- function(x,CSE_GPA_HS_TGP_nona,base_cohort=2015 ){
  fsub <- CSE_GPA_HS_TGP_nona[CSE_GPA_HS_TGP_nona$Cohort == Cohort, c("GPA_HIGHSCHOOL","GPAY1")]
  fsub$GPAgrowth <- (fsub$GPA_HIGHSCHOOL-mean[fsub$GPA_HIGHSCHOOL & fsub$Cohort== base_cohort])*100/mean[fsub$GPA_HIGHSCHOOL & fsub$Cohort== base_cohort]
  fsub
 # ((x-3.5825)/3.5825)*100
}

HS_GPA_growth <- as.data.frame(lapply(CSE_GPA_HS_TGP_nona[2], CSE_GPA_HS_TGP_nona, based_2015GPA))






```

















This is an example of exponential growth function in r
```{r}
x = 1:100
y= x^3

fit = lm(log(y )~x)
plot(x,y)
lines(x, exp(fitted(fit)), col="blue")
```
