---
title: "Graduation Analysis by Program Change"
output: 
 
 flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
 
---


```{r setup, include=FALSE}
library(flexdashboard)
library(gtsummary)
library(readr)
library(dplyr)
metric4anon <- read_csv("G:/Shared drives/HMCSE-PAM Lab/DATA 201901/metric4anon.csv")

```


```{css, echo=FALSE}
.rpivottable{overflow-x: scroll; }
.DataTable{"scrollX": true}
.DataTable{"scrollY": true}

```

```{css, echo=FALSE}
.scroll-100 {
  max-height: 100px;
  overflow-y: auto;
  background-color: inherit;
}
```



Column {.tabset .tabset-fade }
-----------------------------------------------------------------------



### Data Description

* Cohort: 2014 to 2016
* College: HMCSE
* Department: Biology




### Numer of Entered in Biology  from 2014 to 2016

```{r echo=FALSE, message=FALSE, warning=FALSE}

FTIC_201416_BIO <- metric4anon %>% filter((Cohort == 2014 |Cohort ==  2015 | Cohort == 2016) & Stu_DepartmentBegin=="Biology") %>% select(STU_ID,Cohort,Stu_DepartmentBegin)

Total_B_cohort <- tbl_cross(FTIC_201416_BIO,row=Cohort, col=Stu_DepartmentBegin   ,percent = "row")
Total_B_cohort
```



```{r message=FALSE, warning=FALSE, include=FALSE}
### enc
STU_ENROLLMENT_SPRING21 <- read.csv("G:/Shared drives/HMCSE-PAM Lab/DATA 202101/202101/STU_ENROLLMENT_SPRING21.csv") %>% filter(Stu_DEMO_TIME_FRAME >= 201408 & Stu_DEMO_DATA_SOURCE=="SIF") %>% filter(Stu_ClassificationCode <= 4) 

### deg
STUDENT_DEGREE_SPRING21 <- read.csv("G:/Shared drives/HMCSE-PAM Lab/DATA 202101/202101/STUDENT_DEGREE_SPRING21.csv") %>% select(STU_ID, contains("granted"),Deg_Desc,Deg_BannerProgramCode,contains("program"),contains("college"),contains("departm")) %>% filter(Deg_Desc=="Bachelor") %>% group_by(STU_ID) %>% filter(Deg_TermGranted==max(Deg_TermGranted)) %>% top_n(1, Deg_BannerProgram)

### datashaping
FTIC_201416_BIO_enc <- merge(STU_ENROLLMENT_SPRING21,FTIC_201416_BIO, by="STU_ID", all.y = T) %>% 
  select(STU_ID,Cohort,Stu_DepartmentBegin,Stu_DEMO_TIME_FRAME,Stu_Term,contains("classification") ,contains("college"),contains("department"),contains("program") ,contains("total"))

FTIC_201416_BIO_enc_deg <- merge(FTIC_201416_BIO_enc,STUDENT_DEGREE_SPRING21, by="STU_ID", all.x = T) %>% group_by(STU_ID) %>% 
  arrange(Stu_DEMO_TIME_FRAME)
FTIC_201416_BIO_enc_deg #674
```

### FTIC that earned Bachelor Degree

```{r echo=FALSE, message=FALSE, warning=FALSE}
FTIC_201416_BIO_enc_deg_B <- FTIC_201416_BIO_enc_deg %>% filter(Deg_Desc == "Bachelor") %>% 
  mutate(Gradu_Time=ifelse((Cohort==2014 & Deg_TermGranted <= 201805), "Gradu<=4",
                         ifelse((Cohort==2014 & (Deg_TermGranted > 201805 & Deg_TermGranted <= 201905)),"Gradu=5",
                                ifelse((Cohort==2014 & (Deg_TermGranted > 201905 & Deg_TermGranted <= 202005)),'Gradu=6',
                                   ifelse((Cohort==2014 & (Deg_TermGranted > 202005 & Deg_TermGranted <= 202008)),'Gradu=7',    
                                       ifelse((Cohort==2015 & Deg_TermGranted <= 201905), "Gradu<=4",
                         ifelse((Cohort==2015 & (Deg_TermGranted > 201905 & Deg_TermGranted <= 202005)),"Gradu=5",
                                ifelse((Cohort==2015 & (Deg_TermGranted > 202005 & Deg_TermGranted <= 202105)),'Gradu=6',
                                       ifelse((Cohort==2016 & Deg_TermGranted <= 202005), "Gradu<=4",
                         ifelse((Cohort==2016 & (Deg_TermGranted > 202005 & Deg_TermGranted <= 202105)),"Gradu=5",
                                ifelse((Cohort==2016 & (Deg_TermGranted > 202105 & Deg_TermGranted <= 202205)),'Gradu=6',"Others"))))))))))) %>%  mutate(EarnedBDeg=rep("Bachelor"))

#length(unique(FTIC_201416_BIO_enc_deg_B$STU_ID)) #305
FTIC_bachelor <- FTIC_201416_BIO_enc_deg_B %>% group_by(STU_ID) %>% top_n(1, Stu_DEMO_TIME_FRAME)
tbl_cross(FTIC_bachelor,  row=Cohort ,col=Gradu_Time, percent = "row")

#others <- FTIC_201416_BIO_enc_deg_B[FTIC_201416_BIO_enc_deg_B$On_Time =="Others",];others for gradu=7 FT14
```

### FTIC that continue in Fall 2020 (Non-Bachelor Degree)

```{r echo=FALSE}
FTIC_201416_BIO_enc_deg_notB <- FTIC_201416_BIO_enc_deg[which(is.na(FTIC_201416_BIO_enc_deg$Deg_Desc)),] %>% group_by(STU_ID) %>% top_n(1,Stu_DEMO_TIME_FRAME) %>% filter(Stu_DEMO_TIME_FRAME==202008)
#length(unique(FTIC_201416_BIO_enc_deg_notB$STU_ID))#3369
drop_term <- tbl_cross(FTIC_201416_BIO_enc_deg_notB, row= Stu_DEMO_TIME_FRAME, col=Cohort, percent = "col")
drop_term

```



```{r message=FALSE, warning=FALSE, include=FALSE}

#### CODE
library(dplyr)
library(gtsummary)
FTIC_201416_BIO_enc_deg_B_changed <- FTIC_201416_BIO_enc_deg_B %>%  group_by(STU_ID) %>%
  mutate( ChangedProg=ifelse(Stu_ProgramCIPCode == lag(Stu_ProgramCIPCode, default = Stu_ProgramCIPCode[1]), 0,1)) %>% 
  mutate(ChangedDeprt=ifelse(Stu_DepartmentCode == lag(Stu_DepartmentCode, default = Stu_DepartmentCode[1]), 0,1)) %>% 
  mutate(ChangedCol=ifelse(Stu_College == lag(Stu_College, default = Stu_College[1]), 0,1)) 

### Total numbers of changed
Repeated <- FTIC_201416_BIO_enc_deg_B_changed %>% select(STU_ID, ChangedProg,ChangedDeprt,ChangedCol ) %>% group_by(STU_ID) %>% 
  dplyr::summarise(TotalChangedProg=sum(ChangedProg),TotalChangedDepart=sum(ChangedDeprt),TotalChangedCol=sum(ChangedCol))
 

### merged with repeated time
FTIC_201416_BIO_enc_deg_B_changed_total <- merge(FTIC_201416_BIO_enc_deg_B_changed, Repeated, by="STU_ID", all.x = T) %>% group_by(STU_ID) %>% arrange(Stu_DEMO_TIME_FRAME) %>%  group_by(STU_ID) %>% mutate(On_Time=ifelse(Gradu_Time=="Gradu<=4", "Yes", "No")) 

#xtabs(~FTIC_201416_BIO_enc_deg_B_changed_total$On_Time+FTIC_201416_BIO_enc_deg_B_changed_total$Stu_DEMO_TIME_FRAME+FTIC_201416_BIO_enc_deg_B_changed_total$Cohort)


## 1st Term GPA

#glimpse(FTIC_201416_BIO_enc_deg_B_changed_total)
```


###  Numebr of ProgramCIPcode Changes 

```{r}

### Numers of changes and graduation time 
Total_gradu_time <- FTIC_201416_BIO_enc_deg_B_changed_total %>% group_by(STU_ID) %>% top_n(1, Stu_DEMO_TIME_FRAME)


#changed_6times <- FTIC_201416_BIO_enc_deg_B_changed_total[FTIC_201416_BIO_enc_deg_B_changed_total$STU_ID=="7E4B110FD728E884B9389D8BBC408541A5CDD649",];changed_6times


library(gtsummary)
tbl_cross(Total_gradu_time,row = TotalChangedProg, col= Gradu_Time, percent = "col")
```

###  Numebr of Department Changes

```{r}
tbl_cross(Total_gradu_time,row = TotalChangedDepart, col= Gradu_Time, percent = "col")

```

### Number of College Changes

```{r echo=FALSE}
tbl_cross(Total_gradu_time,row = TotalChangedCol, col= Gradu_Time, percent = "col")

```

### Enter Program and End Program among Earned Bachelor Degree

```{r echo=FALSE}

enter_program_end_program <- FTIC_201416_BIO_enc_deg_B_changed_total %>% group_by(STU_ID) %>% filter(Stu_DEMO_TIME_FRAME == min(Stu_DEMO_TIME_FRAME))%>% 
  tbl_cross(row=Stu_BannerProgramDesc , col=Deg_BannerProgram, percent = "row")
enter_program_end_program

```


### Graduation Time and End Department among Earned Bachelor Degree

```{r echo=FALSE}

end_depart <- FTIC_201416_BIO_enc_deg_B_changed_total %>% group_by(STU_ID) %>% filter(Stu_DEMO_TIME_FRAME == min(Stu_DEMO_TIME_FRAME))%>% 
  tbl_cross(row=On_Time  , col=Deg_Department, percent = "col")
end_depart
```

### Enter College and End College among Earned Bachelor Degree


```{r echo=FALSE}
#glimpse(FTIC_201416_BIO_enc_deg_B_changed_total)

end_coll_gradutime <- FTIC_201416_BIO_enc_deg_B_changed_total %>% group_by(STU_ID) %>% filter(Stu_DEMO_TIME_FRAME == min(Stu_DEMO_TIME_FRAME))%>% 
  tbl_cross(row=Deg_College , col=On_Time, percent = "row")
end_coll_gradutime
```

### 1st time change of Major by FTIC Classification

```{r}


#unique(FTIC_201416_BIO_enc_deg_B_changed_total$Stu_ProgramCIPDesc)# #86 #47

### 1st time change of program cip
Change1st_progCIP <- FTIC_201416_BIO_enc_deg_B_changed_total  %>% group_by(STU_ID) %>% top_n(1, ChangedProg ) %>% mutate(ChangedProg_Ind=ifelse(ChangedProg>0, "Yes","No"))%>% filter(Stu_DEMO_TIME_FRAME == min(Stu_DEMO_TIME_FRAME)) %>% select(STU_ID, "FCProg_Class"=Stu_Classification , "FCProg_UniHrs"=Stu_TotalUniversityHours, "FCProg_UWFHrs"=Stu_TotalInstHours, "FCPorg_TGP"=Stu_TotalInstGradePoints, ChangedProg_Ind) %>%  mutate(FCProg_GPA=round(FCPorg_TGP/FCProg_UWFHrs,digits = 4))#305

### on-time and FCProg
Change1st_progCIP_ontime <- FTIC_201416_BIO_enc_deg_B_changed_total  %>% group_by(STU_ID) %>% top_n(1, ChangedProg ) %>% mutate(ChangedProg_Ind=ifelse(ChangedProg>0, "Yes","No"))%>% filter(Stu_DEMO_TIME_FRAME == min(Stu_DEMO_TIME_FRAME)) %>% select(STU_ID, "FCProg_Class"=Stu_Classification , "FCProg_UniHrs"=Stu_TotalUniversityHours, "FCProg_UWFHrs"=Stu_TotalInstHours, "FCPorg_TGP"=Stu_TotalInstGradePoints, ChangedProg_Ind,On_Time) %>%  mutate(FCProg_GPA=round(FCPorg_TGP/FCProg_UWFHrs,digits = 4))

### cross tab with p-value
tbl_cross(Change1st_progCIP_ontime, row=FCProg_Class, col=ChangedProg_Ind, percent ="cell" ) %>% add_p()

### merge with total data
FTIC_201416_BIO_enc_deg_B_changed_total_FCProg <- merge(FTIC_201416_BIO_enc_deg_B_changed_total, Change1st_progCIP, by="STU_ID", all.x = T )




```

### 1st time change of Department by FTIC Classification

```{r echo=FALSE}
### on-time and FCDeprt
Change1st_Deprt_ontime <- FTIC_201416_BIO_enc_deg_B_changed_total  %>% group_by(STU_ID) %>% top_n(1, ChangedDeprt )   %>% mutate(ChangedDeprt_Ind=ifelse(ChangedProg>0, "Yes","No"))%>% filter(Stu_DEMO_TIME_FRAME == min(Stu_DEMO_TIME_FRAME)) %>% select(STU_ID, "FCDepart_Class"=Stu_Classification , "FCDepart_UniHrs"=Stu_TotalUniversityHours, "FCDepart_UWFHrs"=Stu_TotalInstHours, "FCDepart_TGP"=Stu_TotalInstGradePoints,ChangedDeprt_Ind, On_Time) %>%  mutate(FCDepart_GPA=round(FCDepart_TGP/FCDepart_UWFHrs,digits = 4))

### cross tab with p-value
tbl_cross(Change1st_Deprt_ontime, row=FCDepart_Class, col=ChangedDeprt_Ind, percent ="cell" ) %>% add_p()


### 1st time change of department

Change1st_Deprt <- FTIC_201416_BIO_enc_deg_B_changed_total  %>% group_by(STU_ID) %>% top_n(1, ChangedDeprt ) %>% mutate(ChangedDeprt_Ind=ifelse(ChangedProg>0, "Yes","No"))%>% filter(Stu_DEMO_TIME_FRAME == min(Stu_DEMO_TIME_FRAME)) %>% select(STU_ID, "FCDepart_Class"=Stu_Classification , "FCDepart_UniHrs"=Stu_TotalUniversityHours, "FCDepart_UWFHrs"=Stu_TotalInstHours, "FCDepart_TGP"=Stu_TotalInstGradePoints,ChangedDeprt_Ind) %>%  mutate(FCDepart_GPA=round(FCDepart_TGP/FCDepart_UWFHrs,digits = 4))

### merge with total data set: FTIC_201416_BIO_enc_deg_B_changed_total_FCProg
FTIC_201416_BIO_enc_deg_B_changed_total_FCProg_FCDepart <- merge(FTIC_201416_BIO_enc_deg_B_changed_total_FCProg, Change1st_Deprt, by="STU_ID", all.x = T )

```


### 1st time change of College by FTIC Classification

```{r echo=FALSE}
### on-time and FCDeprt
Change1st_col_ontime <- FTIC_201416_BIO_enc_deg_B_changed_total  %>% group_by(STU_ID) %>% top_n(1, ChangedCol ) %>% mutate(ChangedCol_Ind=ifelse(ChangedCol>0, "Yes","No")) %>%  filter(Stu_DEMO_TIME_FRAME == min(Stu_DEMO_TIME_FRAME)) %>% select(STU_ID, "FCCol_Class"=Stu_Classification , "FCCol_UniHrs"=Stu_TotalUniversityHours, "FCCol_UWFHrs"=Stu_TotalInstHours, "FCCol_TGP"=Stu_TotalInstGradePoints, On_Time,ChangedCol_Ind) %>%  mutate(FCDepart_GPA=round(FCCol_TGP/FCCol_UWFHrs,digits = 4))

### cross tab with p-value
tbl_cross(Change1st_col_ontime, row=FCCol_Class, col=ChangedCol_Ind, percent ="cell" ) %>% add_p()


### 1st time change of department

Change1st_Col <-FTIC_201416_BIO_enc_deg_B_changed_total  %>% group_by(STU_ID) %>% top_n(1, ChangedCol ) %>% mutate(ChangedCol_Ind=ifelse(ChangedCol>0, "Yes","No"))%>% filter(Stu_DEMO_TIME_FRAME == min(Stu_DEMO_TIME_FRAME)) %>% select(STU_ID, "FCCol_Class"=Stu_Classification , "FCCol_UniHrs"=Stu_TotalUniversityHours, "FCCol_UWFHrs"=Stu_TotalInstHours, "FCCol_TGP"=Stu_TotalInstGradePoints,ChangedCol_Ind) %>%  mutate(FCCol_GPA=round(FCCol_TGP/FCCol_UWFHrs,digits = 4))

### merge with total data set: FTIC_201416_BIO_enc_deg_B_changed_total_FCProg
FTIC_201416_BIO_enc_deg_B_changed_total_FCProg_FCDepart_FCCol <- merge(FTIC_201416_BIO_enc_deg_B_changed_total_FCProg_FCDepart, Change1st_Col, by="STU_ID", all.x = T )

```

### Variables

```{r echo=FALSE, message=FALSE, warning=FALSE}
#glimpse(FTIC_201416_BIO_enc_deg_B_changed_total_FCProg_FCDepart_FCCol)

### start and end progrma
Start_End_Prog <- FTIC_201416_BIO_enc_deg_B_changed_total_FCProg_FCDepart_FCCol %>% group_by(STU_ID) %>% filter(Stu_DEMO_TIME_FRAME==min(Stu_DEMO_TIME_FRAME)) %>% select(STU_ID, "Start_Prog"=Stu_ProgramCIPDesc, "End_Prog"=Deg_ProgramCIPDesc)

### Year 1 UNIV UWF Hrs
library(stringr)
Year1_CRDInfo <- FTIC_201416_BIO_enc_deg_B_changed_total_FCProg_FCDepart_FCCol %>% group_by(STU_ID) %>% 
  filter(str_detect(Stu_Term, "Fall...")) %>% group_by(Cohort, STU_ID) %>% 
  mutate(Fall_IND=ifelse((Cohort==2014 & Stu_DEMO_TIME_FRAME==201508),"Year1",
                         ifelse((Cohort==2015 & Stu_DEMO_TIME_FRAME==201608),"Year1",
                                ifelse((Cohort==2016 & Stu_DEMO_TIME_FRAME==201708),"Year1","Others")))) %>% 
  filter(Fall_IND=="Year1") %>% select(STU_ID, Fall_IND,"Year1UniHrs"= Stu_TotalUniversityHours, "Year1UWFHrs"=Stu_TotalInstHours )

### start Program
 enter_Prog <- FTIC_201416_BIO_enc_deg_B_changed_total_FCProg_FCDepart_FCCol %>% group_by(STU_ID) %>% 
   filter(Stu_DEMO_TIME_FRAME==min(Stu_DEMO_TIME_FRAME)) %>% select(STU_ID, "Start_Prog" = Stu_ProgramCIPDesc )

 FTIC_201416_BIO_enc_deg_B_changed_total_FCProg_FCDepart_FCCol <- merge(FTIC_201416_BIO_enc_deg_B_changed_total_FCProg_FCDepart_FCCol,enter_Prog, by="STU_ID", all.x = T)
 
### merge with data 
Year1_CRDInfo_ChangeProg<- merge(FTIC_201416_BIO_enc_deg_B_changed_total_FCProg_FCDepart_FCCol,Year1_CRDInfo, by="STU_ID", all.y = T ) %>% group_by(STU_ID)%>%  filter(Stu_TotalUniversityHours==max(Stu_TotalUniversityHours))

#glimpse(Year1_CRDInfo_ChangeProg)

Year1_CRDInfo_ChangeProg1 <- Year1_CRDInfo_ChangeProg[!duplicated(Year1_CRDInfo_ChangeProg$STU_ID),] %>% 
  select(On_Time,"Enter_Prog"=Start_Prog,"End_Prog"=Deg_ProgramCIPDesc,FCProg_Class,TotalChangedProg,TotalChangedDepart,TotalChangedCol,Year1UWFHrs) 


Year1_CRDInfo_ChangeProg12 <- Year1_CRDInfo_ChangeProg1[, -1]
colnames(Year1_CRDInfo_ChangeProg12)
#tbl_cross(Year1_CRDInfo_ChangeProg1, row=End_Prog, col=Enter_Prog, percent = "row")
```

###  Methods

* Using Naive Bayes Classification

```{r include=FALSE}
glimpse(Year1_CRDInfo_ChangeProg12)
Year1_CRDInfo_ChangeProg12$On_Time <- factor(Year1_CRDInfo_ChangeProg12$On_Time)
Year1_CRDInfo_ChangeProg12$Enter_Prog <- factor(Year1_CRDInfo_ChangeProg12$Enter_Prog)
Year1_CRDInfo_ChangeProg12$FCProg_Class <- factor(Year1_CRDInfo_ChangeProg12$FCProg_Class)
Year1_CRDInfo_ChangeProg12$End_Prog <- factor(Year1_CRDInfo_ChangeProg12$End_Prog)

library(caret)
set.seed(123)
TrainingIndex <- createDataPartition(Year1_CRDInfo_ChangeProg12$On_Time, p=0.8, list = FALSE)
TrainingSet <- Year1_CRDInfo_ChangeProg12[TrainingIndex,] # Training Set
TestingSet <- Year1_CRDInfo_ChangeProg12[-TrainingIndex,]
```


### Naive Bayes Classification

```{r}
library(naivebayes)
set.seed(123)
ChangeMajorNB <- naive_bayes(On_Time ~ ., data = TrainingSet, usekernel = TRUE)

confusionMatrix(predict(ChangeMajorNB), TrainingSet$On_Time)
```

### Proportion of On / Over-Time graduation by Degree

```{r}

 DT_CM <- ChangeMajorNB[["tables"]][["End_Prog"]]

DT_CM_DF <- as.data.frame.matrix(DT_CM)
DT_CM_DF$EarnedDegree <- rownames(DT_CM_DF)
DT_CM_DF1 <- DT_CM_DF %>%  select(EarnedDegree, "On_Time"=Yes, "Over_Time"=No)

library(DT)
datatable(DT_CM_DF1, extensions = "Buttons",caption = "Proportion of On/Off-Time by Earned Degree: Biology",
          filter = "top",
          options = list(dom="Blfrtip",buttons=c("copy","csv","excel","pdf","print"), lengthMenu=list(c(10,25,50,-1), c(10,25,50,"All")),pageLength=25,scrollY=800,scrollX=TRUE))

```


### Data Table

```{r message=FALSE, warning=FALSE}
p1 <- round( predict(ChangeMajorNB , TrainingSet, type = "prob"), digits = 2)
p2 <- round(predict(ChangeMajorNB, TestingSet, type = "prob"), digits = 2)
results_train <- cbind(p1, TrainingSet)
results_test <- cbind(p2, TestingSet)
Results_all <- rbind(results_train,results_test) 
colnames(Results_all) <- c("Pred_OverTime", "Pred_OnTime","Actual_OnTime","Entered_Prog","EarnedDeg_Prog","FCProg_Class","TotalChangedProg","TotalChangedDepart",
                           "TotalChangedCol","Year1UWFHrs")
library(DT)
datatable(Results_all, extensions = "Buttons",caption = "Probability of On-Time by Changes Majors: Biology",
          filter = "top",
          options = list(dom="Blfrtip",buttons=c("copy","csv","excel","pdf","print"), lengthMenu=list(c(10,25,50,-1), c(10,25,50,"All")),pageLength=25,scrollY=800,scrollX=TRUE))
head(Results_all)
```





