
---
title: "Summary - Admitted 2021 FTIC HS GPA under 3.5"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---


```{r setup, include=FALSE}
library(flexdashboard)
library(dplyr)
library(readr)
HSGPAunder_1 <- read_csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/DATA/3.0-3.5_Tiers.csv", 
    col_types = cols(Column1 = col_skip())) %>% select(UWFID,TIER="Pred_Tier",HS_OFFER_GPA,HS_CORE_GPA,BEST_SCORE,TRANSFER_HOURS_EARNED,HS_NAME,HS_CNTY,codeCOLLEGE="CURRICULUM_COLL",codeDEPART="CURRICULUM_DEPT",
                                                       contains("PROGRAM"),GENDER,RACE,COUNTY,COUNTY_GROUP,AGE_AT_ENTRY,CNTY1,contains("PELL"))
pMISS <- function(x){sum(is.na(x))/length(x)*100}
apply(HSGPAunder_1,2, pMISS)
sapply(HSGPAunder_1, function(x) sum(is.na(x)))
##HSGPA rank
HSGPAunder <- HSGPAunder_1 %>% mutate(codeHSGPA=ifelse(HS_OFFER_GPA<=3.1000,"GPA<=3.10",
                                                     ifelse(HS_OFFER_GPA<=3.20,"GPA<=3.20",
                                                            ifelse(HS_OFFER_GPA<=3.30,"GPA<=3.30",
                                                                   ifelse(HS_OFFER_GPA<=3.40,"GPA<=3.40","GPA<3.50"))))) %>% 
                                              mutate(COLLEGE=ifelse(codeCOLLEGE==0,"UNA", # college
                                                               ifelse(codeCOLLEGE=="A","HMCSE",
                                                                ifelse(codeCOLLEGE=="P","CEPS",
                                                                 ifelse(codeCOLLEGE=="H","CASSH",
                                                                    ifelse(codeCOLLEGE=="M","UKCOH",
                                                                      ifelse(codeCOLLEGE=="B","COB","NA")))))))
  
# transfer hours NA==0
#HSGPAunder[which(is.na(HSGPAunder$TRANSFER_HOURS_EARNED)),"TRANSFER_HOURS_EARNED"] <- 0
# HS_CORE_GPA
#HSGPAunder[which(is.na(HSGPAunder$HS_CORE_GPA)),"HS_CORE_GPA"] <- 0
# RACE NA=NA
HSGPAunder[which(is.na(HSGPAunder$RACE)),"RACE"] <- "Not Reported"

HSGPAunder$codeRACE <- HSGPAunder$RACE
HSGPAunder[HSGPAunder$codeRACE == "W","codeRACE"] <- "White"
HSGPAunder[HSGPAunder$codeRACE == "B","codeRACE"] <- "African American"
HSGPAunder[HSGPAunder$codeRACE == "A","codeRACE"] <- "Asian"
HSGPAunder[HSGPAunder$codeRACE == "I","codeRACE"] <- "American Indian/Alaskan Native"
twoormore <- c("I,W","P,W","B,W","B,P","A,W","A,B")

HSGPAunder$codeRACE1 <- sapply(HSGPAunder[,23], function(x) replace(x, x %in% c("I,W","P,W","B,W","B,P","A,W","A,B"), 99))
 HSGPAunder$codeRACE1 <- gsub(99,"Two or More", HSGPAunder$codeRACE1)
 
HSGPAunder[HSGPAunder$GENDER=="F","GENDER"] <- "Female"
HSGPAunder[HSGPAunder$GENDER=="M","GENDER"] <- "Male"

### recode Ethnicity
HSGPAunder_new <- HSGPAunder %>% mutate(codeEthnicity=ifelse(codeRACE1=="White","White",
                                                         ifelse(codeRACE1=="African American","AA",
                                                                ifelse(codeRACE1=="Asian","Asian",
                                                                       ifelse(codeRACE1=="Two or More","Two or More","Others")))))

HSGPAunder_new$codePROGRAM <- substr(HSGPAunder_new$PROGRAM_CODE, start = 1, stop = 8)
HSGPAunder_new$codePROGRAM1 <- substr(HSGPAunder_new$PROGRAM_CODE, start = 3, stop = 5)

HSGPAunder_new$codeEthnicity1 <- HSGPAunder_new$codeEthnicity
HSGPAunder_new[HSGPAunder_new$codeDEPART==0,"codeDEPART"] <- "UNA"
HSGPAunder_new[HSGPAunder_new$codeCOLLEGE==0,"codeCOLLEGE"] <- "UNA"
HSGPAunder_new$APR <- NA
write.csv(HSGPAunder_new,"HSGPAunder3_5.csv")
```

```{css, echo=FALSE}
.rpivottable{overflow-x: scroll; }
.DataTable{"scrollX": true}
```
Admitted 2021 FTIC High School GPA under 3.5
======================================================================


Column {.tabset .tabset-fade }
-----------------------------------------------------------------------


### TIER by COLLEGE

```{r}
library(plotly)
library(dplyr)

HSGPAunder_new$TIER <- as.factor(HSGPAunder_new$TIER)
fig_a <- HSGPAunder_new %>% count(TIER, COLLEGE)
fig_a <- fig_a %>% plot_ly(x= ~COLLEGE, y=~n, color=~TIER)
fig_a
```

### High School GPA (by tenths) by Counties

```{r}
HSGPAunder_new$codeHSGPA <- as.factor(HSGPAunder_new$codeHSGPA)
fig_b <- HSGPAunder_new %>% 
  group_by(CNTY1,codeHSGPA) %>% 
  summarise(count=n()) %>% na.omit() %>% 
  plot_ly(x=~CNTY1, y=~count,color =~ codeHSGPA, type="bar") %>% 
  #add_trace(y=~count, name= "CRDGPAAPR") %>% 
  layout(xaxis=list(title="Counties"), yaxis=list(title='Counts by Counties'), barmode="stack")
fig_b
```


### High School GPA by Ethnicity


```{r}
HSGPAunder_new$codeEthnicity <- as.factor(HSGPAunder_new$codeEthnicity)
#unique(HSGPAunder_new$codeEthnicity)
fig_b2 <- HSGPAunder_new %>% 
  group_by(codeEthnicity,codeHSGPA) %>% 
  dplyr::summarise(count=n(), .groups='drop') %>%  
  plot_ly(x=~codeHSGPA, y=~count, color=~codeEthnicity, type="bar") %>% 
  layout(xaxis=list(title="High School GPA by tenths"), yaxis=list(title='High Scholl GPA by Ethnicity'), barmode="stack")
fig_b2
```




### Pivot Table by HS GPA (by tenths)

```{r}
library(rpivotTable)
rpivotTable(HSGPAunder_new, aggregatorName = "Count",
            rows = c("COLLEGE","TIER"),
            
            cols = c("codeHSGPA"),
           
            subtotals = T, rendererName = "Table With Subtotal")
```

### DATA Table

```{r }
library(DT)
datatble <- DT::datatable(HSGPAunder_new, extensions = "Buttons",caption = "HS GPA under 3.5",
          filter = "top",
          options = list(dom="Blfrtip",buttons=c("copy","csv","excel","pdf","print"), lengthMenu=list(c(10,25,50,-1), c(10,25,50,"All")),pageLength=25,scrollY=800,scrollX=TRUE))
datatble
```


### Numbers of 2021 FTIC in HMCSE with HS GPA under 3.5 (by tenths) 

```{r}
HSGPAunder_new$TIER <- as.factor(HSGPAunder_new$TIER)
fig_C <- HSGPAunder_new[HSGPAunder_new$COLLEGE=="HMCSE",] %>% 
  group_by(codeHSGPA,PROGRAM_DESC) %>% 
  summarise(count=n()) %>% na.omit() %>% 
  plot_ly(x=~PROGRAM_DESC, y=~count,color =~ codeHSGPA, type="bar") %>% 
  #add_trace(y=~count, name= "CRDGPAAPR") %>% 
  layout(xaxis=list(title="PROGRAMS in HMCSE"), yaxis=list(title='Counts by HS GPA'), barmode="stack")
fig_C
```

### Average HS GPA by College and Program 2021 FTIC

```{r}
library(dplyr)
summaryall <- HSGPAunder_new %>% group_by(COLLEGE,PROGRAM_DESC) %>% dplyr::summarise(meanHSGPA=round(mean(HS_OFFER_GPA, na.rm=T),4), meanHSCoreGPA=round(mean(HS_CORE_GPA, na.rm=T),4), meanTransHrs=mean(TRANSFER_HOURS_EARNED,na.rm=T), Counts=n())

library(DT)
datatable(summaryall, extensions = "Buttons",caption = "Summary  HS GPA under 3.5",
          filter = "top",
          options = list(dom="Blfrtip",buttons=c("copy","csv","excel","pdf","print"), lengthMenu=list(c(10,25,50,-1), c(10,25,50,"All")),pageLength=25,scrollY=800,scrollX=TRUE))
```

Prediction of APR for 2021 HSGPA under3.5 - Logistic Regression Analysis
==============================================================================================



Column {.tabset .tabset-fade }
-----------------------------------------------------------------------


```{r  include=FALSE}

########## APR ###########
##################################
# build a model using CSE data
CSE_ALL_ENR_up2020 <- read.csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/DATA/CSE_ALL_ENR_up2020.csv", stringsAsFactors=TRUE) 
HSGPAunder3.5_DF <- read.csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/DATA/CSE_ALL_ENR_up2020.csv", stringsAsFactors=TRUE) %>% 
  filter(COHORT_YEAR<=20192020,GPA_HIGHSCHOOL<3.500) %>%
  select(GENDER,codeRACE="ETHNICITY",AGE_AT_ENTRY,GPA_HIGHSCHOOL,COUNTY_GROUP,COUNTY,TIER="APPLICANT_TIER",AveGPA2.77,APR, HIGH_SCHOOL_NAME,contains("PROGRAM"),contains("PROPORTION"),contains("DEPART"),contains("code")) %>% 
  mutate(codeHSGPA=ifelse(GPA_HIGHSCHOOL<=3.1000,"GPA<=3.10",
                                                     ifelse(GPA_HIGHSCHOOL<=3.2000,"GPA<=3.20",
                                                            ifelse(GPA_HIGHSCHOOL<=3.30000,"GPA<=3.30",
                                                                   ifelse(GPA_HIGHSCHOOL<=3.40000,"GPA<=3.40",
                                                                          ifelse(GPA_HIGHSCHOOL<3.50000,"GPA<3.50","GPA>=3.50")))))) %>% 
  mutate(codeEthnicity=ifelse(codeRACE=="White","White",
                              ifelse(codeRACE=="African American","AA",
                                     ifelse(codeRACE=="Asian","Asian",
                                            ifelse(codeRACE=="Two or More","Two or More","Others"))))) 
 
codePROGRAM <- HSGPAunder3.5_DF$ENTRY_PROGRAM_CODE 
## code program
codePROGRAM1 <- substr(codePROGRAM, start=3, stop=5)
HSGPAunder3.5_DF$codePROGRAM1 <- codePROGRAM1
HSGPAunder3.5_DF$BestTest_PROPORTION <- pmax(HSGPAunder3.5_DF$ACT_PROPORTION,HSGPAunder3.5_DF$SAT_PROPORTION)
#range(HSGPAunder3.5_DF$BestTest_PROPORTION)
## missing values
HSGPAunder3.5_DF[which(HSGPAunder3.5_DF$BestTest_PROPORTION==0),"BestTest_PROPORTION"] <- NA
sapply(HSGPAunder3.5_DF,pMISS)



### To get an probability of APR by Counties and HS Names
tab_counties <- addmargins(table(HSGPAunder3.5_DF$COUNTY, HSGPAunder3.5_DF$APR))
DF_counites <- as.data.frame.matrix(tab_counties)
DF_counites$COUNTY <- rownames(tab_counties)
DF_counites <- DF_counites  %>% filter(COUNTY != "Sum")
DF_counites$APRRATE_counties <- round( DF_counites$Yes/DF_counites$Sum*100,4)


### To get an probability of APR by Counties and HS Names
tab_HSNAME <- addmargins(table(HSGPAunder3.5_DF$HIGH_SCHOOL_NAME, HSGPAunder3.5_DF$APR))
DF_HSNAME <- as.data.frame.matrix(tab_HSNAME)
DF_HSNAME$HIGH_SCHOOL_NAME <- rownames(tab_HSNAME)
DF_HSNAME <- DF_HSNAME  %>% filter(HIGH_SCHOOL_NAME != "Sum")
DF_HSNAME$APRRATE_HSNAME <- round(DF_HSNAME$Yes/DF_HSNAME$Sum*100,4)

HSGPAunder3.5_dataset_rate <- merge(DF_counites,HSGPAunder3.5_DF, by="COUNTY", all=T)
HSGPAunder3.5_dataset_rate2 <- merge(HSGPAunder3.5_dataset_rate,DF_HSNAME, by="HIGH_SCHOOL_NAME", all=T)
write.csv(HSGPAunder3.5_dataset_rate2,"HSGPAunder3.5_dataset_rate2.csv")
### To get an APR Rate by program
tab_PROGRAMcode <- addmargins(table(HSGPAunder3.5_DF$codePROGRAM1, HSGPAunder3.5_DF$APR))
DF_tab_PROGRAMcode <- as.data.frame.matrix(tab_PROGRAMcode)
DF_tab_PROGRAMcode$codePROGRAM1 <- rownames(tab_PROGRAMcode)
DF_tab_PROGRAMcode <- DF_tab_PROGRAMcode %>% filter(codePROGRAM1 != "Sum")
DF_tab_PROGRAMcode$APRRATE_PROGRAM <- round( DF_tab_PROGRAMcode$Yes/DF_tab_PROGRAMcode$Sum*100,4)
HSGPAunder3.5_dataset_rate3 <- merge(DF_tab_PROGRAMcode,HSGPAunder3.5_dataset_rate2, by.x="codePROGRAM1",by.y="codePROGRAM1", all=T)
write.csv(HSGPAunder3.5_dataset_rate3,"HSGPAunder3.5_dataset_rate3.csv")
### To get an average APR HSGPA for under APR FTIC by codeprogramn
APR_HSGPA_PROGRAM_DF <- read.csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/DATA/CSE_ALL_ENR_up2020.csv", stringsAsFactors=FALSE) %>% 
  filter((COHORT_YEAR<=20192020 & COHORT_YEAR>=20172018)) %>% select(contains("APR"),contains("PROGRAM"),contains("GPA"))
APR_HSGPA_PROGRAM_DF$codePROGRAM1 <- substr(APR_HSGPA_PROGRAM_DF$ENTRY_PROGRAM_CODE, start = 3,stop = 5)
APR_HSGPA_PROGRAM_DF1 <- APR_HSGPA_PROGRAM_DF %>% group_by(codePROGRAM1,APR) %>% dplyr::summarise(meanHSGPAAPR=mean(GPA_HIGHSCHOOL, na.rm=T), Counts=n()) %>% arrange(codePROGRAM1)
# merge
library(tidyr)
spreadDF <- APR_HSGPA_PROGRAM_DF1 %>% spread(APR,meanHSGPAAPR)
colnames(spreadDF) <- c("codePROGRAM1","CountFTIC","AveHSGPA_NoAPR","AveHSGPA_YesAPR")
df1 <- spreadDF %>% select(1,2,3) %>% na.omit()
df2 <- spreadDF %>% select(1,2,4) %>% na.omit()
AveHSGPA_APR <- merge(df1,df2, by="codePROGRAM1", all=T)
colnames(AveHSGPA_APR) <- c("codePROGRAM1","CountNoAPR","AveHSGPA_NoAPR","CountYesAPR","AveHSGPA_YesAPR")

HSGPAunder3.5_dataset_rate4 <- merge(AveHSGPA_APR,HSGPAunder3.5_dataset_rate3, by.x="codePROGRAM1",by.y="codePROGRAM1", all=T)
write.csv(HSGPAunder3.5_dataset_rate4,"HSGPAunder3.5_dataset_rate4.csv")



# two way table
xtabs(~GENDER+APR, data = HSGPAunder3.5_dataset_rate4)
xtabs(~codeEthnicity+APR, data = HSGPAunder3.5_dataset_rate4)
xtabs(~TIER+APR, data = HSGPAunder3.5_dataset_rate4)
xtabs(~codeHSGPA+APR, data = HSGPAunder3.5_dataset_rate4)
xtabs(~codePROGRAM1+APR,data = HSGPAunder3.5_dataset_rate4)

### my data set
HSGPAunder3.5_dataset_rate4[which(is.na(HSGPAunder3.5_dataset_rate4$APRRATE_PROGRAM)),"APRRATE_PROGRAM"] <- 0
HSGPAunder3.5_dataset_rate4[which(is.na(HSGPAunder3.5_dataset_rate4$APRRATE_counties)),"APRRATE_counties"] <- 0
HSGPAunder3.5_dataset_rate4[which(is.na(HSGPAunder3.5_dataset_rate4$APRRATE_HSNAME)),"APRRATE_HSNAME"] <- 0
HSGPAunder3.5_dataset_rate4[which(is.na(HSGPAunder3.5_dataset_rate4$AveHSGPA_NoAPR)),"AveHSGPA_NoAPR"] <- 0
HSGPAunder3.5_dataset_rate4[which(is.na(HSGPAunder3.5_dataset_rate4$AveHSGPA_YesAPR)),"AveHSGPA_YesAPR"] <- 0
HSGPAunder3.5_dataset_rate4[which(is.na(HSGPAunder3.5_dataset_rate4$GPA_HIGHSCHOOL)),"GPA_HIGHSCHOOL"] <- NA


HSGPA_data <- HSGPAunder3.5_dataset_rate4 %>%   filter(TIER != 1) %>%  select(APRRATE_PROGRAM,APRRATE_counties,APRRATE_HSNAME,APRAveHSGPA_PROGRAM="AveHSGPA_YesAPR",GENDER,codeEthnicity,codeHSGPA,AGE_AT_ENTRY,TIER,APR) %>% na.omit()#1889

## replacing missing values
library(mice)
library(VIM)
#HSGPA_data$GPA_HIGHSCHOOL <- as.numeric(gsub("\\.","",HSGPA_data$GPA_HIGHSCHOOL))
#HSGPA_data$TIER <- as.numeric(gsub("\\.","",HSGPA_data$TIER))
#HSGPA_data$BestTest_PROPORTION <- as.numeric(gsub("\\.","",HSGPA_data$BestTest_PROPORTION))
#imp_regression <- regressionImp(BestTest_PROPORTION ~ GPA_HIGHSCHOOL+TIER, HSGPAunder3.5_DF)
#aggr(imp_regression, delimiter = "_imp")
#imputeTest <- mice(HSGPA_data[, c("BestTest_PROPORTION","GPA_HIGHSCHOOL","TIER"),m=3,seed=123])
#imputeHS$imp$GPA_HIGHSCHOOL
```

### Prediction of APR for HS GPA under3.5 2021 FTIC (sample)

```{r}
### logistice modeling
library(caret)
# Performs stratified random split of the data set
TrainingIndex <- createDataPartition(HSGPA_data$APR, p=0.8, list = FALSE)
TrainingSet <- HSGPA_data[TrainingIndex,] # Training Set
TestingSet <- HSGPA_data[-TrainingIndex,] # Test Set

# Build model using all factors and labels
set.seed(111)
HSGPAunderlgm <- glm(APR ~ ., data = TrainingSet, family="binomial")
summary(HSGPAunderlgm)
exp(coef(HSGPAunderlgm))
set.seed(222)
HSGPAunderlgm_re <- glm(APR ~APRRATE_PROGRAM+APRRATE_HSNAME+GENDER+codeHSGPA+codeHSGPA*codeEthnicity, data = TrainingSet, family="binomial")
summary(HSGPAunderlgm_re)
library(tidyverse)  # data manipulation and visualization
library(modelr)     # provides easy pipeline modeling functions
library(broom)      # helps to tidy up model outputs

P_value <- tidy(HSGPAunderlgm_re)
#coefficient
coef1 <- exp(coef(HSGPAunderlgm_re))
coef_table <- knitr::kable(coef1)
coef_table
#confint(HSGPAunderlgm_re)
vari_imp <- knitr::kable(caret::varImp(HSGPAunderlgm_re))
vari_imp

#accuracy 
p_1 <- predict(HSGPAunderlgm_re, TrainingSet, type="response")
#head(p_1,10)
#head(TrainingSet,10)
pred_1 <- ifelse(p_1>0.5, 1,0)
tab_1 <- table(Predicted=pred_1, Actural=TrainingSet$APR)
tab_1
round(sum(diag(tab_1))/sum(tab_1),4) # 0.2876387

#test data
p_2 <- predict(HSGPAunderlgm_re, TestingSet, type="response")
#head(p_2,10)
#head(TestingSet,10)
pred_2 <- ifelse(p_2>0.5, 1,0)
tab_2 <- table(Predicted=pred_2, Actural=TestingSet$APR)
tab_2
round(sum(diag(tab_2))/sum(tab_2),4) #0.0.2761905

### goodness of fit
with(HSGPAunderlgm_re, pchisq(null.deviance - deviance, df.null-df.residual, lower.tail=F))#p-value==  1.127255e-80


```



```{r eval=FALSE, include=FALSE}
# building new data set
library(dplyr)
# for admitted 2021 FTIC

#test_2021FTICadmission <- HSGPAunder %>% 
#  select(TIER,HIGH_SCHOOL_NAME="HS_NAME")
HS_NAME_APR <- HSGPAunder3.5_dataset_rate4 %>% select(HIGH_SCHOOL_NAME, APRRATE_HSNAME,CountTotalHS_NAME="Sum.y" )
HS_NAME_APR <- HS_NAME_APR[!duplicated(HS_NAME_APR),]
test_HSNAMEAPR <- merge(HS_NAME_APR, HSGPAunder_new, by.x="HIGH_SCHOOL_NAME",by.y="HS_NAME", all.y=T) 
test_HSNAMEAPR$codeEthnicity <- as.factor(test_HSNAMEAPR$codeEthnicity)

PROGRAM_APR <- HSGPAunder3.5_dataset_rate4 %>% select(codePROGRAM1, APRRATE_PROGRAM,CounTotal_HSGPAunder="Sum" ,AveHSGPA_NoAPR,CountNoAPR,AveHSGPA_YesAPR,CountYesAPR)
PROGRAM_APR <- PROGRAM_APR[!duplicated(PROGRAM_APR),]
test_APRRATE_PROGRAM <- merge(PROGRAM_APR, test_HSNAMEAPR, by.x=c("codePROGRAM1"),by.y=c("codePROGRAM1"), all.y=T)
test_APRRATE_PROGRAM <- test_APRRATE_PROGRAM[!duplicated(test_APRRATE_PROGRAM$UWFID),]
test_APRRATE_PROGRAM$codePROGRAM <- as.factor(test_APRRATE_PROGRAM$codePROGRAM1)
# for APRAveHSGPA, need to average GPA by program
HSGPA_by_program <- HSGPAunder_new %>% group_by(codePROGRAM1) %>% dplyr::summarise(APRAveHSGPA_PROGRAM=mean(HS_OFFER_GPA,na.rm=T))
test_APRRATE_PROGRAM1 <- merge(test_APRRATE_PROGRAM,HSGPA_by_program,by="codePROGRAM1",all.x = T)
test_APRRATE_PROGRAM1[which(is.na(test_APRRATE_PROGRAM1$APRRATE_HSNAME)),"APRRATE_HSNAME"] <- 0
test_APRRATE_PROGRAM1[which(is.na(test_APRRATE_PROGRAM1$APRRATE_PROGRAM)),"APRRATE_PROGRAM"] <- 0

# APRRATE_BYHS
HS_meanPGA <- HSGPA_data %>% group_by(codeHSGPA) %>% dplyr::summarise(meanAPRRATE_HS=mean(APRRATE_HSNAME, na.rm=T), count=n())
HS_meanPGA
APRRAT_HS <- HSGPAunder3.5_dataset_rate4 %>% group_by(codeCOLLEGE) %>% dplyr::summarise(meanAPR=mean(APRRATE_PROGRAM,na.rm=T),Count=n())
APRRAT_HS
#dataset
APRRAT_test_APRRATE_PROGRAM1 <- test_APRRATE_PROGRAM1 %>% group_by(codeHSGPA,codePROGRAM1) %>% dplyr::summarise(meanAPR=mean(APRRATE_PROGRAM,na.rm=T),Count=n())
APRRAT_test_APRRATE_PROGRAM1

test_APRRATE_PROGRAM1[(test_APRRATE_PROGRAM1$codeHSGPA=="GPA<=3.10" & test_APRRATE_PROGRAM1$APRRATE_HSNAME==0),"APRRATE_HSNAME"] <- 61.80
test_APRRATE_PROGRAM1[(test_APRRATE_PROGRAM1$codeHSGPA=="GPA<=3.20" & test_APRRATE_PROGRAM1$APRRATE_HSNAME==0),"APRRATE_HSNAME"] <- 62.90
test_APRRATE_PROGRAM1[(test_APRRATE_PROGRAM1$codeHSGPA=="GPA<=3.30" & test_APRRATE_PROGRAM1$APRRATE_HSNAME==0),"APRRATE_HSNAME"] <- 63.20
test_APRRATE_PROGRAM1[(test_APRRATE_PROGRAM1$codeHSGPA=="GPA<=3.40" & test_APRRATE_PROGRAM1$APRRATE_HSNAME==0),"APRRATE_HSNAME"] <- 62.90
test_APRRATE_PROGRAM1[(test_APRRATE_PROGRAM1$codeHSGPA=="GPA<3.50"  & test_APRRATE_PROGRAM1$APRRATE_HSNAME==0),"APRRATE_HSNAME"] <- 64.80
test_APRRATE_PROGRAM1[(test_APRRATE_PROGRAM1$codePROGRAM1=="ISS"  & test_APRRATE_PROGRAM1$APRRATE_PROGRAM==0),"APRRATE_PROGRAM"] <- 62.80 #codeHSGPA3.29
```

```{r include=FALSE}
####################################################################
###### replacing APRRATE_HS using codeHSGPA and APRRATE_PROGRAM ###
###################################################################
# prediction data
glimpse(HSGPAunder_new) #codePROGRAM1,codeHSGPA,HS_NAME,HS_NAME
glimpse(HSGPAunder3.5_dataset_rate4) #APRRATE_HSNAME,codeHSGPA,codePROGRAM1,APRRATE_PROGRAM,HIGH_SCHOOL_NAME
program1 <- HSGPAunder3.5_dataset_rate4 %>% select(codePROGRAM1, APRRATE_PROGRAM)
program2 <-program1[!duplicated(program1),]
program3 <- program2[!is.na(program2$codePROGRAM1),]
# for replacing
programa <- HSGPAunder3.5_dataset_rate4 %>% select(codeHSGPA,codePROGRAM1, APRRATE_PROGRAM)
programb <-programa[!duplicated(programa),]
programc <- programb[!is.na(programb$codePROGRAM1),]
colnames(programc) <- c("recodeHSGPA","recodePROGRAM","recodeAPRRATE_PROGRAM")

# HSPAR
hsname <- HSGPAunder3.5_dataset_rate4 %>% select(HIGH_SCHOOL_NAME,APRRATE_HSNAME)
hsname1 <- hsname[!duplicated(hsname),]
hsname2 <- hsname1[!is.na(hsname1$HIGH_SCHOOL_NAME),]
merge1 <- merge(HSGPAunder_new, program3, by.x=c("codePROGRAM1"), by.y="codePROGRAM1", all.x=T)
merge1[!duplicated(merge1),]

merge2 <- merge(merge1, hsname2, by.x=c("HS_NAME"), by.y=c("HIGH_SCHOOL_NAME"), all.x=T)
merge3 <- merge(merge2,programc, by.x=c("codeHSGPA","codePROGRAM1"),by.y=c("recodeHSGPA","recodePROGRAM"), all.x=T)

merge3[which(is.na(merge3$APRRATE_HSNAME)), "APRRATE_HSNAME"] <- 0
merge3[which(is.na(merge3$recodeAPRRATE_PROGRAM)), "recodeAPRRATE_PROGRAM"] <- 0

merge3$replacedAPRRATE_HSNAME <- ifelse(merge3$APRRATE_HSNAME==0, merge3$recodeAPRRATE_PROGRAM, merge3$APRRATE_HSNAME)
merge3[which(is.na(merge3$replacedAPRRATE_HSNAME)),"replacedAPRRATE_HSNAME"] <- 0
merge3$replacedAPRRATE_HSNAME <- ifelse(merge3$replacedAPRRATE_HSNAME==0, merge3$APRRATE_PROGRAM, merge3$replacedAPRRATE_HSNAME)
merge3[merge3$codePROGRAM1=="ISS","replacedAPRRATE_HSNAME"] <- 63.52577# average program APR for CEPS
merge3[merge3$codePROGRAM1=="ISS","APRRATE_PROGRAM"] <- 63.52577# average program APR for CEPS

#merge3[which(is.na(merge3$APRRATE_PROGRAM)),]
merge5 <- merge3 %>% select(1:30,APRRATE_HSNAME="replacedAPRRATE_HSNAME")
sapply(merge5, pMISS)

test_APRRATE_PROGRAM1 <- merge5

```

### Predicting a Probability of APR among 2021 FTIC - table(sample)

```{r}
# prediction
#HSGPAunderlgm_re$xlevels[["codeHSGPA"]] <- union(HSGPAunderlgm_re$xlevels[["codeHSGPA"]], levels(test_HSNAMEAPR$codeHSGPA))
#HSGPAunderlgm_re$xlevels[["codeEthnicity"]] <- union(HSGPAunderlgm_re$xlevels[["codeEthnicity"]], levels(test_HSNAMEAPR$codeEthnicity))
# prediction
test_pre <- round(predict(HSGPAunderlgm_re, test_APRRATE_PROGRAM1, type="response"),4)
#head(test_pre,10)
#head(test_HSNAMEAPR,10)
Results_test2021admitted <- cbind(test_pre, test_APRRATE_PROGRAM1)

results_table_all <- Results_test2021admitted %>%
  mutate(codeProb_APR1=ifelse(test_pre<0.5,"high-risk",
                             ifelse(test_pre<0.75,"moderate-risk","low-risk"))) %>% 
   mutate(codeProb_APR=ifelse(test_pre<0.5,"Pr<0.50",
                             ifelse(test_pre<0.75,"0.50<=Pr<0.75","Pr>=0.75"))) %>% 
  select(UWFID,34,33,Prob_APR="test_pre",Admitted_PROGRAM="codePROGRAM1",PROGRAM_CODE,PROGRAM_DESC,APR_PROGRAM="APRRATE_PROGRAM",
         HS_NAME,APR_HS_NAME="APRRATE_HSNAME",TIER,HS_OFFER_GPA,HS_CORE_GPA,COLLEGE,GENDER,
         COUNTY,COUNTY_GROUP,CNTY1,AGE_AT_ENTRY,contains("PELL"),codeHSGPA,codeEthnicity="codeEthnicity1",APR)

#%>% 
#  select(UWFID,Prob_APR="test_pre",Sum_underHSGPA3.5="Sum.y",TIER,HIGH_SCHOOL_NAME,APRRATE_HSNAME,HS_OFFER_GPA,codeHSGPA,COLLEGE,CURRICULUM_DEPT,PROGRAM_DESC,GENDER,codeEthnicity,APR)
write.csv(results_table_all,"results_table_all_02092021.csv")
library(DT)
results_table <- datatable(results_table_all, extensions = "Buttons",caption = "Probability of APR",
          filter = "top",
          options = list(dom="Blfrtip",buttons=c("copy","csv","excel","pdf","print"), lengthMenu=list(c(10,25,50,-1), c(10,25,50,"All")),pageLength=25,scrollY=800,scrollX=TRUE))

results_table
```

### APR Rates among GPA under3.5 by High School Name

```{r}
library(highcharter)
library(dplyr)
library(readr)
#data from 2015to19
HSGPAunder3_5rate2 <- read_csv("HSGPAunder3.5_dataset_rate4.csv") %>% select(HIGH_SCHOOL_NAME,APR_Rate_by_HS_Name="APRRATE_HSNAME",Sum_HS_Name="Sum.y")
HSGPAunder3_5rate21 <- HSGPAunder3_5rate2[!duplicated(HSGPAunder3_5rate2),]
#results
Results_table_1 <- read.csv("G:/My Drive/#1_Jay's Space/AdmittedFTIC2021/Results_table_1.csv") %>% select(UWFID,Prob_APR)
HSGPAunder_2 <- read.csv("G:/My Drive/#1_Jay's Space/AdmittedFTIC2021/V1HSGPAunder3_5.csv")
#filter for 2021
admittedHSNAME <- merge(HSGPAunder3_5rate21, HSGPAunder_2, by.x="HIGH_SCHOOL_NAME", by.y="HS_NAME",all.y=T)
admittedHSNAME1 <- merge(admittedHSNAME,Results_table_1,by="UWFID", all.x=T)
myhchart <- admittedHSNAME1 %>% select(Sum_HS_Name,APR_Rate_by_HS_Name_percentage="APR_Rate_by_HS_Name",HIGH_SCHOOL_NAME)
fig_b3 <- hchart(myhchart, "scatter", hcaes(x = Sum_HS_Name, y = APR_Rate_by_HS_Name_percentage , group = HIGH_SCHOOL_NAME ))
fig_b3
```

### Pivot Table-Prediction

```{r}
results_table_all$codeProb_APR <- as.factor(results_table_all$codeProb_APR)
results_table_all <- results_table_all %>% arrange(-Prob_APR)
library(rpivotTable)
rpivotTable(results_table_all, aggregatorName = "Count",
            rows = c("COLLEGE"),
            
            cols = c("codeProb_APR"),
           
            subtotals = T, rendererName = "Table With Subtotal")

```


```{r  include=FALSE}
library(readr)
hist_HSNAME <- read_csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/DATA/CSE_ALL_ENR_up2020.csv") %>% 
  select(COHORT_YEAR,APR,HIGH_SCHOOL_NAME,ENTRY_PROGRAM,GPA_HIGHSCHOOL) %>% filter(COHORT_YEAR<=2019202, COHORT_YEAR>=20172018) %>% na.omit() %>% 
  mutate(codeHSGPA=ifelse(GPA_HIGHSCHOOL<3.500,"<3.50",">=3.50"))
sapply(hist_HSNAME, function(x) sum(is.na(x)))

hist_HSNAME_under3.5 <- hist_HSNAME %>% filter(codeHSGPA=="<3.50")

new_HSNAME <- HSGPAunder_new %>% select(HS_NAME)
```
### DATA Summary

* Total number of FTIC from 2015 to 2020 is 6,851.
* The total with HS GPA under 3.5 is 1,955.
* For this analysis, 1,900 were used after removing Tier 1 and NA FTIC.
* The train and test data sets are 80% and 20% (1,521 and 379).
* Only these FTIC are used when calculating APR by Program and HS name.
* The average program APR by HS GPA was substituted for HS names with no history.
* For example, if a student was admitted to MAT with no HS name history, then their APR by HS name would be replaced with the average of the program APR for the same GPA group.
* Program Code: Updated Cyber-security and Computer Science codes as "CYB" and "CSC" in the Admitted FTIC program codes (old code was "CPS").
* The number of unique high school names is `r  length(unique(CSE_ALL_ENR_up2020$HIGH_SCHOOL_NAME))`.
* The number used for the prediction is `r  length(unique(HSGPAunder3.5_DF$HIGH_SCHOOL_NAME))`.
* The number of unique high school names for HS GPA under 3.5 for 2021 FTIC is `r  length(unique(HSGPAunder_new$HS_NAME))`.
* 173 out of 272 high schools did not have recorded APRs for HS GPA under 3.5.


### NEW High School Name & Average HS GPA (Overall 2021 FTIC)

```{r}
library(readr)
HSNAME_recorded <- read.csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/DATA/CSE_ALL_ENR_up2020.csv", stringsAsFactors=FALSE) %>% select(HIGH_SCHOOL_NAME,GPA_HIGHSCHOOL) %>% group_by(HIGH_SCHOOL_NAME) %>% dplyr::summarise(meanHSGPA_recorded=mean(GPA_HIGHSCHOOL,na.rm=T), Counts_recorded=n())
colnames(HSNAME_recorded) <- c("HIGH_SCHOOL_NAME","HSGPA_Recorded","Counts_recorded")

HSNAME_2021 <- read.csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/DATA/over3_0_Tiers.csv", stringsAsFactors=FALSE) %>% select(HS_NAME,HS_OFFER_GPA ) %>% 
  group_by(HS_NAME) %>% dplyr::summarise(meanHSGPA_2021=mean(HS_OFFER_GPA,na.rm=T), Counts_2021=n())
colnames(HSNAME_2021) <- c("HIGH_SCHOOL_NAME_2021","HSGPA2021","Counts_2021")

All_HSNAME <- merge(HSNAME_recorded,HSNAME_2021,by.x="HIGH_SCHOOL_NAME",by.y="HIGH_SCHOOL_NAME_2021",all=T)
All_HSNAME[which(is.na(All_HSNAME$Counts_recorded)),"Counts_recorded"] <- 0
All_HSNAME[which(is.na(All_HSNAME$Counts_2021)),"Counts_2021"] <- 0
All_HSNAME1 <- All_HSNAME %>% mutate(NEW_HSNAME = ifelse((Counts_recorded==0 & Counts_2021>0) , "Yes","No"))

#under3.5
HSNAME_2021_under3.5 <- read.csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/DATA/over3_0_Tiers.csv", stringsAsFactors=FALSE) %>% select(HS_NAME,HS_OFFER_GPA ) %>% 
  filter(HS_OFFER_GPA<3.500) %>%  group_by(HS_NAME) %>% dplyr::summarise(meanHSGPAunder3.5_2021=mean(HS_OFFER_GPA,na.rm=T), Counts_2021under3.5=n())
colnames(HSNAME_2021_under3.5) <- c("HIGH_SCHOOL_NAME_2021under3.5","HSGPA2021uner3.5","Counts_2021under3.5")

All_HSNAME2 <- merge(All_HSNAME1,HSNAME_2021_under3.5,by.x="HIGH_SCHOOL_NAME",by.y="HIGH_SCHOOL_NAME_2021under3.5",all=T)
All_HSNAME2[which(is.na(All_HSNAME2$Counts_2021under3.5)),"Counts_2021under3.5"] <- 0
All_HSNAME3 <- All_HSNAME2 %>% mutate(Ratio_HSGPA3.5= round(Counts_2021under3.5/Counts_2021,4))
#write.csv(All_HSNAME3,"All_HSNAMEs_aveHSGPA_2021.csv")
library(DT)
HSNAME_table <- datatable(All_HSNAME3, extensions = "Buttons",caption = "Average HS GPA by HS Name",
          filter = "top",
          options = list(scrollX=TRUE,dom="Blfrtip",buttons=c("copy","csv","excel","pdf","print"), lengthMenu=list(c(10,25,50,-1), c(10,25,50,"All")),pageLength=25,scrollY=800,scrollX=TRUE))

HSNAME_table
```

### High School Achievements...

```{r eval=FALSE, include=FALSE}
#all high shcool name from 2015 to 2020
HSNAME_recorded_all <- read.csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/DATA/CSE_ALL_ENR_up2020.csv", stringsAsFactors=FALSE)
HSNAME_recorded_all$HIGH_SCHOOL_NAME <- toupper(HSNAME_recorded_all$HIGH_SCHOOL_NAME)

HS_Data_achievement <- read.csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/DATA/HS_Data_sum.csv") %>% arrange()
HS_Data_achievement$HS_State <- rep("FL")
HS_Data_achievement$HIGH_SCHOOL_NAME <- toupper(HS_Data_achievement$School_Name)
All_HSNAME_achievement <- merge(HSNAME_recorded_all,HS_Data_achievement, by.x="HIGH_SCHOOL_NAME", by.y="HIGH_SCHOOL_NAME", all.x =T )
# for FTIC 2021
HSNAME_2021$HIGH_SCHOOL_NAME <- toupper(HSNAME_2021$HIGH_SCHOOL_NAME_2021)
All_HSNAME_achievement2021 <- merge(HSNAME_2021,HS_Data_achievement, by.x="HIGH_SCHOOL_NAME", by.y="School_Name", all.x =T )

All_HSNAME_achievement <- merge(All_HSNAME3,HS_Data_achievement, by.x="HIGH_SCHOOL_NAME", by.y="School_Name", all.x =T ) #77.835% missing values all HS
apply(All_HSNAME_achievement, 2, pMISS)

results_table_all$School_Name <- toupper(results_table_all$HIGH_SCHOOL_NAME)
All_results_HSNAME_Achievement <- merge(results_table_all, HS_Data_achievement, by.x="School_Name",by.y="School_Name", all.x=T)# 69.069% missing values under3.5 HS
hist(All_HSNAME_achievement$Sum_Achievement)
```


