---
title: "2020FTIC_1st Fall completed"
author: "JS"
date: "12/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/Data_IR_2020")
```

## Prediction 1st Fall below / aAbove GPA 2.77 for FTIC 2020

This is an prediction accuracy for FTIC 2020 1st Fall GPA document.

This document is only for the  **FTIC 2020 1st Fall** and included APR and Average GPA Rate and Accuracy by Tiers

```{r}
library(readxl)
CSE_enr2020<- read_excel("C:/Users/jsparks3/Downloads/CSE Analysis Full Data 2020 Cohort 12-19-20.xlsx", 
    sheet = "CSE Analysis Data 2020 Cohort")

CSE_crs1stTerm2020 <- read_excel("C:/Users/jsparks3/Downloads/CSE Analysis Full Data 2020 Cohort 12-19-20.xlsx", 
     sheet = "CSE Analysis Courses 2020 Cohor")
CSE_crs2ndTerm2020 <- read_excel("C:/Users/jsparks3/Downloads/CSE Analysis Full Data 2020 Cohort 12-19-20.xlsx", 
    sheet = "Registered Spring courses")

```

### FTIC 2020 Size
```{r}
length(CSE_enr2020$UNIV_ROW_ID)
xtabs(~CSE_enr2020$COHORT_YEAR + CSE_enr2020$ENTRY_COLLEGE)
```
### NAs and mean GPA Histogram
```{r}
sum(is.na(CSE_enr2020$GPA)) #31
CSE_enr2020[is.na(CSE_enr2020$GPA),] # all had Ws
CSE_enr2020[which(is.na(CSE_enr2020$GPA)),"GPA"] <- 0 # all withdrawn for GPA == 0 1029/31
CSE_crs1stTerm2020[CSE_crs1stTerm2020$UNIV_ROW_ID==970576675,]
hist(CSE_enr2020$GPA)
mean(CSE_enr2020$GPA, na.rm = T)


```


### Proportion of GPA2.77 and APR 2.00
```{r}
library(dplyr)
CSE_enr20201 <- CSE_enr2020 %>% mutate(Act_GPAAPR1stTerm=ifelse(GPA >=2.77,"Above2.77","Belowe2.77")) %>% 
  mutate(APR1stTerm=ifelse(GPA >=2.00,"Above2.00","Belowe2.00"))
xtabs(~CSE_enr20201$Act_GPAAPR1stTerm)
764/1029 #0.74
xtabs(~CSE_enr20201$APR1stTerm)
886/1029 #0.86
```

### FTIC size for each dataset
```{r}
library(readr)
CSEdata2020 <- read_csv("Data_IR/CSEdata2020.csv")
CRSCSEData <- read_csv("Data_IR/CRSCSEData.csv")
updated1CSEdata <- read_csv("Data_IR/updated1CSEdata.csv")
# CSE data
length(CSEdata2020$UNIV_ROW_ID)
library(readxl)
Prediction_NBC_By_Tier_FTIC2020_UWF <- read_excel("Data_IR/output/prediction-results/Prediction_NBC_By_Tier_FTIC2020_UWF.xlsx")
# for prediction
length(Prediction_NBC_By_Tier_FTIC2020_UWF$UNIV_ROW_ID)
xtabs(~Prediction_NBC_By_Tier_FTIC2020_UWF$APPLICANT_TIER)
```

Prediction and Actural data table
```{r}
library(DT)
Pred_matched1 <- CSE_enr20201 %>% select(UNIV_ROW_ID,GPA,Act_GPAAPR1stTerm)
Pred_matched2 <- Prediction_NBC_By_Tier_FTIC2020_UWF %>% mutate(Pred_GPAAPR1stTerm=ifelse(`Pr(AboveGPA)`>=0.5,"Above2.77","Belowe2.77")) %>%  select(UNIV_ROW_ID,Pred_GPAAPR1stTerm,APPLICANT_TIER,ENTRY_COLLEGE)

Matched_2020 <- merge(Pred_matched1,Pred_matched2, by="UNIV_ROW_ID",all.y = T) # merge for 2019
Matched_2020[which(is.na(Matched_2020$Act_GPAAPR1stTerm)),] # 34 NA from original data

#Matched_20201 <- Matched_2020[!is.na(Matched_2020$Pred_GPAAPR1stTerm),] 
Matched_20202 <- Matched_2020 %>% 
      mutate(Matched=ifelse((Act_GPAAPR1stTerm=="Above2.77" & Pred_GPAAPR1stTerm=="Above2.77")|(Act_GPAAPR1stTerm=="Belowe2.77" & Pred_GPAAPR1stTerm=="Belowe2.77"), "Yes","No")) 
datatable(Matched_20202)
```
### Average GPA Rate
```{r}
xtabs(~Matched_20202$Act_GPAAPR1stTerm + Matched_20202$Pred_GPAAPR1stTerm)
xtabs(~Matched_20202$Pred_GPAAPR1stTerm)
762/1026 #actual prop for above 2.77
667/1026 #Prediction
```
### Average GPA 2019

```{r}
CSEGPA2019 <- updated1CSEdata %>% filter(COHORT_YEAR==20192020 & TERM_TO_USE==201908) %>% select(ENTRY_COLLEGE,APPLICANT_TIER,FIRST_FALL_GPA)
CSEGPA2019[which(is.na(CSEGPA2019$FIRST_FALL_GPA)),"FIRST_FALL_GPA"] <- 0
mean(CSEGPA2019$FIRST_FALL_GPA)
meanGPAbyCol2019 <- CSEGPA2019 %>% group_by(APPLICANT_TIER) %>% summarise(meanGPA=mean(FIRST_FALL_GPA), Count=n())
meanGPAbyCol2019
```
### Average GPA 2020
```{r}
Matched_20202[which(is.na(Matched_20202$GPA)),"GPA"] <- 0
mean(Matched_20202$GPA)
meanGPAbyCol2020 <- Matched_20202 %>% group_by(APPLICANT_TIER) %>% summarise(meanGPA=mean(GPA), Count=n())
meanGPAbyCol2020
```




### Actural/Prediction GPA by tier
```{r}
xtabs(~Matched_20202$APPLICANT_TIER+Matched_20202$Act_GPAAPR1stTerm)
```
```{r}
xtabs(~Matched_20202$APPLICANT_TIER+Matched_20202$Pred_GPAAPR1stTerm)
```


### Overall Accuracy
```{r}
xtabs(~Matched_20202$Pred_GPAAPR1stTerm+Matched_20202$Act_GPAAPR1stTerm)
# Accuracy
641/1026
```


### Acurracy by Tier
```{r}
xtabs(~Matched_20202$APPLICANT_TIER)
xtabs(~Matched_20202$APPLICANT_TIER+Matched_20202$Matched)
# Tier 1:5
124/149
252/378
199/374
59/144
7/15

```
### Accuracy by college

```{r}
tabcol <- table(Matched_20202$Matched,Matched_20202$ENTRY_COLLEGE);tabcol
tabcol["Yes",]/colSums(tabcol)

```


### Missed 34 FTIC from original dataset
```{r}
datatable(Matched_20202[is.na(Matched_20202$Act_GPAAPR1stTerm),])
```

### CSE CRS 2019 vs. 2020

```{r}
CSE2019 <- CRSCSEData %>% filter(COHORT_YEAR==20192020 & DEMO_TIME_FRAME==201908
                                 ) %>% select(-2)
colnames(CSE2019) <-    paste("A", colnames(CSE2019), sep="_")
write.csv(CSE2019,"CSE2019.csv")
CSECRS2019 <- CSE2019 %>% select(3,6)

CSE2020 <- CSE_crs1stTerm2020 %>% filter(DEMO_TIME_FRAME==202008) 
colnames(CSE2020) <-paste("B", colnames(CSE2020), sep="_")
write.csv(CSE2020,"CSE2020.csv")

CSECRS2020 <- CSE2020 %>% select(3,6)

CSECRS20192020 <- merge(CSECRS2019,CSECRS2020, by.x = "A_COURSE_NAME", by.y="B_COURSE_NAME", all=T)
CSECRS20192020[CSECRS20192020$A_COURSE_NAME=="23-Mar", "A_COURSE_NAME"] <- "MAR3023"

colnames(CSECRS20192020) <- c("CRSName","Grade2019","Grade2020")
write.csv(CSECRS20192020,"CSECRSGrade20192020.csv")

cbind.fill <- function(...){
  nm <- list(...) 
  nm <- lapply(nm, as.matrix)
  n <- max(sapply(nm, nrow)) 
  do.call(cbind, lapply(nm, function (x) 
    rbind(x, matrix(, n-nrow(x), ncol(x))))) 
}
All_CSECRS <- cbind.fill(CSE2019, CSE2020)
as.data.frame(All_CSECRS)

```
### Numbers of CRS
```{r}
my_DWF_grades <- c("D","D+","W","F","NF","I","I*") # No grades S and XX
CSE2019$DWF <- sapply(CSE2019[,6], function(x) replace(x, x %in% c("D","D+","W","F","NF","I","I*"), "failed"))
CSE2019 <- CSE2019 %>% mutate(DWF1 = ifelse(DWF=="failed","DWF","Pass")) 
AA <- CSE2019 %>% group_by(A_COURSE_NAME,DWF1) %>%   summarise(Count=n(), .groups="drop")

```
```{r}

tabcol2019 <- table(CSE2019$DWF1,CSE2019$A_COURSE_NAME);tabcol2019
tabcol2019["1",]/colSums(tabcol2019)
```


```{r}

CSE2020$DWF <- sapply(CSE2020[,6], function(x) replace(x, x %in% c("D","D+","W","F","NF","I","I*"), "failed"))
CSE2020 <- CSE2020 %>% mutate(DWF1 = ifelse(DWF=="failed","DWF","Pass")) 
CSE2020 %>% group_by(B_COURSE_NAME) %>%  summarise(Count=n(), .groups='drop') %>% arrange(-Count)
```
### prop.table
```{r}
tblFun <- function(x,){
    tbl <- table(x)
    res <- cbind(tbl,round(prop.table(tbl)*100,1))
    colnames(res) <- c('Count','Percentage')
    res
}

do.call(cbind,lapply(CSE2020$B_COURSE_NAME,tblFun))
```

### prop.table1

```{r}
tab2019 <- table(CSE2019$A_COURSE_NAME,CSE2019$DWF1)
tab2019.add <- addmargins(tab2019) %>% as.data.table()
colnames(tab2019.add) <- c("CRSName2019","CRSGrade2019","Count2019")
tab2019.prop <- round(prop.table(tab2019, margin = 1),2) %>% as.data.frame()
colnames(tab2019.prop) <- c("CRSName2019","CRSGrade2019","CRSProp2019")

tab2019DTprop <- merge(tab2019.prop,tab2019.add, by=c("CRSName2019","CRSGrade2019"))


tab2020 <- table(CSE2020$B_COURSE_NAME,CSE2020$DWF1)
tab2020.add <- addmargins(tab2020) %>% as.data.table()
colnames(tab2020.add) <- c("CRSName2020","CRSGrade2020","Count2020")
tab2020.prop <- round(prop.table(tab2020, margin = 1),2) %>% as.data.frame()
colnames(tab2020.prop) <- c("CRSName2020","CRSGrade2020","CRSProp2020")

tab2020DTprop <- merge(tab2020.prop,tab2020.add, by=c("CRSName2020","CRSGrade2020"))


tab20192020 <- merge(tab2019DTprop,tab2020DTprop, by.x = c("CRSName2019","CRSGrade2019"),by.y =c("CRSName2020","CRSGrade2020"), all = T)

write.csv(tab20192020,"tab20192020_CRS_prop.csv")
```

### prop.table2
```{r}
MeanGPA20 <- Matched_20202 %>% group_by(ENTRY_COLLEGE) %>% summarise(meanGPA= round(mean(GPA),2), Count=n(), .groups='drop') %>% as.data.frame();MeanGPA20
colnames(MeanGPA20) <- c("Entry_College","meanGPA2020","count2020")

updated1CSEdata <- read_csv("Data_IR/updated1CSEdata.csv")

updated1CSEdata[which(is.na(updated1CSEdata$FIRST_FALL_GPA)),"FIRST_FALL_GPA"] <- 0
MeanGPA19 <- updated1CSEdata %>% filter(COHORT_YEAR==20192020) %>% group_by(ENTRY_COLLEGE) %>%  summarise(meanGPA= round(mean(FIRST_FALL_GPA),2), Count=n(), .groups='drop') %>% as.data.frame();MeanGPA19
colnames(MeanGPA19) <- c("Entry_College","meanGPA2019","count2019")

meanGPA20192020 <- merge(MeanGPA20,MeanGPA19,by=("Entry_College"),all = T)
write.csv(meanGPA20192020, "meanGPA20192020.csv")
```

```{r}
TierGPA20 <- Matched_20202 %>% group_by(APPLICANT_TIER) %>% summarise(TierGPA= round(mean(GPA),2), Count=n(), .groups='drop') %>% as.data.frame();TierGPA20
colnames(TierGPA20) <- c("Tier","TierGPA2020","count2020")

updated1CSEdata <- read_csv("Data_IR/updated1CSEdata.csv")

updated1CSEdata[which(is.na(updated1CSEdata$FIRST_FALL_GPA)),"FIRST_FALL_GPA"] <- 0
TierGPA19 <- updated1CSEdata %>% filter(COHORT_YEAR==20192020) %>% group_by(APPLICANT_TIER) %>%  summarise(TierGPA= round(mean(FIRST_FALL_GPA),2), Count=n(), .groups='drop') %>% as.data.frame();TierGPA19
colnames(TierGPA19) <- c("Tier","TierGPA2019","count2019")

TierGPA20192020 <- merge(TierGPA20,TierGPA19,by=("Tier"),all = T)
write.csv(TierGPA20192020, "TierGPA20192020.csv")
```



```{r}

MeanGPA20 <- Matched_20202 %>% group_by(ENTRY_COLLEGE) %>% summarise(meanGPA=mean(GPA), Count=n(), .groups='drop') %>% as.data.frame();MeanGPA20
MeanGPA20a <- Matched_20202 %>% group_by(ENTRY_COLLEGE,APPLICANT_TIER) %>% summarise(meanGPA=mean(GPA), Count=n(), .groups='drop') %>% as.data.frame();MeanGPA20a

mean(CSE_enr2020$GPA, na.rm = T)
```

### GPA 2015-2020 1st Fall
```{r}

cse2019GPA <- updated1CSEdata %>% select(COHORT_YEAR, FIRST_FALL_GPA)
colnames(cse2019GPA) <- c("Cohort","GPA1st")
cse2020GPA <- CSE_enr2020 %>% select(COHORT_YEAR,GPA)
colnames(cse2020GPA) <- c("Cohort","GPA1st")

cseall <- rbind(cse2019GPA, cse2020GPA)
cseall[which(is.na(cseall$GPA1st)), "GPA1st"] <- 0

cseallGPA <- cseall1 %>% group_by(Cohort) %>% summarise(mean1stGPA=round(mean(GPA1st),2), count=n());cseallGPA
write.csv(cseallGPA,"CSEALL1stGPA.csv")

cseallGPARank <- cseall %>% mutate(APRGPA=ifelse(GPA1st>=2.77,"Above2.77", ifelse(GPA1st>=2.00,"2.00<=GPA<2.77","Beloww2.00")))



write.csv(cseallGPARank,"cseallGPARank.csv")
```

### comparison by departments
```{r}
# 2019
cse2019GPAdepart <- updated1CSEdata %>% select(COHORT_YEAR, FIRST_FALL_GPA,ENTRY_DEPARTMENT,ENTRY_PROGRAM ) 
cse2019GPAdepart[which(is.na(cse2019GPAdepart$FIRST_FALL_GPA)), "FIRST_FALL_GPA"] <- 0
cse2019GPAdepart1 <- cse2019GPAdepart %>% mutate(APRGPA=ifelse(FIRST_FALL_GPA>=2.77,"Above2.77", ifelse(FIRST_FALL_GPA>=2.00,"2.00<=GPA<2.77","Beloww2.00")))
colnames(cse2019GPAdepart1) <- c("Cohort","GPA1st","Entry_Depart","Entry_Prog","APRGPA")


#2020
cse2020GPAdepart <- CSE_enr2020 %>% select(COHORT_YEAR, GPA,ENTRY_DEPARTMENT,ENTRY_PROGRAM ) 
cse2020GPAdepart[which(is.na(cse2020GPAdepart$GPA)), "GPA"] <- 0
cse2020GPAdepart1 <- cse2020GPAdepart %>% mutate(APRGPA=ifelse(GPA>=2.77,"Above2.77", ifelse(GPA>=2.00,"2.00<=GPA<2.77","Beloww2.00")))
colnames(cse2020GPAdepart1) <- c("Cohort","GPA1st","Entry_Depart","Entry_Prog","APRGPA")

cse20192020GPAdepart <- rbind(cse2019GPAdepart1,cse2020GPAdepart1)
df2019 <-cse20192020GPAdepart %>% filter(Cohort==20192020) %>%  group_by(Entry_Depart) %>% summarise(DepartGPA2019=round(mean(GPA1st),2),count2019=n()) %>% arrange(-count2019);df2019

df2020 <-cse20192020GPAdepart %>% filter(Cohort==20202021) %>%  group_by(Entry_Depart) %>% summarise(DepartGPA2020=round(mean(GPA1st),2),count2020=n()) %>% arrange(-count2020);df2020


df20192020departGPA <- merge(df2020,df2019, by="Entry_Depart",all=T)
df20192020departGPA <- df20192020departGPA %>%  mutate(diffGPA = round(DepartGPA2020-DepartGPA2019, digits = 2), diffcount=count2020-count2019)

write.csv(df20192020departGPA,"df20192020departGPA.csv")
```

## table for grades distribution

```{r}
CSE2020re <- CSE2020[CSE2020$B_COURSE_NAME != 410228,]
tab2020CRS_Grds <- table(CSE2020re$B_COURSE_NAME, CSE2020re$B_GRADE_AWARDED)

tab2020.add.Grds_cut <- addmargins(tab2020CRS_Grds)   # sum for margin 1 and 2
### Convert table to data frmae
tab2020.add.Grds_cut_frame <- as.data.frame.matrix(tab2020.add.Grds_cut, row.names = F)
colnames(tab2020.add.Grds_cut_frame) <- paste("FT2020", colnames(tab2020.add.Grds_cut_frame), sep="_")
tab2020.add.Grds_cut_frame$FT2020CRSName <- rownames(tab2020.add.Grds_cut)

tab2020.add.Grds_cut_framere15 <- tab2020.add.Grds_cut_frame %>% filter(FT2020_Sum>=15)

listCRS <- tab2020.add.Grds_cut[(tab2020.add.Grds_cut$V2=="Sum" & tab2020.add.Grds_cut$N >=20),"V1"] %>% as.data.frame() 

#2019
CSE2019re <- CSE2019[CSE2019$A_COURSE_NAME != 410228,]
tab2019CRS_Grds <- table(CSE2019re$A_COURSE_NAME, CSE2019re$A_GRADE_AWARDED)
tab2019.add.Grds_cut <- addmargins(tab2019CRS_Grds) %>% as.data.table()   # sum for margin 1 and 2
tab2019.add.Grds_cut[(tab2019.add.Grds_cut$V1 != "Sum" & tab2019.add.Grds_cut$V2!="Sum"),]
colnames(tab2019.add.Grds_cut) <- c("CRSName2019","CRSGrade2019","Count2019")
tab2019.add.Grds_cut[tab2019.add.Grds_cut$CRSName2019=="23-Mar","CRSName2019"] <- "MAR3023"

# remove sum for each courses
listCRS_grads_cut <- merge(listCRS, tab2020.add.Grds_cut, by="V1", all.x=T) %>% filter(V2 != "Sum")
colnames(listCRS_grads_cut) <- c("CRSName2020","CRSGrade2020","Count2020")



tab2020.prop.Grds <- round(prop.table(tab2020CRS_Grds, margin = 1),2) %>% as.data.frame()
colnames(tab2020.prop.Grds) <- c("CRSName2020","CRSGrade2020","CRSProp2020")


write.csv(tab2020.prop.Grds,"tab2020_prop_Grds.csv")
write.csv(listCRS_grads_cut,"listCRS_grads_cut.csv")#2020
write.csv(tab2020.add.Grds_cut_framere30,"tab2020.add.Grds_cut_framere30.csv")
write.csv(tab2019.add.Grds_cut,"tab2019_add_Grds_cut.csv") #2019
```

