---
title: "V2-Scholarships - ACEX"
author: "PAM Lab - UWF"
date: "3/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
setwd("G:/My Drive/#1_Jay's Space/Scholarships-YOY")
library(xfun)
```

```{r eval=FALSE, include=FALSE}
# need to select variables and created matching ID with ACEX and data (filter SIF=="Stu_DEMO_DATA_SOURCE")
library(readr)
library(dplyr)
COURSE_STUDENT_INSTRUCTION_SPRING21 <- read_csv("G:/Shared drives/HMCSE-PAM Lab/DATA 202101/202101/COURSE_STUDENT_INSTRUCTION_SPRING21.csv") %>% filter(Course_DEMO_TIME_FRAME>=201505) %>% 
   select(STU_ID,Course_DEMO_TIME_FRAME,contains("term"),Course_CrsTitle, contains("combined"),Course_CrsPrefix,Course_CrsSuffix, contains("number"), contains("award"),contains("credit"))
write.csv(COURSE_STUDENT_INSTRUCTION_SPRING21, "mycrs_acex.csv")

STU_ENROLLMENT_SPRING21 <- read_csv("G:/Shared drives/HMCSE-PAM Lab/DATA 202101/202101/STU_ENROLLMENT_SPRING21.csv") %>%  filter(Stu_DEMO_DATA_SOURCE=="SIF" & Stu_DEMO_TIME_FRAME>= 201505) %>% 
  select(STU_ID, contains("demo"),contains("major"),Stu_CurrentTermLoad,contains("GPA"), contains("inst"), contains("total"),contains("credit"), contains("semester"),contains("coll"),contains("depart")) %>% arrange(Stu_DEMO_TIME_FRAME)
write.csv(STU_ENROLLMENT_SPRING21, "myenc_acex.csv")


STUDENT_DEGREE_SPRING21 <- read_csv("G:/Shared drives/HMCSE-PAM Lab/DATA 202101/202101/STUDENT_DEGREE_SPRING21.csv") %>%  select(STU_ID,  contains("data"), Deg_TermGranted, contains("desc"), contains("major"), contains("banner"))
write.csv(STUDENT_DEGREE_SPRING21, "mydeg_acex.csv")

CSE_MATCH_IDS_1_ <- read_csv("C:/Users/jsparks3/Downloads/CSE_MATCH_IDS (1).csv") # FTIC2015 to 2019

library(readxl)
ACEX_Data_by_Cohort2_22_21 <- read_excel("G:/Shared drives/HMCSE-PAM Lab/ACEX/ACEX Data by Cohort2-22-21.xlsx")
ACEX_Data_by_Cohort2_22_211 <- ACEX_Data_by_Cohort2_22_21[!is.na(ACEX_Data_by_Cohort2_22_21$`Aid Year`),]
#colSums(is.na(ACEX_Data_by_Cohort2_22_211))

Final_id_cohort <- Final_acex_enc_DF %>% select(uwfid, cohort_year) %>% mutate(Cohort=ifelse(cohort_year=="17-18", "FTIC2017",
                                                                                             ifelse(cohort_year=="18-19","FTIC2018",
                                                                                                    ifelse(cohort_year=="19-20", "FTIC2019","FTIC2020")))) %>% select(uwfid, Cohort)
Final_id_cohort_nodup <- Final_id_cohort[!duplicated(Final_id_cohort),] #1709
Final_acex_enc_DFtwo <- merge( Final_acex_enc_DF, Final_id_cohort_nodup, by=c("uwfid"), all.x = T)

Final_acex_enc_DF3 <- Final_acex_enc_DFtwo[!is.na(Final_acex_enc_DFtwo$Cohort),] %>% group_by(uwfid) %>% arrange(Index)
# export data
write.csv(Final_acex_enc_DF3, "V1_Final_acex_enc_DF.csv") 
```

## 1. Questions:

  1. How many FTIC with AXEC:
  
    a. maintained it and stayed at UWF?
    b. maintained it and left UWF?
    c. lost it and stayed at UWF?
    d. lost it and left UWF?
    
  2. What are the proportions of A and B+C+D?
  
  3. What factors best predict an FTIC with AXEC being in A as opposed to B, C, or D?


## 2. Requirements to Remain Eligible for Financial Aid

  * GPA: 3.00 cumulative UWF GPA
  * Full-time enrollment of 12+ credit hours (not summer)
  * At least 30 credit hours / academic year
  * Not be reinstated
  
 [Freshman Scholoarships-UWF](https://uwf.edu/admissions/undergraduate/cost-and-financial-aid/awards-and-scholarships/freshman/)

## 3. Factors should be Considered
- Prior Hours
- Gender
- CRS+GPA: maintained requirements
- Impact of switched majors: scholarship and extra hours (needs info)

## 4. Cleaning Data
- Total of 3,645 FTIC ACEX cases 
- Total of 1,935 unique FTICs in the ACEX data (FT17 to FT20)
- 555 FTIC ACEX (2020) were removed for lack of enrollment information (enrollments data ends Fall 2020)
- 36 UWFID with ACEX but not in CSE data were removed
- Total of 591 FTICs were removed from ACEX data 
- Total of 1,344 FTICs were used for this document
- For degree information, the most recent (Fall 2020) was used 
- For dual-degrees, the first alphabetically was used
- ID = 970522062 is FTIC 2017, but was mislabeled 2019; this was corrected


```{r read data, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(readr)
library(dplyr)
library(readxl)
#used dataname
myenc_acex <- read_csv("myenc_acex.csv", 
    col_types = cols(X1 = col_skip())) %>% filter(Stu_DEMO_TIME_FRAME== 201708 | Stu_DEMO_TIME_FRAME ==201808 | Stu_DEMO_TIME_FRAME ==201908 |Stu_DEMO_TIME_FRAME==202008)

myenc_acex_dropY1 <- read_csv("myenc_acex.csv") %>% filter(Stu_DEMO_TIME_FRAME>=201708 & Stu_DEMO_TIME_FRAME <=201808)

#used dataname
mycrs_acex <- read_csv("mycrs_acex.csv", 
    col_types = cols(X1 = col_skip()))


#used dataname
mydeg_acex <- read_csv("mydeg_acex.csv", 
    col_types = cols(X1 = col_skip()))

# Id match
CSE_MATCH_IDS_1_ <- read_csv("G:/My Drive/#1_Jay's Space/Scholarships-YOY/CSE_MATCH_IDS (1).csv") # FTIC2015 to 2019


#used dataname
mydeg_acex <- read_csv("mydeg_acex.csv", 
    col_types = cols(X1 = col_skip()))

# usded dataname


acex_cleandata <- read.csv("G:/My Drive/#1_Jay's Space/Scholarships-YOY/acex_cleandata.csv")

# used final data

Final_acex_enc_DF <- read_csv("V1_Final_acex_enc_DF.csv") %>% select(-1)


#length(unique(Final_acex_enc_DF$STU_ID))
#xtabs(~Final_acex_enc_DF$cohort_year+ Final_acex_enc_DF$Num_aidYear)
```


```{r eval=FALSE, include=FALSE}
head(acex_cleandata[,-1], 50)

```

## 5. Statistical Data

### Numbers of FTIC by Aid Years (before removed NA) 

```{r message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(scales)
acex_cleandata <- read.csv("G:/My Drive/#1_Jay's Space/Scholarships-YOY/acex_cleandata.csv")
#length(unique(acex_cleandata$uwfid))


library(plotly)
cleannames_ACEX1_fig1 <- acex_cleandata %>% group_by(cohort_year, Num_aidYear ) %>% dplyr::summarise(count=n()) %>% 
  plot_ly( x=~cohort_year, y=~ count, color=~ Num_aidYear, type="bar" ) %>% 
  layout( xaxis=list(title="COHORT YEAR"), yaxis=list(title="Counts by Aid Year"), barmode="stack")

cleannames_ACEX1_fig1

```

### ACEX Offered and Paid Amounts by FTIC (before removed NA)


```{r}
Summary_acex <- acex_cleandata %>% group_by(cohort_year, aid_year) %>% dplyr::summarise(Count=n(), SumOffered=sum(acex_offered), SumPaid=sum(acex_paid), .groups='drop') %>%  mutate(Diff_YOY= SumPaid -lag(SumPaid, default = SumPaid[1])) %>%  mutate(Diff_YOY1= Count -lag(Count, default = Count[1]))


SumOffered <- dollar(Summary_acex$SumOffered)
Summary_acex$SumOffered <- SumOffered
SumPaid <- dollar(Summary_acex$SumPaid)
Summary_acex$SumPaid <- SumPaid
Diff_YOY <- dollar(Summary_acex$Diff_YOY)
Summary_acex$Diff_YOY <- Diff_YOY

#%>%  arrange(-Count)
knitr::kable(Summary_acex, caption = " Numbers of Cohort vs. Sum of Aid Amounts (before NA)")
# loosing by Tier
```


### ACEX FTIC by Tiers (before removed NA)

```{r summary by tier, echo=FALSE, message=FALSE, warning=FALSE}

Summary_acex_tiers <- acex_cleandata %>% group_by(cohort_year,tier,Num_aidYear) %>% dplyr::summarise(Count=n(),SumOffered=sum(acex_offered), SumPaid=sum(acex_paid)) %>%  mutate(Diff_YOY= SumPaid -lag(SumPaid, default = SumPaid[1])) %>%  mutate(Diff_YOY1= Count -lag(Count, default = Count[1]))

SumOffered <- dollar(Summary_acex_tiers$SumOffered)
Summary_acex_tiers$SumOffered <- SumOffered
SumPaid <- dollar(Summary_acex_tiers$SumPaid)
Summary_acex_tiers$SumPaid <- SumPaid
Diff_YOY <- dollar(Summary_acex_tiers$Diff_YOY)
Summary_acex_tiers$Diff_YOY <- Diff_YOY

#Summary table
library(DT)
summary_acex_tiers_DT <- datatable(Summary_acex_tiers, extensions = "Buttons",caption = "ACEX-FTIC Trends by Tiers",
          filter = "top",
          options = list(dom="Blfrtip",buttons=c("copy","csv","excel","pdf","print"), lengthMenu=list(c(10,25,50,-1), c(10,25,50,"All")),pageLength=25) )
summary_acex_tiers_DT
```



```{r cleaned data, eval=FALSE, include=FALSE}
library(rvest)
library(tidyverse)
library(janitor)
colSums(is.na(cleannames_ACEX1))
cleannames_ACEX3 <- cleannames_ACEX1[, c(9:14)] %>% replace(is.na(.),0)
cleannames_ACEX4 <- cleannames_ACEX1[, c(1:8)]
cleannames_ACEX1 <- cbind(cleannames_ACEX4,cleannames_ACEX3) %>% group_by( uwfid ) %>%  mutate(Num_aidYear=row_number())
AidYear <- rep("AidYear")
cleannames_ACEX1$Num_aidYear <- as.factor(cleannames_ACEX1$Num_aidYear)
cleannames_ACEX1$Num_aidYear <- paste(AidYear,cleannames_ACEX1$Num_aidYear)
#write.csv(cleannames_ACEX1,"acex_cleandata.csv")
```


### No information in dataset STU_IDs


```{r echo=FALSE, message=FALSE, warning=FALSE}

acex_cleandata_STU_ID <- merge(acex_cleandata, CSE_MATCH_IDS_1_, by.x= "uwfid", by.y= "UNIV_ROW_ID", all.x = T ) %>% group_by(uwfid) %>% arrange(aid_year)%>%  mutate(Index=row_number())
#colSums(is.na(acex_cleandata_STU_ID))
cohort_size <- xtabs(~acex_cleandata_STU_ID$cohort_year+ acex_cleandata_STU_ID$Index)
acex_cleandata_STU_ID_17to19 <- acex_cleandata_STU_ID %>% filter(cohort_year != "20-21")
NAs <- acex_cleandata_STU_ID_17to19[which(is.na(acex_cleandata_STU_ID_17to19$STU_ID)),] %>% group_by(uwfid) %>% arrange(aid_year) %>% select(-Num_aidYear)
Missing <- xtabs(~NAs$cohort_year+NAs$aid_year)

acex_cleandata_STU_ID_17to19_nona <- acex_cleandata_STU_ID_17to19[!is.na(acex_cleandata_STU_ID_17to19$STU_ID),] #3,042
acex_id_only <- acex_cleandata_STU_ID_17to19_nona %>% select(uwfid, STU_ID)
all_id_acex_only <- acex_id_only[!duplicated(acex_id_only$uwfid),] #,1344

knitr::kable(Missing)
#NAs
```

### Total FTIC and Aid Years after removed NAs

```{r}
total_sum <- xtabs(~Final_acex_enc_DF$cohort_year + Final_acex_enc_DF$aid_year)
knitr::kable(total_sum, caption = "Numbers of ACEX Cohort vs. Aid Year")
```

```{r ACEX, eval=FALSE, include=FALSE}
CSE_MATCH_IDS_1_[CSE_MATCH_IDS_1_$STU_ID=="DC8FE3C332A7D51614C81674186529DAA8A35825","UNIV_ROW_ID"]
acex <- acex_cleandata[acex_cleandata$uwfid== "970398543", ]
acex
```

### Example: Lost ACEX but Maintaned at UWF

```{r merging two data set, eval=FALSE, include=FALSE}

myenc_acex_term_index <- myenc_acex %>% group_by(STU_ID) %>% arrange(Stu_DEMO_TIME_FRAME) %>%  mutate(Index=row_number())

# all about enc 3,642
acex_cleandata_STU_ID_enc <- merge(all_id_acex_only, myenc_acex_term_index, by=c("STU_ID"), all.x = T) %>% group_by(STU_ID) %>% arrange(Index) 

# all about acex 3,042
enc_acex_all_STU_ID <- merge(all_id_acex_only, acex_cleandata_STU_ID, by=c("uwfid","STU_ID"), all.x = T ) %>% group_by(STU_ID) %>% arrange(Index)

enc_acex_all_STU_ID2 <-merge(acex_cleandata_STU_ID_enc,enc_acex_all_STU_ID, by=c("uwfid","STU_ID","Index"), all.x = T)

ex1 <- enc_acex_all_STU_ID2[enc_acex_all_STU_ID2$STU_ID=="DC8FE3C332A7D51614C81674186529DAA8A35825",]
```

```{r}
library(DT)
EX1 <- Final_acex_enc_DF[Final_acex_enc_DF$STU_ID=="DC8FE3C332A7D51614C81674186529DAA8A35825",]

knitr::kable(EX1[, c(-1,-2,-3)])
```

```{r with degree info, eval=FALSE, include=FALSE}

deg_acex <- merge(all_id_acex_only, mydeg_acex, by="STU_ID", all.x = T) %>% group_by(STU_ID) %>% filter(Deg_TermGranted==max(Deg_TermGranted))
deg_acex1 <- deg_acex[!duplicated(deg_acex$STU_ID),]

enc_acex_all_STU_ID3 <- merge(enc_acex_all_STU_ID2, deg_acex1, by=c("STU_ID","uwfid"), all.x = T) %>% group_by(uwfid) %>% arrange(Index)
write.csv(enc_acex_all_STU_ID3, "Final_acex_enc_DF.csv")
```

```{r used Final_acex_enc_DF, eval=FALSE, include=FALSE}
ex_deg <- Final_acex_enc_DF[Final_acex_enc_DF$STU_ID=="DC8FE3C332A7D51614C81674186529DAA8A35825",]

ex_deg
```

### Example: Lost ACEX and left at UWF

```{r}
library(DT)
EX2 <- Final_acex_enc_DF[Final_acex_enc_DF$STU_ID=="74AAECD2F0EC58EA82CDE143B404AF0041E24C6E",]
knitr::kable(EX2[, c(-1,-2,-3)])
```
### Example: Remained ACEX and staying at UWF

```{r}
EX3 <- Final_acex_enc_DF[Final_acex_enc_DF$STU_ID=="3B6593E215D1346FFB4DC10580A270C8C86DA559",]
knitr::kable(EX3[, c(-1,-2,-3)])
```

### Example: Remained ACEX and graduated 

```{r}

EX4_a <- Final_acex_enc_DF[Final_acex_enc_DF$STU_ID=="111D28829BE5F0630859445BC0BF02028FD59654",]
knitr::kable(EX4_a[, c(-1,-2)])
```


### Number of UWF graduates in cohort (Fall 2020)

```{r echo=FALSE, message=FALSE, warning=FALSE}

# method 1 (had fill cohort_year 625 NAs that did not created when it merged)
Max_index <- Final_acex_enc_DF %>%   group_by(uwfid) %>%  top_n(1, Index) # choose max index by uwfid
Max_index$B_deg <- ifelse(Max_index$Deg_Desc=="Bachelor",1,0)
Max_index[which(is.na(Max_index$B_deg)), "B_deg"] <- 0 # for nas
Max_index <- Max_index %>% mutate(leftUWF=ifelse(B_deg ==1, "Bachelor" ,
                                                 ifelse((B_deg==0 & Stu_DEMO_TIME_FRAME != 202008), "Dropped", "Stayed")))
table_degree <-Max_index %>% group_by(Cohort,B_deg ) %>% dplyr::summarise(Count=n())
knitr::kable(table_degree)
```

  - Total 1,718 ACEX in FTIC 2017 
  - 516 unique FTIC 2017 earned ACEX in their first year
  - 151 FTIC earned bachelor degree by Fall 2020
  - 23 of 151 were in graduate programs


```{r cleaningFTIC2017, eval=FALSE, include=FALSE}
#create B-deg
FTIC2017_YOY <- Final_acex_enc_DF %>% filter(Cohort=="FTIC2017") %>% mutate(B_deg=ifelse(Deg_Desc=="Bachelor","1","0"))
FTIC2017_YOY[which(is.na(FTIC2017_YOY$B_deg)), "B_deg"] <- "0"
#create earned hours by year
FTIC2017_YOY_up <- FTIC2017_YOY %>% group_by(uwfid) %>% mutate(EarnedHrs_Year = Stu_TotalInstHours - lag(Stu_TotalInstHours, default = Stu_TotalInstHours[1])) %>%group_by(uwfid) %>%  mutate(TGP_Year=Stu_TotalInstGradePoints- lag(Stu_TotalInstGradePoints, default = Stu_TotalInstGradePoints[1])) %>% mutate(GPA_Year = round(TGP_Year/EarnedHrs_Year,digits = 4))
#colSums(is.na(FTIC2017_YOY_up))
negative_values <- FTIC2017_YOY_up %>% filter(EarnedHrs_Year<0) %>% select(uwfid, Index, Stu_DEMO_TIME_FRAME) %>%  mutate(Mater_UWF=rep("Master_UWF"))
#length(unique(negative_values$uwfid)) #23 in graduate level
#omit master enc info #1,718 master 23, so 1,695
Final_FTIC2017 <- anti_join(FTIC2017_YOY_up,negative_values, by=c("uwfid","Index","Stu_DEMO_TIME_FRAME"),all.x=T ) 

ACEX_FTIC2017 <- Final_FTIC2017 %>% group_by(uwfid) %>% mutate(GPACRD_Year= ifelse( (GPA_Year>=3.0000 & EarnedHrs_Year >= 30.0000), "On_Track",
                                                                                 ifelse((GPA_Year>=3.0000 & EarnedHrs_Year < 30.0000),"On_GPA",
                                                                                 ifelse((GPA_Year < 3.0000 & EarnedHrs_Year >= 30.0000), "On_CRD30",
                                                                                        ifelse((GPA_Year < 3.0000 & EarnedHrs_Year < 30.0000), "Off_Track","NA")))))

ACEX_FTIC2017[which(is.na(ACEX_FTIC2017$Num_aidYear)), "Num_aidYear"] <- "No_ACEX"
ACEX_FTIC2017[which(is.na(ACEX_FTIC2017$GPACRD_Year)), "GPACRD_Year"] <- "ACEX_1st"
Final_ACEX_FTIC2017 <- ACEX_FTIC2017
write.csv(Final_ACEX_FTIC2017, "ACEX_FTIC2017.csv")
```

### Numbers of On-Track FTIC by Terms

- On-Track: earned >=30 credits and kept GPA >= 3.00 per year
- Off-Track: did not earn >=30 credits and below GPA 3.00 per year
- On-CRD20: earned >=30 credits but GPA below 3.00 per year
- On-GPA: kept GPA >= 3.00 but earned less than 30 credits per year

```{r eval=FALSE, message=FALSE, warning=FALSE,  echo=FALSE, include=FALSE}
library(dplyr)
ACEX_FTIC2017 <- read.csv("G:/My Drive/#1_Jay's Space/Scholarships-YOY/ACEX_FTIC2017.csv", stringsAsFactors=TRUE) %>% select(-1)
strange <- ACEX_FTIC2017[ACEX_FTIC2017$GPACRD_Year=="ACEX_1st" & ACEX_FTIC2017$Stu_DEMO_TIME_FRAME=="202008",]
ACEX_FTIC2017[ACEX_FTIC2017$uwfid=="970518096" & ACEX_FTIC2017$Index=="3", "GPACRD_Year"] <- "Off_Track"
ACEX_FTIC2017[ACEX_FTIC2017$uwfid=="970527857" & ACEX_FTIC2017$Index=="3", "GPACRD_Year"] <- "Off_Track"
ACEX_FTIC2017[ACEX_FTIC2017$uwfid=="970525139" & ACEX_FTIC2017$Index=="4", "GPACRD_Year"] <- "Off_Track"
ACEX_FTIC2017 <- ACEX_FTIC2017 %>% mutate(GPACRD_Year1=ifelse(GPACRD_Year =="On_Track", "On_Track", "Off_Track")) 
#970525139
write.csv(ACEX_FTIC2017, "ACEX_FTIC2017_03_05_21.csv")

```


```{r message=FALSE, warning=FALSE, include=FALSE}
ACEX_FTIC2017_03_05_21 <- read.csv("G:/My Drive/#1_Jay's Space/Scholarships-YOY/ACEX_FTIC2017_03_05_21.csv", stringsAsFactors=TRUE) %>% select(-1)
Summary_track <- ACEX_FTIC2017_03_05_21 %>% group_by(GPACRD_Year,Stu_DEMO_TIME_FRAME) %>% dplyr::summarise(Count=n()) %>% 
  tidyr::pivot_wider(names_from = Stu_DEMO_TIME_FRAME, values_from=Count) %>% replace(is.na(.),0)
knitr::kable(Summary_track, caption = "CRD+GPA Trend for FTIC 2017")

```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
summary_ontrack <- ACEX_FTIC2017_03_05_21 %>% group_by(GPACRD_Year1, Stu_DEMO_TIME_FRAME) %>% dplyr::summarise(Count=n()) %>% 
  tidyr::pivot_wider(names_from = GPACRD_Year1, values_from= Count) %>% replace(is.na(.),0) %>% mutate(Not_Fall2020=516-(Off_Track+On_Track)) %>% 
  mutate(Prop_OnTrack= round(On_Track/516, digits = 4))
knitr::kable( summary_ontrack[ , -4], caption = "Proportion of On-track per Year (FTIC 2017)")

```

### Proportion of Maintained ACEX and On-track FTIC by Year 

```{r echo=FALSE, message=FALSE, warning=FALSE}
ACEX_ontrack <- ACEX_FTIC2017_03_05_21 %>%filter(Num_aidYear != "AidYear 1")  %>%  group_by(Num_aidYear, GPACRD_Year1) %>% dplyr::summarise(Count=n()) %>% 
  tidyr::pivot_wider(names_from = GPACRD_Year1, values_from=Count) %>% replace(is.na(.),0) %>% 
mutate(SumCountAidFT=(Off_Track+On_Track),  Prop_OnTrack= round(On_Track/SumCountAidFT, digits = 4), Prop_MaintainedACEX= round(SumCountAidFT/516, digits = 4), Prop_MaintainedACEXandOnTrack=round(On_Track/516, digits = 4))
knitr::kable(ACEX_ontrack[ -4,])

```
* 60.12% (208) of the ACEX were on-track in their 2nd year
* 58.21% (163) were on-track in their 3rd year
* 59.41% (120) were on-track in their 4th year


```{r Datashaping, eval=FALSE, include=FALSE}

length(unique(ACEX_FTIC2017_03_05_21$uwfid))#516
head(Datashaping_acex2017)
ACEX_FTIC2017_03_05_21$Deg_MajorCode <- as.numeric(ACEX_FTIC2017_03_05_21$Deg_MajorCode)
### data shaping
Datashaping_acex2017 <- ACEX_FTIC2017_03_05_21 %>% mutate(EarnedBDeg=ifelse(B_deg==0, "No","Yes")) %>% 
  transform(housing=ifelse(is.na(housing), 0, housing)) %>%  
  mutate(OfferedHousing=ifelse(housing>0, "Yes","No")) %>% 
  transform(Deg_Desc=ifelse(is.na(Deg_Desc), "NoDegInfo", Deg_Desc)) %>% 
  transform(acex_paid=ifelse(is.na(acex_paid), 0, acex_paid)) %>% 
   group_by(uwfid) %>% mutate(TotalACEXOffered=sum(acex_offered, na.rm = T),TotalACEXPaid=sum(acex_paid, na.rm = T) ) %>% group_by(uwfid) %>%           mutate(CumSumACEXPaid=cumsum(acex_paid)) %>% group_by(uwfid) %>% 
  mutate(UWFStatus=ifelse((B_deg == 1), "Graduated", "Conti")) %>% 
  mutate(LastFall=ifelse(UWFStatus== "Conti", max(Stu_DEMO_TIME_FRAME), Deg_TermGranted)) %>% 
  mutate(LastUWFStatus=ifelse((UWFStatus !="Graduated" & LastFall != 202008), "Dropped",ifelse(UWFStatus == "Graduated","Graduated", "Conti")))
### how to fil na

library(data.table)
Datashaping_acex2017DT <- setDT(Datashaping_acex2017)[ , comptier := unique(tier[!is.na(tier)]), by=uwfid]

### checking  
Datashaping_acex2017DT[Datashaping_acex2017DT$uwfid=="970520853",]# conti
Datashaping_acex2017DT[Datashaping_acex2017DT$uwfid=="970522062",]# graduated
Datashaping_acex2017DT[Datashaping_acex2017DT$uwfid=="970398552",]# NotFall2020


### saving
DataView <- Datashaping_acex2017DT[, c(1,2,3,50:57)]



write.csv(Datashaping_acex2017DT,"ACEX_FTIC2017_03_08_21.csv")


```

### Summary of ACEX FTIC 2017 by UWF Status (Fall 2020)

```{r readdata, echo=FALSE, message=FALSE, warning=FALSE}
ACEX_FTIC2017_03_08_21 <- read.csv("G:/My Drive/#1_Jay's Space/Scholarships-YOY/ACEX_FTIC2017_03_08_21.csv", stringsAsFactors=TRUE)

ACEX_MAX_index <- ACEX_FTIC2017_03_08_21 %>% group_by(uwfid) %>% filter(Index==max(Index)) %>% group_by(LastUWFStatus,"LastGPACRD"=GPACRD_Year1 ) %>% 
  dplyr::summarise(MeanUnivHrsbyFall2020=mean(Stu_TotalUniversityHours, na.rm=T, .groups='drop'),MeanUWFHrsbyFall2020=mean(Stu_TotalInstHours),MeanUWFHrsbyFall2020=mean(EarnedHrs_Year),MeanGPAbyFall2020=mean(GPA_Year, na.rm=T),Count=n(), TotalACEXPaid1byFall2020= sum(TotalACEXPaid, na.rm=T))
TotalACEXPaid1byFall2020 <- dollar(ACEX_MAX_index$TotalACEXPaid1byFall2020)
ACEX_MAX_index$TotalACEXPaid1byFall2020 <- TotalACEXPaid1byFall2020


knitr::kable(ACEX_MAX_index, caption = "Summary of ACEX FTIC 2017 Status by On/off-Track (Fall 2020)")
```

### Summary ACEX FTIC 2017 at thier 1st Year by UWF Status and Tiers

```{r echo=FALSE, message=FALSE}
ACEX_1stTerm_index <- ACEX_FTIC2017_03_08_21 %>%  filter(Index==1) %>% group_by(B_deg,LastUWFStatus,comptier ) %>% 
  dplyr::summarise(MeanUnivPrioHrs=mean(Stu_TotalUniversityHours, na.rm=T), Mean1stUWFPriorHrs=mean(Stu_TotalInstHours, na.rm=T), MeanUWFHrsby1stTerm=mean(Stu_GPATermHours,  na.rm=T), MeanGPAby1stTerm=mean(Stu_GPASemester, na.rm=T), Count=n(), SumACEXPaid= sum(acex_paid), .groups="drop") %>% group_by(LastUWFStatus) %>% mutate(SumACEXPaidbyStatus=sum(SumACEXPaid))
SumACEXPaid <- dollar(ACEX_1stTerm_index$SumACEXPaid)
ACEX_1stTerm_index$SumACEXPaid <- SumACEXPaid
SumACEXPaidbyStatus <- dollar(ACEX_1stTerm_index$SumACEXPaidbyStatus)
ACEX_1stTerm_index$SumACEXPaidbyStatus <- SumACEXPaidbyStatus

knitr::kable(ACEX_1stTerm_index, caption = "Summary of ACEX FTIC 2017 Status by Tiers (Fall 2020)")

```



```{r message=FALSE, include=FALSE}
### Summary ACEX FTIC 2017 at thier 1st Year by UWF Status and Aid Years

ACEX_aidYears_index <- ACEX_FTIC2017_03_08_21  %>% group_by(Stu_DEMO_TIME_FRAME,Num_aidYear, LastUWFStatus, B_deg ) %>% 
  dplyr::summarise(MeanUnivHrsbyYear=mean(Stu_TotalUniversityHours, na.rm=T),  MeanUWFHrsbyYear=mean(EarnedHrs_Year,  na.rm=T), MeanGPAby1stYear=mean(GPA_Year, na.rm=T), Count=n(), SumACEXPaid= sum(acex_paid, na.rm=T), .groups="drop") %>% group_by(Num_aidYear) %>% mutate(SumCumACEXPaidbyAidYear=sum(SumACEXPaid, na.rm=T))
SumACEXPaid <- dollar(ACEX_aidYears_index$SumACEXPaid)
ACEX_aidYears_index$SumACEXPaid <- SumACEXPaid
SumCumACEXPaidbyAidYear <- dollar(ACEX_aidYears_index$SumCumACEXPaidbyAidYear)
ACEX_aidYears_index$SumCumACEXPaidbyAidYear <- SumCumACEXPaidbyAidYear

knitr::kable(ACEX_aidYears_index, caption = "Summary of ACEX FTIC 2017 by Aid Years (Fall 2020)")

```

```{r mergedata with CSE, include=FALSE}
CSE_ALL_ENR_2017<- read.csv("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/DATA/CSE_ALL_ENR_up2020.csv", stringsAsFactors=TRUE) %>% 
  filter(COHORT_YEAR==20172018) %>% select(STU_ID,"PriorHours"=HOURS_BROUGHT_TO_UNIVERSITY, AP_CREDITS,GENDER,ETHNICITY,ENTRY_PROGRAM,"ENTRY_DEPARTMENT"=codeDEPART, PSE,codeCOUNTY,FIRST_GENERATION_STUDENT,ATHLETE,GPA_HIGHSCHOOL,HIGH_SCHOOL_NAME, ACT_PROPORTION,SAT_PROPORTION,FIRST_FALL_GPA, APR) %>% 
  mutate(BestTest=pmax(SAT_PROPORTION, ACT_PROPORTION))


CSE_ALL_ENR_20171 <- CSE_ALL_ENR_2017[!duplicated(CSE_ALL_ENR_2017$STU_ID),]

ACEX_CSE_2017 <- merge(ACEX_FTIC2017_03_08_21, CSE_ALL_ENR_20171, by="STU_ID", all.x = T) %>% select(-2) %>% 
  mutate(OnTrackwithAid=ifelse((GPACRD_Year1=="On_Track" & acex_paid >0), "OnTrackwithAid_Yes","OnTrackwithAid_No")) %>% 
  mutate(codeEthnicity=ifelse(ETHNICITY=="African American","African American",
                              ifelse(ETHNICITY=="Asian","Asian",
                                     ifelse(ETHNICITY=="Hispanic","Hispanic",
                                            ifelse(ETHNICITY=="White","White","Others")))))

ACEX_CSE_2017[ACEX_CSE_2017$BestTest ==0.141666, "BestTest"] <- 0.458333 # outliers
ACEX_CSE_2017[ACEX_CSE_2017$ETHNICITY =="Test", "ETHNICITY"] <- "Not Reported" # outliers
ACEX_CSE_2017[ACEX_CSE_2017$PSE =="NO", "PSE"] <- "No" # outliers
ACEX_CSE_2017[ACEX_CSE_2017$PSE =="YES", "PSE"] <- "Yes" # outliers

### saving data
write.csv(ACEX_CSE_2017, "ACEX_CSE_2017_03_09_21.csv")
library(Hmisc)
Hmisc::describe(ACEX_CSE_2017[, c(-1,-2,-4,-5,-6)])
```

```{r read new data, include=FALSE}
 ACEX_CSE_2017_03_09_21 <- read.csv("G:/My Drive/#1_Jay's Space/Scholarships-YOY/ACEX_CSE_2017_03_09_21.csv", stringsAsFactors=TRUE)
hist(ACEX_CSE_2017_03_09_21$BestTest)
glimpse(ACEX_CSE_2017_03_09_21)
xtabs(~ACEX_CSE_2017_03_09_21$Stu_DEMO_TIME_FRAME+ACEX_CSE_2017_03_09_21$B_deg +ACEX_CSE_2017_03_09_21$OnTrackwithAid)

## select the last term at uwf
ACEX_CSE_myvariables <- ACEX_CSE_2017_03_09_21 %>%  filter(Index==2) %>%  select(OnTrackwithAid,BestTest,GPA_HIGHSCHOOL,PriorHours,AP_CREDITS,HIGH_SCHOOL_NAME,county,GENDER,codeEthnicity, Stu_CollegeCode,Stu_DepartmentCode ,Stu_MajorDesc,PSE,FIRST_FALL_GPA)

## Dropped Before 2nd Fall 54 are not included
ACEX_CSE_myvariables$OnTrackwithAid <- as.factor(ACEX_CSE_myvariables$OnTrackwithAid)
ACEX_CSE_myvariables$codeEthnicity <- as.factor(ACEX_CSE_myvariables$codeEthnicity)
glimpse(ACEX_CSE_myvariables)
```

```{r eval=FALSE, include=FALSE}
## Who is more likely On-Track or Graduated ?

* Considering the last status at UWF
  
  + Total ACEX awarded in their 2nd Year
  + Majors
  + Departments
  + Gender
  + Ethnicity
  + Prior Hours
  + AP credits
  + High school names
  + County
  + 1st Term GPA
  + HS GPA


### Methods: Decision Tree
  

#####################
#### Index ==2 ######
###################
library(party)
library(rpart)
library(rpart.plot)
library(dplyr)
library(caret)
library(e1071)
#Data partition
# Performs stratified random split of the data set
TrainingIndex <- createDataPartition(ACEX_CSE_myvariables$OnTrackwithAid, p=0.8, list = FALSE)
TrainingSet <- ACEX_CSE_myvariables[TrainingIndex,] # Training Set 677
TestingSet <- ACEX_CSE_myvariables[-TrainingIndex,] 
set.seed(123)
treeACEXOnTrack <- ctree(OnTrackwithAid ~ ., data=TrainingSet, controls = ctree_control(mincriterion = 0.9,minsplit = 10))
treeACEXOnTrack
```

```{R eval=FALSE, include=FALSE}
### Plot

plot(treeACEXOnTrack)


### Confusion Matrix and Statistics

#confusion matrix

confusionMatrix(predict(treeACEXOnTrack), TrainingSet$OnTrackwithAid)

### Methods: Naive Bayes Classification


library(rsample)  # data splitting 
library(dplyr)    # data transformation
library(ggplot2)  # data visualization
library(caret)    # implementing with caret
#library(h2o)      # implementing with h2o
library(naivebayes)
NaiveACEXOnTrack <- naive_bayes(OnTrackwithAid ~., 
                                data = TrainingSet, usekernel = TRUE)
confusionMatrix(predict(NaiveACEXOnTrack), TrainingSet$OnTrackwithAid)


### Tables

tables(NaiveACEXOnTrack)
```

```{r included dropped at their 1st year, include=FALSE}
Index_1_ids <- ACEX_CSE_2017_03_09_21 %>% filter( Index==1) %>% select( STU_ID, uwfid,B_deg)
Index_1_ <- ACEX_CSE_2017_03_09_21 %>% filter( Index==1) 

Index_1_enc_max <- merge(myenc_acex_dropY1,Index_1_ids, by="STU_ID", all.y = T) %>% group_by(STU_ID) %>% filter(Stu_DEMO_TIME_FRAME==max(Stu_DEMO_TIME_FRAME) | Stu_DEMO_TIME_FRAME==201708) %>% arrange(uwfid,Stu_DEMO_TIME_FRAME ) %>% 
  mutate(reEarnedHrs_Year=Stu_TotalInstHours-lag(Stu_TotalInstHours, default = Stu_TotalInstHours[1]),reTGP_Year=Stu_TotalInstGradePoints-lag(Stu_TotalInstGradePoints, default = Stu_TotalInstGradePoints[1]),reGPA_Year=reTGP_Year/reEarnedHrs_Year) %>% 
  mutate(reTotalUnivHrsY1=Stu_TotalUniversityHours, reTotalUWFHrsY1=Stu_TotalInstHours, reOnTrackwithAid=ifelse((reGPA_Year>=3.00 & reEarnedHrs_Year>= 30.00 ),"OnTrackwithAid_Yes","OnTrackwithAid_No" )) %>% group_by(uwfid) %>% filter(Stu_DEMO_TIME_FRAME==max(Stu_DEMO_TIME_FRAME)) %>% select(21:28)

### Last dataset
Index_1_Y1516 <- merge(Index_1_enc_max, Index_1_, by="uwfid", all=T ) %>% select(-X,-GPACRD_Year1,-OnTrackwithAid)
aa <- Index_1_Y1516[(Index_1_Y1516$reOnTrackwithAid=="OnTrackwithAid_Yes" & Index_1_Y1516$comptier==4),] 
aa
### checking
Index_1_Y1516[Index_1_Y1516$uwfid=="970398552",] #OnTrackwithAid_No
Index_1_Y1516[Index_1_Y1516$uwfid=="970399429",] #OnTrackwithAid Yes

glimpse(Index_1_enc_max)

#######################
##### My variables 516 ###
#####################
ACEX_Year_1_OnTrack_vari <- Index_1_Y1516 %>% select(reOnTrackwithAid,reTotalUnivHrsY1,OfferedHousing, BestTest,GPA_HIGHSCHOOL,PriorHours,AP_CREDITS,HIGH_SCHOOL_NAME,county,GENDER,codeEthnicity, Stu_CollegeCode,Stu_DepartmentCode ,Stu_MajorDesc,PSE,FIRST_FALL_GPA, acex_paid,OfferedHousing,meal_plan,flms_paid,flut_paid)


ACEX_Year_1_OnTrack_vari$reOnTrackwithAid <- as.factor(ACEX_Year_1_OnTrack_vari$reOnTrackwithAid)
glimpse(Index_1_Y1516)
```

### ACEX Aid Difference between On / Off Track  FTIC 2017 (graduates were included in On-Track)

```{r echo=FALSE, message=FALSE}
Summary_516 <- Index_1_Y1516 %>% group_by(reOnTrackwithAid,B_deg.y ) %>%  dplyr::summarise(Count=n(), SumACEXPaid=sum(acex_paid),   GPA1stTerm=mean(FIRST_FALL_GPA,na.rm=T),                                                 meanEarnedHrsY1=mean(reEarnedHrs_Year), meanGPAY1=mean(reGPA_Year, na.rm=T)) %>% group_by(reOnTrackwithAid) %>% mutate(SumACEXbyMaintained=sum(SumACEXPaid))
SumACEXPaid <- dollar(Summary_516$SumACEXPaid)
Summary_516$SumACEXPaid <- SumACEXPaid
SumACEXbyMaintained <- dollar(Summary_516$SumACEXbyMaintained)
Summary_516$SumACEXbyMaintained <- SumACEXbyMaintained
Summary_516_DT <- as.data.frame(Summary_516)
knitr::kable(Summary_516)
```

### Plot - On-Off Track at their 1st Year by Tiers

```{r echo=FALSE, message=FALSE, warning=FALSE}

library(plotly)
cleannames_ACEX1_fig2 <- Index_1_Y1516 %>% group_by(reOnTrackwithAid, comptier ) %>% dplyr::summarise(count=n()) %>% 
  plot_ly( x=~reOnTrackwithAid, y=~ count, color=~ comptier, type="bar" ) %>% 
  layout( xaxis=list(title="On/Off Track"), yaxis=list(title="ACEX On/Off Track at their 1st Year by Tiers"), barmode="stack")

cleannames_ACEX1_fig2
```

### Summary Table for Graduates among On/Off Track at their 1st Year

```{r echo=FALSE}

DT_1 <- addmargins(table(Index_1_Y1516$reOnTrackwithAid, Index_1_Y1516$EarnedBDeg))
DT <- as.data.frame.matrix(DT_1)
DT$OnTrackwithAid <- rownames(DT_1)
DT2 <- DT %>% select(OnTrackwithAid,Yes,No, Sum ) %>% mutate(Prop_Gradu = round(Yes/Sum, digits = 4))
knitr::kable(DT2[, -1])
```


## 6. Conclusion


* The situation with ACEX is complex considering that some students don't meet the requirements to maintain it, but still keep it
* So, we considered only their 1st year to simply interpret the impact of ACEX
* For the impact of ACEX, we determined who earned a bachelor degree by Fall 2020 to show the impact of aid
* In the 2017 cohort, 516 FTIC started with an ACEX
* After their first year, 218 FTIC still met the ACEX requirements (i.e. 298 FTIC didn't meet the requirements)
* Of those 218 that met the requirements in their first year, 97 graduated by Fall 2020
* In other words,  18.78% of the 2017 FTIC cohort (or 44.5% of those who maintained through their first year) finished their degrees while maintaining ACEX requirements
* Additionally, 54 (10.4% of the 2017 FTIC cohort) started with an ACEX (or 18.12% of those who did not maintain requirements), but graduated by Fall 2020

  + In the future, we need to capture another cohort to more accurately see the trends
  + Additionally, more details about the award processes are needed to explain those who did not meet the requirements but kept the award
