---
title: "R Change college, Department, and Program"
output:
  html_document:
    df_print: paged
---
### import data
```{r}

setwd("C:/Users/jsparks3/Documents/Data_IR")
library(dplyr)
deg0 <- read.csv("G:/Shared drives/HMCSE-PAM Lab/DATA 202001/STUDENT_DEGREE_SPRING20.csv")

degspring2020 <- read.csv("G:/Shared drives/HMCSE-PAM Lab/DATA 202001/STUDENT_DEGREE_SPRING20.csv") %>%
  filter(Deg_DEMO_TIME_FRAME>=200805 ) %>% 
  group_by(STU_ID) %>% filter(Deg_Desc=="Bachelor", .by_group=TRUE) %>% mutate(number_Bachelor=row_number(), endDay=rep("01")) %>% 
   select(1:2,contains("college"),contains("Department"),contains("type"),contains("Semester"),contains("CIP"),contains("GPA"),contains("Major"),
          Deg_TermDate,Deg_Desc,contains("Major"),number_Bachelor)

library(stringr)
library(tidyr)
### FTIC filter with first term only
enr<- read.csv("G:/Shared drives/HMCSE-PAM Lab/DATA 202001/STU_ENROLLMENT_SPRING20.csv")
enrspring2020 <- read.csv("G:/Shared drives/HMCSE-PAM Lab/DATA 202001/STU_ENROLLMENT_SPRING20.csv") %>% 
  filter(Stu_LoadIPEDSFTPT=="Full Time" & # full time
             Stu_StudentTypeCode =="B" &
           Stu_AdmissionRecentTypeCode=="B" & Stu_DEMO_TIME_FRAME >= 200805 ) %>% # data quality since 2008
  group_by(STU_ID) %>% 
  filter(str_detect(Stu_Term, c("Fall..."))) %>% 
  separate(Stu_AdmissionTerm, into = c("Term","Cohort"),sep=" ") %>% 
  select(1,Stu_TermDate,Cohort) %>% filter(Cohort >= 2008)
  xtabs(~enrspring2020$Cohort )
```


```{r}
  ### Merge all fall semesters information and create index that not include summer term
FTIC_all_semesterFall <- merge(enrspring2020, enr, by="STU_ID", all.x = TRUE ) %>% 
  #filter(Stu_AdmissionRecentTypeCode == "B") %>% 
   filter(str_detect(.$Stu_Term , "Fall")) %>%  # No summer terms
   group_by(STU_ID) %>%  arrange(Stu_DEMO_TIME_FRAME, .by_group=TRUE) %>% 
  mutate(number_semester=row_number()) %>% 
  select(1:5,contains("ethnicity"),contains("gender"),contains("county"),contains("Nation"),contains("college"),contains("Department"),contains("type"),contains("Semester"),contains("term"),contains("CIP"),contains("GPA"),
         contains("Hour"),contains("Insti"),contains("Major"),Cohort,number_semester)


FTIC_all_semesterFall$Term = rep("Fall") 

FTIC_all_semesterSpring <- merge(enrspring2020, enr, by="STU_ID", all.x = TRUE ) %>% 
  #filter(Stu_AdmissionRecentTypeCode == "B") %>% 
   filter(str_detect(.$Stu_Term , "Spring")) %>%  # No summer terms
   group_by(STU_ID) %>%  arrange(Stu_DEMO_TIME_FRAME, .by_group=TRUE) %>% 
  mutate(number_semester=row_number()) %>% 
  select(1:5,contains("ethnicity") ,contains("gender"),contains("county"),contains("Nation"), contains("college"),contains("Department"),contains("type"),contains("Semester"),contains("term"),contains("CIP"),contains("GPA"),
         contains("Hour"),contains("Insti"),contains("Major"),Cohort,number_semester)
    
FTIC_all_semesterSpring$Term = rep("Spring") 


FTIC_all_semesterSummer <- merge(enrspring2020, enr, by="STU_ID", all.x = TRUE ) %>% 
  #filter(Stu_AdmissionRecentTypeCode == "B") %>% 
   filter(str_detect(.$Stu_Term , "Summer")) %>%  # No summer terms
   group_by(STU_ID) %>%  arrange(Stu_DEMO_TIME_FRAME, .by_group=TRUE) %>% 
  mutate(number_semester=row_number()) %>% 
  select(1:5,contains("ethnicity") ,contains("gender"),contains("county"), contains("Nation"),contains("college"),contains("Department"),contains("type"),contains("Semester"),contains("term"),contains("CIP"),contains("GPA"),
         contains("Hour"),contains("Insti"),contains("Major"),Cohort,number_semester)
    
FTIC_all_semesterSummer$Term = rep("Summer") 

FTIC_all_semester <- rbind(FTIC_all_semesterFall,FTIC_all_semesterSpring,FTIC_all_semesterSummer)

```


```{r} 

FTIC_all_semester1  <- FTIC_all_semester %>%  group_by(STU_ID) %>%  arrange(Stu_DEMO_TIME_FRAME, .by_group=TRUE) %>% 
  mutate(Total_number_semester=row_number())

xtabs(~FTIC_all_semester1$Cohort+FTIC_all_semester1$Term)
```
```{r}
xtabs(~FTIC_all_semester1$Cohort+FTIC_all_semester1$number_semester+FTIC_all_semester1$Term )
```


```{r}
### get the bachelor degree information and merge with enr
FTIC_all_semester_deg <- merge(FTIC_all_semester1, degspring2020, by="STU_ID", all.x = TRUE)

  sub_B <- FTIC_all_semester_deg %>%  filter(Stu_StudentTypeCode=="B" )
  xtabs(~sub_B$Cohort+sub_B$number_Bachelor)

```


```{r}
FTIC_all_semester_deg$Stu_TermDate <- as.Date.character(FTIC_all_semester_deg$Stu_TermDate.x)
FTIC_all_semester_deg$Deg_TermDate <- as.Date.character(FTIC_all_semester_deg$Deg_TermDate)

# create col for data
FTIC_all_semester_deg$Days <- as.character(round(FTIC_all_semester_deg$Deg_TermDate - as.Date(FTIC_all_semester_deg$Stu_TermDate)))

FTIC_all_semester_deg$Months <- as.character(round(FTIC_all_semester_deg$Deg_TermDate - as.Date(FTIC_all_semester_deg$Stu_TermDate))/(365.25/12))

FTIC_all_semester_deg$Years <- as.character(round(FTIC_all_semester_deg$Deg_TermDate - as.Date(FTIC_all_semester_deg$Stu_TermDate))/(365))
FTIC_all_semester_deg$Ontime_gradu <- ifelse(FTIC_all_semester_deg$Years < 4.00, "Ontime","Overtime")

FTIC_all_semester_deg$Semester <- as.character(round(FTIC_all_semester_deg$Deg_TermDate - as.Date(FTIC_all_semester_deg$Stu_TermDate))/(365/3))

#merge with deg and start info
FTIC_all_semester_deg_semester1 <- FTIC_all_semester_deg %>% group_by(STU_ID) %>% filter(number_semester==max(number_semester))

xtabs(~FTIC_all_semester_deg_semester1$Cohort+FTIC_all_semester_deg_semester1$Ontime_gradu)
```


```{r}
### changing variables from start and end  for changes
FTIC_all_semester_deg_semester1 <- FTIC_all_semester_deg_semester1[!is.na(FTIC_all_semester_deg_semester1$Ontime_gradu), ] %>% 
 mutate(ChangeCIP= ifelse(Stu_ProgramCIPCode==Deg_ProgramCIPCode, "No","Yes"), ChangeColl= ifelse(Stu_CollegeCode==Deg_CollegeCode, "No","Yes"),ChangeDeprt= ifelse(Stu_DepartmentCode==Deg_DepartmentCode, "No","Yes"))
xtabs(~FTIC_all_semester_deg_semester1$Ontime_gradu)
xtabs(~FTIC_all_semester_deg_semester1$Ontime_gradu+FTIC_all_semester_deg_semester1$ChangeColl)
xtabs(~FTIC_all_semester_deg_semester1$Ontime_gradu+FTIC_all_semester_deg_semester1$ChangeDeprt)
xtabs(~FTIC_all_semester_deg_semester1$Ontime_gradu+FTIC_all_semester_deg_semester1$ChangeCIP)
xtabs(~FTIC_all_semester_deg_semester1$Cohort +FTIC_all_semester_deg_semester1$ChangeDeprt)
```

```{r}
library(ggplot2)
myCIPMAXsemester <- FTIC_all_semester_deg_semester1
p_deg <- ggplot(data=myCIPMAXsemester)
P_Time <- p_deg+ geom_bar(aes(x=number_semester, color=Cohort), size=1.2)+ facet_grid(Ontime_gradu~. ) + labs(title = "Number of Semesters by Cohort (included summer term)", subtitle = "source(PAM Lab 2020)")
           
P_Time


```


### each colleges and department
```{r}
library(tidyr)
myCIPMAXsemester_N <- myCIPMAXsemester %>% group_by(Deg_College, ChangeColl, Ontime_gradu) %>% summarise(meanSemester=mean(number_semester), .groups="drop");myCIPMAXsemester_N



```
### changed departments
```{r}
myCIPMAXsemester_ND <- myCIPMAXsemester %>% group_by(Deg_Department, ChangeDeprt, Ontime_gradu) %>% summarise(meanSemester=mean(number_semester), .groups="drop");myCIPMAXsemester_ND

```

```{r}
meanvalue <- myCIPMAXsemester %>% 
  group_by( Deg_College,  Ontime_gradu) %>% 
  summarise(meanSemester=mean(number_semester, na.rm = T), CountFTIC=n()); meanvalue
```


```{r}
myCIPmaxminsemester_OntimePlot <- ggplot(myCIPMAXsemester[!is.na(myCIPMAXsemester$Ontime_gradu) & myCIPMAXsemester$Stu_TotalInstHours>=100,], aes(Stu_TotalInstHours, Stu_TotalUniversityHours, colour = Ontime_gradu)) + geom_point() +labs(title = "UWF Hours vs. University Hours by Degree Completion") 
myCIPmaxminsemester_OntimePlot


```


```{r}
p_ontime <- ggplot(data = myCIPMAXsemester_N[myCIPMAXsemester_N$Ontime_gradu=="Ontime",])
p_ontime + geom_point(aes(x=Deg_College,y=meanSemester,color=ChangeColl),size=5)+labs(xlabs="College", ylabs="Average of Semester", title="Average Semester for Ontime degree FTIC", subtitle = "Source: PAM Lab 2020") +facet_grid(ChangeColl~.)

p_overtime <- ggplot(data = myCIPMAXsemester_N[myCIPMAXsemester_N$Ontime_gradu=="Overtime",])
p_overtime + geom_point(aes(x=Deg_College,y=meanSemester,color=ChangeColl),size=5)+labs(xlabs="College", ylabs="Average of Semester", title="Average Semester for Overtime degree FTIC", subtitle = "Source: PAM Lab 2020")+facet_grid(ChangeColl~.)
```

```{r}


myCIPmaxminsemester_OntimePlot1 <- ggplot(myCIPMAXsemester[complete.cases(myCIPMAXsemester$Ontime_gradu),], aes(x=Stu_TotalInstHours)) + geom_histogram(aes(fill=ChangeColl)) +facet_grid(Ontime_gradu~.) +labs(title = "UWF Hours  by Degree Completion and Changed College") 
myCIPmaxminsemester_OntimePlot1

myCIPmaxminsemester_OntimePlot2 <- ggplot(myCIPMAXsemester[complete.cases(myCIPMAXsemester$Ontime_gradu),], aes(x=Stu_TotalInstHours)) + geom_histogram(aes(fill=ChangeDeprt)) +facet_grid(Ontime_gradu~.) +labs(title = "UWF Hours  by Degree Completion and Changed College") 
myCIPmaxminsemester_OntimePlot2

myCIPmaxminsemester_OntimePlot3 <- ggplot(myCIPMAXsemester[complete.cases(myCIPMAXsemester$Ontime_gradu),], aes(x=Stu_TotalInstHours)) + geom_histogram(aes(fill=ChangeCIP)) +facet_grid(Ontime_gradu~.) +labs(title = "UWF Hours  by Degree Completion and Changed College") 
myCIPmaxminsemester_OntimePlot3


```
### no degree FTIC

```{r}
str(myCIPMAXsemester)
myCIPMAXsemesterNodeg <- myCIPMAXsemester %>% group_by(STU_ID) %>% filter( number_semester==max(number_semester))
which(is.na(FTIC_all_semester_deg$Ontime_gradu))
```
### Using Term index dataset
### Noitce that GPA HS values are missing for updated 2018:2019 Cohort
```{r}
FTIC_Fall_semester2 <- FTIC_all_semester1 %>% filter(Term == "Fall") %>% select(1:3,contains("ethnicity") ,contains("gender"), contains("Nation"),number_semester,contains("College"),contains("county"), contains("department"),Stu_Department,contains("Major"),contains("HighSchool"),contains("Total"),Stu_TotalInstHours,Stu_GPAUndergraduateTerm,Stu_GPATermHours,Stu_CurrentTermLoad,Stu_GPASemester, contains("gradepoint")) %>% 
  mutate(GPAT1= Stu_GPAGradePoints/Stu_GPATermHours) %>% mutate(diffGPAt1=round(GPAT1-Stu_GPAUndergraduateTerm, digits = 2))


# fitler the first fall 
FTIC_Fall_semester3 <- select(FTIC_Fall_semester2,-Total_number_semester) %>% filter(Cohort >= 2015 & number_semester == 1) ;glimpse(FTIC_Fall_semester3)
```

Merge with CRS data to combine the first term courses
```{r}
my_id_3from3 <- FTIC_Fall_semester3 %>%  select(1:3) %>% group_by(STU_ID)

my_id_crs <- COURSE_STUDENT_INSTRUCTION_SPRING20 %>%  select(1,5,"Course_CrsCombined","Course_CrsGradeAwarded", contains("credit"))

FTIC_frist_crs <- merge(my_id_3from3,my_id_crs, by.x=c("STU_ID", "Stu_TermDate.x"), by.y=c("STU_ID","Course_TermDate"), all.x = T)
#check duplication
library(tidyverse)
FTIC_frist_crs_nodup <- FTIC_frist_crs %>% distinct(STU_ID,Stu_TermDate.x,Course_CrsCombined, Course_CrsGradeAwarded,  .keep_all = T)
FTIC_frist_crs_keep <- FTIC_frist_crs_nodup %>% select(1:7) %>% group_by(STU_ID) %>% mutate(Number_Crs=row_number());FTIC_frist_crs_keep[1:10,]
length(unique(FTIC_frist_crs_keep$Number_Crs)) # all first term courses

# merge with orig data
FTIC_Fall_semester_all_CRS <- merge(FTIC_Fall_semester3, FTIC_frist_crs_keep, by="STU_ID", all.y = T)# this is FTCI 2015-2019 with first term GPA and CRS info

```
This is GradePoint  in data set
```{r}
unique(FTIC_Fall_semester_all_CRS$Course_CrsGradeAwarded)
# [1] "C"  "C-" "B"  "A"  "A-" "B+" "C+" "D"  "D+" "XX" "W"  "F"  "B-" "NF" "I*" "I"  "P"  "S" 

FTIC_Fall_semester_all_CRS$GradePoints <- FTIC_Fall_semester_all_CRS$Course_CrsGradeAwarded
FTIC_Fall_semester_all_CRS[FTIC_Fall_semester_all_CRS$GradePoints=="A", "GradePoints"] <- 4.0
FTIC_Fall_semester_all_CRS[FTIC_Fall_semester_all_CRS$GradePoints=="A-", "GradePoints"] <- 3.70
FTIC_Fall_semester_all_CRS[FTIC_Fall_semester_all_CRS$GradePoints=="B+", "GradePoints"] <- 3.30
FTIC_Fall_semester_all_CRS[FTIC_Fall_semester_all_CRS$GradePoints=="B", "GradePoints"] <- 3.00
FTIC_Fall_semester_all_CRS[FTIC_Fall_semester_all_CRS$GradePoints=="B-", "GradePoints"] <- 2.70
FTIC_Fall_semester_all_CRS[FTIC_Fall_semester_all_CRS$GradePoints=="C", "GradePoints"] <- 2.00
FTIC_Fall_semester_all_CRS[FTIC_Fall_semester_all_CRS$GradePoints=="C+", "GradePoints"] <- 2.30
FTIC_Fall_semester_all_CRS[FTIC_Fall_semester_all_CRS$GradePoints=="C-", "GradePoints"] <- 1.70
FTIC_Fall_semester_all_CRS[FTIC_Fall_semester_all_CRS$GradePoints=="D", "GradePoints"] <- 1.00
FTIC_Fall_semester_all_CRS[FTIC_Fall_semester_all_CRS$GradePoints=="D+", "GradePoints"] <- 1.30
FTIC_Fall_semester_all_CRS[FTIC_Fall_semester_all_CRS$GradePoints=="F", "GradePoints"] <- 0.00
FTIC_Fall_semester_all_CRS[FTIC_Fall_semester_all_CRS$GradePoints=="NF", "GradePoints"] <- 0.00
attach(FTIC_Fall_semester_all_CRS)
na_if(GradePoints,"P")
na_if(GradePoints,"S")
na_if(GradePoints,"XX")
na_if(GradePoints,"W")
na_if(GradePoints,"I*")
na_if(GradePoints,"I")



FTIC_Fall_semester_all_CRS[FTIC_Fall_semester_all_CRS$Course_CrsGradeAwarded=="D", 30:36]
```











plot for difference term1 and semester1

```{r}

qplot( data= FTIC_Fall_semester3, x=GPAT1 , y= Stu_GPAGradePoints)
ggplot(FTIC_Fall_semester, aes(x =GPAT1   , y = Stu_GPAGradePoints , color = Stu_GPAGradePoints)) + 
  geom_point(show.legend = T) +
  labs(x = 'First Term GPA (GPAGT/GPATH)', y='Stu_GPAGradePoints',  title = "First Term GPA vs.Stu_GPAGradePoints", subtitle = "Missing HS GPA (red)",
       color = 'Stu_GPAGradePoints') + 
  scale_color_gradient(low = "green", high = "darkgreen",
                       na.value = "red", guide = "legend") +
  theme_minimal()+theme(legend.position="bottom")
```
Find mean sd for HS GPA and TERM1 GPA
```{r}
my_norm_variables <- FTIC_Fall_semester3[, 18:26]
my_norm_variables[is.infinite(my_norm_variables$GPAT1),] <- 0
my_norm_variables[is.na(my_norm_variables$GPAT1),] <- 0
my_norm_variables[which(my_norm_variables$Stu_GPAHighSchool==0), "Stu_GPAHighSchool"] <- NA
my_norm_variables[which(is.na(my_norm_variables$Stu_GPAHighSchool)), ] 

summary(my_norm_variables)
```
GPA high school normalize
```{r}
GPAHS_normal <- my_norm_variables[!is.na(my_norm_variables$Stu_GPAHighSchool),"Stu_GPAHighSchool"] %>% 
  filter(Stu_GPAHighSchool<= 4.0)
normGPAHS_normal <- apply(GPAHS_normal, 2, function(x) (x-mean(x))/sd(x))
normGPAHS_normal[1:20]
range(normGPAHS_normal)
```
GPAT1 normalize
```{r}
GPAT1_normal <- my_norm_variables[!is.na(my_norm_variables$GPAT1),"GPAT1"]
normGPAT1 <- apply(GPAT1_normal, 2, function(x) (x-mean(x))/sd(x))
normGPAT1[1:20]
range(normGPAT1)

```



```{r}
mean_GPA_HS_FTIC <- FTIC_Fall_semester %>% group_by(Cohort) %>% summarise(meanGPaHS = mean(Stu_GPAHighSchool, na.rm = T), medianGPAHS=median(Stu_GPAHighSchool, na.rm = T), minGPAHS=min(Stu_GPAHighSchool,na.rm = T), maxGPAHS=max(Stu_GPAHighSchool, na.rm = T)); mean_GPA_HS_FTIC
```

```{r}
library(Hmisc)
Hmisc::describe(FTIC_Fall_semester)
```

```{r}
ggplot(FTIC_Fall_semester, aes(x =Stu_GPASemester  , y = Stu_GPATermHours , color = Stu_GPAHighSchool)) + 
  geom_point(show.legend = T) +
  labs(x = 'First Term GPA', y='First Term Earned Hours',  title = "First Term GPA vs. Earned Credit Hours", subtitle = "Missing HS GPA (red)",
       color = 'Stu_GPAHighSchool') + 
  scale_color_gradient(low = "green", high = "darkgreen",
                       na.value = "blue", guide = "legend") +
  theme_minimal()+theme(legend.position="bottom") #+facet_grid(Stu_CountyCode~.)
```
### missing values using KNN
```{r}
sapply(FTIC_Fall_semester, function(x){sum(is.na(x))*100/length(x)}) # 21.34889 Stu_GPAHighSchool 
```
### replace GPA HS >= 4.01 as 4 and na to zero values to replace
```{r}
GPA_HS_max <- 4
FTIC_Fall_semester$Stu_GPAHighSchool[FTIC_Fall_semester$Stu_GPAHighSchool >= 4.00] <- GPA_HS_max
FTIC_Fall_semester$Stu_GPAHighSchool[FTIC_Fall_semester$Stu_GPAHighSchool <= 2.00] <- NA

#FTIC_Fall_semester$Stu_GPAHighSchool[which(is.na(FTIC_Fall_semester$Stu_GPAHighSchool))] <- 0

ggplot(FTIC_Fall_semester[FTIC_Fall_semester$Stu_TotalUniversityHours==0 & FTIC_Fall_semester$Stu_TotalInstHours==0 & FTIC_Fall_semester$Cohort<= 2017,], aes(x = Stu_GPAUndergraduateTerm , y = Stu_GPATermHours , color = Stu_GPAHighSchool)) + 
  geom_point(show.legend = T) +
  labs(x = 'First Term GPA', y='First Term Earned Hours',  title = "First Term GPA UWF vs. First Term Earned Credit ", subtitle = "Missing HS GPA (red)",
       color = 'Stu_GPAHighSchool') + 
  scale_color_gradient(low = "green", high = "blue",
                       na.value = "gray", guide = "legend") +
  theme_minimal()+theme(legend.position="bottom") #+facet_grid(Stu_CountyCode~.)
```
### replace GPA HS
```{r}
library(mice)
library(VIM)
imp_regression <- regressionImp(Stu_GPAHighSchool~Stu_GPAUndergraduateTerm+Stu_TotalUniversityHours, FTIC_Fall_semester)
aggr(imp_regression, delimiter = "_imp")
marginplot(kNN(log10(FTIC_Fall_semester[,c("Stu_GPAHighSchool","Stu_GPAUndergraduateTerm")])), delimiter = "_imp")
set.seed(123)
imputeHS <- mice(FTIC_Fall_semester[, c("Stu_GPAHighSchool","Stu_GPAUndergraduateTerm")])
imputeHS$imp$Stu_GPAHighSchool
apply(imputeHS$imp$Stu_GPAHighSchool, 2,function(x) mean(x)) 
```
### choose highest mean
```{r}
FTIC_Fall_semesterGPAHS <- complete(imputeHS, 5);colnames(FTIC_Fall_semesterGPAHS) <- c("NewGPAHS","NewGPAFirstTerm");FTIC_Fall_semesterGPAHS
FTIC_Fall_semesterNew <- cbind(FTIC_Fall_semester,FTIC_Fall_semesterGPAHS);FTIC_Fall_semesterNew
```
```{r}
ggplot(FTIC_Fall_semesterNew[FTIC_Fall_semesterNew$Stu_TotalUniversityHours==0 & FTIC_Fall_semesterNew$Stu_TotalInstHours==0 & FTIC_Fall_semesterNew$Cohort<= 2017,], aes(x = Stu_GPAUndergraduateTerm , y = NewGPAHS , color = NewGPAHS)) + 
  geom_point(show.legend = T) +
  labs(x = 'First Term GPA', y='New GPA High School',  title = "First Term GPA UWF vs. KNN GPA HS", subtitle = "Missing HS GPA (red)",
       color = 'NewGPAHS') + 
  scale_color_gradient(low = "green", high = "blue",
                       na.value = "red", guide = "legend") +
  theme_minimal()+theme(legend.position="bottom") 
```
```{r}
library(caret)
library(RANN)
FTIC_Fall_semesterNew <- data.frame(FTIC_Fall_semesterNew)
preGPAHS <- preProcess(FTIC_Fall_semesterNew %>% 
                          dplyr::select(Stu_TotalUniversityHours, Stu_TotalInstHours, Stu_GPAUndergraduateTerm, Stu_GPAHighSchool),
                            method = c("knnImpute"),
                            k = 76,
                            knnSummary = mean)
impute_GPAHS_info <- predict(preGPAHS, FTIC_Fall_semesterNew, na.action = na.pass)
procNames <- data.frame(col = names(preGPAHS$mean), mean = preGPAHS$mean, sd = preGPAHS$std)
for(i in procNames$col){
 impute_GPAHS_info[i] <- impute_GPAHS_info[i]*preGPAHS$std[i]+preGPAHS$mean[i] 
}
```
### mean of HS GPA, imputed GSGPA, mean credit hours, withdrawn hours, sum withdrawn hours
```{r}
meanGPAHS <- impute_GPAHS_info %>% group_by(Cohort) %>% 
  summarise(meanGPA=mean(Stu_GPAHighSchool), meanGPA1=mean(NewGPAHS), meanGPAT1=mean(Stu_GPAUndergraduateTerm), meanCRT1=mean(Stu_GPATermHours, na.rm = T),meanWRate=mean((Stu_GPATermHours/Stu_CurrentTermLoad), na.rm = T),meanSumWHours=mean(sum(Stu_CurrentTermLoad-Stu_GPATermHours),na.rm = T),CountF=n());meanGPAHS
```

```{r}

gpahs <- boxplot(impute_GPAHS_info$Stu_GPAHighSchool)
gpahs$stats

```

```{r}
CSEdata[CSEdata$UNIV_ROW_ID==970577458,]
```


### ANOVA t find optimal value HS GPA for subgroups
```{r}

CSEdatascan <- CSEdatascan %>% mutate(aveGPAT1=ifelse(FIRST_FALL_GPA>=2.77,"Strong",ifelse(FIRST_FALL_GPA>=2.00,"Safe","Warning")))
CSEdatascan <- CSEdatascan %>% mutate(aveGPAHS=ifelse(GPA_HIGHSCHOOL>=3.88,"Strong",ifelse(GPA_HIGHSCHOOL>=3.33,"Safe","Warning")))


tieran <- aov(GPA_HIGHSCHOOL~APPLICANT_TIER, data =CSEdatascan )
summary(tieran)
mytukeytier <-TukeyHSD(tieran, "APPLICANT_TIER")
mytukeytier

CSEdatascan$APPLICANT_TIER <- factor(CSEdatascan$APPLICANT_TIER)
tieran1 <- aov(FIRST_FALL_GPA~APPLICANT_TIER, data =CSEdatascan )
summary(tieran1)
mytukeytier1 <-TukeyHSD(tieran1, "APPLICANT_TIER")
mytukeytier1

CSEdatascan$aveGPAT1 <- factor(CSEdatascan$aveGPAT1)
tieran2 <- aov(GPA_HIGHSCHOOL~aveGPAT1, data =CSEdatascan )
summary(tieran2)
mytukeytier2 <-TukeyHSD(tieran2, "aveGPAT1")
mytukeytier2

CSEdatascan$aveGPAT1 <- factor(CSEdatascan$aveGPAT1)
tieran3 <- aov(FIRST_FALL_GPA~aveGPAT1 + aveGPAHS, data =CSEdatascan )
summary(tieran3)
mytukeytierGPAHS <-TukeyHSD(tieran3, "aveGPAHS")
mytukeytierGPAT1 <- TukeyHSD(tieran3, "aveGPAT1")
mytukeytierGPAHS
mytukeytierGPAT1
```


### replace GPAHS>=4.01 as 4.00
```{r}
CSEdatascan <- CSEdata
CSEdatascan[which(is.na(CSEdatascan$FIRST_FALL_GPA)),] <- 0
CSEdatascan[which(is.na(CSEdatascan$GPA_ENTERING_SECOND_FALL)),] <- 0
CSEdatascan$GPA_HIGHSCHOOL[CSEdatascan$GPA_HIGHSCHOOL >=  4.01] <- 4


CSEdatascan1 <- CSEdatascan %>% group_by(APPLICANT_TIER,COHORT_YEAR) %>% 
  summarise(meanGPAHS=mean(GPA_HIGHSCHOOL, na.rm = T), meanGPAY1=mean(GPA_ENTERING_SECOND_FALL), meanGPAT1=mean(FIRST_FALL_GPA), CountF=n());CSEdatascan1
gpahs1 <- boxplot(CSEdatascan$GPA_HIGHSCHOOL)
gpahs1$stats

```
### ANOVA for HS GPA by TIer
```{r}
CSEdatascan$APPLICANT_TIER <- factor(CSEdatascan$APPLICANT_TIER)
CSEdatascan <- CSEdatascan %>% mutate(aveGPAT1=ifelse(FIRST_FALL_GPA>=2.77,"Strong",ifelse(FIRST_FALL_GPA>=2.00,"Safe","Warning")))
CSEdatascan <- CSEdatascan %>% mutate(aveGPAHS=ifelse(GPA_HIGHSCHOOL>=3.88,"Strong",ifelse(GPA_HIGHSCHOOL>=3.33,"Safe","Warning")))


tieran <- aov(GPA_HIGHSCHOOL~APPLICANT_TIER, data =CSEdatascan )
summary(tieran)
mytukeytier <-TukeyHSD(tieran, "APPLICANT_TIER")
mytukeytier

CSEdatascan$APPLICANT_TIER <- factor(CSEdatascan$APPLICANT_TIER)
tieran1 <- aov(FIRST_FALL_GPA~APPLICANT_TIER, data =CSEdatascan )
summary(tieran1)
mytukeytier1 <-TukeyHSD(tieran1, "APPLICANT_TIER")
mytukeytier1

CSEdatascan$aveGPAT1 <- factor(CSEdatascan$aveGPAT1)
tieran2 <- aov(GPA_HIGHSCHOOL~aveGPAT1, data =CSEdatascan )
summary(tieran2)
mytukeytier2 <-TukeyHSD(tieran2, "aveGPAT1")
mytukeytier2

CSEdatascan$aveGPAT1 <- factor(CSEdatascan$aveGPAT1)
tieran3 <- aov(FIRST_FALL_GPA~aveGPAT1 + aveGPAHS, data =CSEdatascan )
summary(tieran3)
mytukeytierGPAHS <-TukeyHSD(tieran3, "aveGPAHS")
mytukeytierGPAT1 <- TukeyHSD(tieran3, "aveGPAT1")
mytukeytierGPAHS
mytukeytierGPAT1


```




### CSE data

```{r}
CSEdata <- read.csv("~/Data_IR/upCSEupto2020.csv")# %>% filter(COHORT_YEAR<=20192020)
unique(CSEdata$COHORT_YEAR)
CSEdata[which(CSEdata$ACT_PROPORTION==0),"ACT_PROPORTION"] <- NA
CSEdata[which(CSEdata$SAT_PROPORTION==0),"SAT_PROPORTION"] <- NA
CSEdata[which(CSEdata$GPA_HIGHSCHOOL==0),"GPA_HIGHSCHOOL"] <- NA
CSEdata[which(CSEdata$AP_CREDITS==0),"AP_CREDITS"] <- NA
CSEdata[which(CSEdata$HOURS_BROUGHT_TO_UNIVERSITY==0),"HOURS_BROUGHT_TO_UNIVERSITY"] <- NA


## GPA by cohort
meanGPAHSFTIC <- CSEdata %>% group_by(COHORT_YEAR) %>% 
  summarise(meanGPAHS=mean(GPA_HIGHSCHOOL, na.rm = T),meanPH=mean(HOURS_BROUGHT_TO_UNIVERSITY,na.rm = T),meanAP=mean(AP_CREDITS , na.rm = T),  CountF=n());meanGPAHSFTIC

### GPA by TIER
meanGPAHS1 <- CSEdata %>% group_by(COHORT_YEAR,APPLICANT_TIER) %>% 
  summarise(meanGPAHS=mean(GPA_HIGHSCHOOL, na.rm = T), meanACT=mean(ACT_PROPORTION, na.rm = T), meanSAT=mean(SAT_PROPORTION, na.rm = T), CountF=n());meanGPAHS1
gpahs1 <- boxplot(CSEdata$GPA_HIGHSCHOOL)
gpahs1$stats

## HS name
meanGPAHSNAME <- CSEdata %>% group_by(HIGH_SCHOOL_NAME) %>% 
  summarise(meanGPAHS=mean(GPA_HIGHSCHOOL, na.rm = T),meanGPAT1=mean(FIRST_FALL_GPA, na.rm=T),meanPH=mean(HOURS_BROUGHT_TO_UNIVERSITY,na.rm = T),meanAP=mean(AP_CREDITS , na.rm = T),  CountF=n())
meanGPAHSNAME1 <- meanGPAHSNAME %>%  arrange(-CountF);meanGPAHSNAME1

```
```{r}
lmHS <- lm(GPA_HIGHSCHOOL~ACT_PROPORTION, data=CSEdata)
summary(lmHS)
CSEdata$APPLICANT_TIER <- as.factor(CSEdata$APPLICANT_TIER)
pHS <- ggplot(data=CSEdata, aes(x=SAT_PROPORTION,y=ACT_PROPORTION)) +geom_point(col="blue")
pHS + geom_smooth(col="red")
```
```{r}
ggplot(CSEdata[CSEdata$HOURS_BROUGHT_TO_UNIVERSITY <= 45 & CSEdata$COHORT_YEAR <= 20192020,], aes(x = GPA_HIGHSCHOOL , y = FIRST_FALL_GPA , color = ACT_PROPORTION) )+ 
  geom_point(show.legend = T) +
  labs(x = 'GAP HS', y='GPA Term1',  title = "Preparation vs. Persistence ", subtitle = "Missing value (yellow)",
       color = 'ACT_PROPORTION') + geom_smooth(col="red")+
  scale_color_gradient(low = "green", high = "blue",
                       na.value = "yellow", guide = "legend") +
  theme_minimal()+theme(legend.position="bottom") 
```



### geting Tier and HS GPA
```{r}
CSEdata <- read.csv("~/Data_IR/CSEdata.csv") %>% filter(COHORT_YEAR<=20192020)
CSEdata$COHORT_YEAR1 <- stringr::str_extract( CSEdata$COHORT_YEAR, "^.{4}")

## merge with cse and enr data set
FTIC_Fall_CSEinfo <- merge(FTIC_Fall_semester,CSEdata, by.x = c("Cohort","Stu_Ethnicity","Stu_Gender","Stu_NationCode","Stu_County","Stu_Department","Stu_TotalUniversityHours"),
                           by.y = c("COHORT_YEAR1","ETHNICITY","GENDER","NATION_GROUP","COUNTY","ENTRY_DEPARTMENT","HOURS_BROUGHT_TO_UNIVERSITY")) 

FTIC_ALL_semester_CSE_info <- merge(FTIC_Fall_CSEinfo, FTIC_Fall_semester) # only half of them captured 2636/5827
```


### Coding for factors
```{r}
library(dplyr)
recodingdemographicFACTORS <- MASTER_CRSCSE_allcourse_tier %>% 
  #filter(ENTRY_COLLEGE=="CASSH" | ENTRY_COLLEGE=="HMCSE"|ENTRY_COLLEGE=="UKCOH") %>% 
  mutate(TIER_THREE = ifelse(APPLICANT_TIER>=4,3,ifelse(APPLICANT_TIER==3,2,1))) %>% 
  mutate(ETHNICITY_recode1 = ifelse(ETHNICITY=="White","White",ifelse(ETHNICITY=="African American",              "AA", ifelse(ETHNICITY=="Asian","Asian", ifelse(ETHNICITY=="Hispanic","HSP",                     ifelse(ETHNICITY=="Two or More","TM","OTHERS")))))) %>% 
  mutate(ETHNICITY_recode2 = ifelse(ETHNICITY=="White","White","Non_White")) %>% 
  mutate(ETHNICITY_recode3 = ifelse(ETHNICITY=="White","White",ifelse(ETHNICITY=="African American"|
         ETHNICITY=="Hispanic","AAH","OTHERS"))) %>% 
  filter(str_detect(CRSNAME1, "MAC") | str_detect(CRSNAME2,"MAC")|str_detect(CRSNAME3,"MAC")|str_detect(CRSNAME4,"MAC")|str_detect(CRSNAME5,"MAC")|str_detect(CRSNAME6,"MAC")) # from 3225,3216 so choose CRSNAME6




do.call(rbind,lapply(recodingdemographicFACTORS[29],tblFun))

Majorrecoding <- recodingdemographicFACTORS %>% 
  mutate(ENTRY_DEPARTMENT.recode1=ifelse(ENTRY_COLLEGE=="HMCSE"&ENTRY_DEPARTMENT=="Biology","Biology","Others"))
unique(Majorrecoding$ENTRY_DEPARTMENT.recode1)
```

```{r}
qplot(data = recodingdemographicFACTORS, x=recodingdemographicFACTORS$GPA_HIGHSCHOOL, y= recodingdemographicFACTORS$FIRST_FALL_BRIGHT_FUTURES_AMOUNT, color=recodingdemographicFACTORS$TIER_THREE, facets = APR ~ETHNICITY_recode1)

qplot(data = recodingdemographicFACTORS,   x=recodingdemographicFACTORS$GPA_HIGHSCHOOL, y= recodingdemographicFACTORS$FIRST_FALL_PELL_AMOUNT, color=APR, facets = GENDER ~ ENTRY_COLLEGE)

qplot(data = recodingdemographicFACTORS, x=recodingdemographicFACTORS$TIER_THREE, y= recodingdemographicFACTORS$FIRST_FALL_BRIGHT_FUTURES_AMOUNT, color=recodingdemographicFACTORS$ETHNICITY_recode2, facets = APR ~ETHNICITY_recode2)

qplot(data = recodingdemographicFACTORS,   x=FIRST_FALL_GPA, y= FIRST_FALL_BRIGHT_FUTURES_AMOUNT, color=ENTRY_COLLEGE)

```
```{r}
glimpse(recodingdemographicFACTORS)
```

# manage the data for three college
```{r}
  whichcollege <- recodingdemographicFACTORS %>% 
  filter(COHORT_YEAR < 20202021) %>% 
  filter(COUNTY_GROUP=="Tri-County") %>% 
  select(CRSNAME1, HOURS_BROUGHT_TO_UNIVERSITY,AP_CREDITS,FIRST_FALL_NEED_BASED_LOANS_AMOUNT,GENDER,COUNTY,GPA_HIGHSCHOOL,ETHNICITY_recode3,FIRST_FALL_GPA,ENTRY_COLLEGE) %>% 
  na.omit()

with(whichcollege, table(ENTRY_COLLEGE, COUNTY))
with(whichcollege, table(ENTRY_COLLEGE, ETHNICITY_recode3))
with(whichcollege, do.call(rbind, tapply(GPA_HIGHSCHOOL, COUNTY, function(x) c(M=mean(x), SD=sd(x)))))
with(whichcollege, do.call(rbind, tapply(FIRST_FALL_GPA, COUNTY, function(x) c(M=mean(x), SD=sd(x)))))
glimpse(whichcollege)



```
```{r}
library(foreign)
library(nnet)
library(ggplot2)
library(reshape2)
library(caret)
# the outcome should be the second year college
whichcollege$ENTRY_COLLEGE <- factor(whichcollege$ENTRY_COLLEGE)
whichcollege$ENTRY_COLLEGE_out <- relevel(whichcollege$ENTRY_COLLEGE, ref="UKCOH")
whichcollege1 <- whichcollege[,-10]
set.seed(123)
model_first <- multinom(ENTRY_COLLEGE_out~., data = whichcollege1)
summary(model_first)
z <- summary(model_first)$coefficients/summary(model_first)$standard.errors
z

```
```{r}
# 2-tailed z test
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
```


```{r}
## extract the coefficients from the model and exponentiate
exp(coef(model_first))
```
```{r}
confMatRank <- table(predict(model_first), whichcollege$ENTRY_COLLEGE)
confMatRank
sum(diag(confMatRank)/sum(confMatRank))
1-sum(diag(confMatRank)/sum(confMatRank)) 

p <- predict(model_first, whichcollege)
pred <- predict(model_first, whichcollege[c(1:10),], type="prob")
round(pred, digits = 5)

n <- table(whichcollege$ENTRY_COLLEGE)
n
n/sum(n)

p <- predict(model_first, whichcollege)
tab <- table(p, whichcollege$ENTRY_COLLEGE)
tab/colSums(tab)
```
```{r}
head(p <- fitted(model_first),10)
head(whichcollege[,10],10)
```
### multinomial naive bayes classification
```{r}

  whichcollege <- recodingdemographicFACTORS %>% 
    filter(COHORT_YEAR < 20202021) %>% 
  filter(COUNTY_GROUP=="Tri-County") %>% 
  select(CRSNAME1,CRSNAME2,CRSNAME3,CRSNAME1,CRSNAME4,
         CRSgrade1,CRSgrade2,CRSgrade3,CRSgrade4,
         ENTRY_COLLEGE) %>% 
  na.omit()
whichcollege$CRSNAME1 <- factor(whichcollege$CRSNAME1)
whichcollege$CRSNAME2 <- factor(whichcollege$CRSNAME2)
whichcollege$CRSNAME3 <- factor(whichcollege$CRSNAME3)
whichcollege$CRSNAME4 <- factor(whichcollege$CRSNAME4)
whichcollege$CRSgrade1 <- factor(whichcollege$CRSgrade1)
whichcollege$CRSgrade2 <- factor(whichcollege$CRSgrade2)
whichcollege$CRSgrade3 <- factor(whichcollege$CRSgrade3)
whichcollege$CRSgrade4 <- factor(whichcollege$CRSgrade4)
#whichcollege$COUNTY <- factor(whichcollege$COUNTY)
#whichcollege$ETHNICITY_recode3 <- factor(whichcollege$ETHNICITY_recode3)
whichcollege$ENTRY_COLLEGE <- factor(whichcollege$ENTRY_COLLEGE)
apply(whichcollege, 2, get.missing.prop)
whichcollege <- whichcollege[-33,]
glimpse(whichcollege)
```
### Data partition and modeling
```{r}
TrainingIndex <- createDataPartition(whichcollege$ENTRY_COLLEGE, p=0.7, list=FALSE)
whichcollegeTraingSet <-  whichcollege[TrainingIndex, ]
whichcollegeTestingSet <- whichcollege[-TrainingIndex, ]
table(whichcollegeTraingSet$ENTRY_COLLEGE) %>% prop.table()
table(whichcollegeTestingSet$ENTRY_COLLEGE) %>% prop.table()
library(naivebayes)
model_naive <- naive_bayes(ENTRY_COLLEGE~., data=whichcollegeTraingSet)
confusionMatrix(predict(model_naive), whichcollegeTraingSet$ENTRY_COLLEGE)
p <- round(predict(model_naive, whichcollegeTraingSet, type="prob"), digits = 5)
model_naivet <- naive_bayes(ENTRY_COLLEGE~., data=whichcollegeTestingSet)
confusionMatrix(predict(model_naivet), whichcollegeTestingSet$ENTRY_COLLEGE)
pt <- round(predict(model_naivet, whichcollegeTestingSet, type="prob"), digit=5)
whichcollege_p <- cbind(whichcollegeTraingSet,p)
whichcollege_pt <- cbind(whichcollegeTestingSet,pt)
whichcollege_results <- rbind(whichcollege_p,whichcollege_pt)
write.csv(whichcollege_results,"V1_whichcollege_results.csv")

#export the resutls


```

