---
title: "Admitted 2021 FTIC - Predicting APR"
author: "PAM Lab -  UWF"
date: "02-26-2021"
output: 
  html_document:
    fig_height: 3
    fig_width: 5
  pdf_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

```{r, setup, include=FALSE}
require(mosaic)   # Load additional packages here 
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code


#final_admission <- readRDS("./final_admissionFTIC_model.rds")
library(htmlTable)
library(dplyr)
library(readr)
all_vari_PREDICTION <- read_csv("v5all_vari_PREDICTION.csv") %>% # clean the final dataset and complete this report
        mutate(College=ifelse(CURRICULUM_COLL=="UNA","UNA", # college
                                                               ifelse(CURRICULUM_COLL=="A","HMCSE",
                                                                ifelse(CURRICULUM_COLL=="P","CEPS",
                                                                 ifelse(CURRICULUM_COLL=="H","CASSH",
                                                                    ifelse(CURRICULUM_COLL=="M","UKCOH",
                                                                      ifelse(CURRICULUM_COLL=="B","COB","UNA")))))))
```

```{r eval=FALSE, include=FALSE}
library(readxl)
V1_updated_tier_pred_incoming_class_02_22_2021 <- read_excel("G:/Shared drives/HMCSE-PAM Lab/Jay's Space/DATA/V1-updated_tier_pred_incoming_class_02_22_2021.xlsx", 
    sheet = "updated_tier_pred_incoming_clas")
V2_updated_tier_pred_incoming_class_02_22_2021 <- V1_updated_tier_pred_incoming_class_02_22_2021[!is.na(V1_updated_tier_pred_incoming_class_02_22_2021$HS_OFFER_GPA),] %>% 
  filter(HS_OFFER_GPA != 0) %>% 
  mutate(codeHSGPA=ifelse(HS_OFFER_GPA<3.1000,"GPA<3.10",
                          ifelse(HS_OFFER_GPA<3.20,"GPA<3.20",
                                 ifelse(HS_OFFER_GPA<3.30,"GPA<3.30",
                                        ifelse(HS_OFFER_GPA<3.40,"GPA<3.40",
                                               ifelse(HS_OFFER_GPA<3.50,"GPA<3.50",
                                                      ifelse(HS_OFFER_GPA<3.60,"GPA<3.60",
                                                             ifelse(HS_OFFER_GPA<3.70,"GPA<3.70",
                                                                    ifelse(HS_OFFER_GPA<3.80,"GPA<3.80",
                                                                           ifelse(HS_OFFER_GPA<3.90,"GPA<3.90",
                                                                                  ifelse(HS_OFFER_GPA<4.00,"GPA<4.00","GPA>=4.00"))))))))))) 

V2_updated_tier_pred_incoming_class_02_22_2021[V2_updated_tier_pred_incoming_class_02_22_2021$PROGRAM_CODE=="U-EGT-BS-03","PROGRAM_CODE"] <- "U-EGT-BS-01"
V2_updated_tier_pred_incoming_class_02_22_2021[V2_updated_tier_pred_incoming_class_02_22_2021$PROGRAM_CODE=="U-CON-BS-00","PROGRAM_CODE"] <- "U-EGT-BS-01"

V2_updated_tier_pred_incoming_class_02_22_2021[V2_updated_tier_pred_incoming_class_02_22_2021$PROGRAM_CODE=="U-HOS-BS-09","PROGRAM_CODE"] <- "U-HOS-BS-00"
V2_updated_tier_pred_incoming_class_02_22_2021[V2_updated_tier_pred_incoming_class_02_22_2021$PROGRAM_CODE=="U-MUP-BA-00","PROGRAM_CODE"] <- "U-MUP-BM-00"
#3,466
xtabs(~all_vari_PREDICTION$Pred_Tier)
xtabs(~V1_updated_tier_pred_incoming_class_02_22_2021$Pred_Tier)
glimpse(all_vari_PREDICTION)L

# merge with APR hs code and program
library(readr)
V1_HIST_APR_PROGRAM_CODE <- read_csv("G:/My Drive/#1_Jay's Space/Application Data/V1-HIST_APR_PROGRAM_CODE.csv", 
    col_types = cols(X1 = col_skip()))

library(readr)
HIST_APR_HSNAME_withCODE <- read_csv("G:/My Drive/#1_Jay's Space/Application Data/HIST_APR_HSNAME_withCODE.csv", 
    col_types = cols(X1 = col_skip()))

# merge with program code
V3_updated_tier_pred_incoming_class_02_22_2021 <- merge(V2_updated_tier_pred_incoming_class_02_22_2021, V1_HIST_APR_PROGRAM_CODE, by.x="PROGRAM_CODE" , by.y="ENTRY_PROGRAM_CODE",
                                                        all.x = T)


colSums(is.na(V3_updated_tier_pred_incoming_class_02_22_2021))
NA_PROGRAM <- V3_updated_tier_pred_incoming_class_02_22_2021[which(is.na(V3_updated_tier_pred_incoming_class_02_22_2021$APR_PROGRAM)),]
NA_PROGRAM

# merge with HS code
V4_updated_tier_pred_incoming_class_02_22_2021 <- merge(V3_updated_tier_pred_incoming_class_02_22_2021, HIST_APR_HSNAME_withCODE, by.x="HS_CEEB" , by.y="Code",
                                                        all.x = T)
colSums(is.na(V4_updated_tier_pred_incoming_class_02_22_2021))

V4_updated_tier_pred_incoming_class_02_22_2021[which(is.na(V4_updated_tier_pred_incoming_class_02_22_2021$APR_HSNAME)), "APR_HSNAME"] <- "0"
V4_updated_tier_pred_incoming_class_02_22_2021$APR_HSNAME <- as.numeric(V4_updated_tier_pred_incoming_class_02_22_2021$APR_HSNAME)
V4_updated_tier_pred_incoming_class_02_22_2021 <- V4_updated_tier_pred_incoming_class_02_22_2021 %>% mutate(APR_HSNAME1=ifelse(APR_HSNAME<=0, 0.7309942, V4_updated_tier_pred_incoming_class_02_22_2021$APR_HSNAME))


colSums(is.na(V4_updated_tier_pred_incoming_class_02_22_2021)) #517 na
valueHScode999999 <- V4_updated_tier_pred_incoming_class_02_22_2021[V4_updated_tier_pred_incoming_class_02_22_2021$HS_CEEB==999999,] #NOTAVAILABLE/FL == 0.7309942
valueHScode999999

write.csv(V4_updated_tier_pred_incoming_class_02_22_2021, "applicants2021_3466DF.csv")

library(readxl)
FTIC_202108_202105_Admitted_as_of_2_22_21 <- read_excel("G:/Shared drives/HMCSE-PAM Lab/DATA 202101/Admissions Data 2018-2020/FTIC 202108 202105 Admitted as of 2-22-21.xlsx")
hist(FTIC_202108_202105_Admitted_as_of_2_22_21$TRANSFER_HOURS_EARNED)
colSums(is.na(FTIC_202108_202105_Admitted_as_of_2_22_21))
xtabs(~FTIC_202108_202105_Admitted_as_of_2_22_21$TRANSFER_HOURS_EARNED)

library(ggplot2)
ggplot(FTIC_202108_202105_Admitted_as_of_2_22_21, aes(GENDER)) + geom_bar(aes(fill=factor(GENDER)))
FTIC_202108_202105_CNTYTOP <- FTIC_202108_202105_Admitted_as_of_2_22_21 %>% group_by(CNTY) %>% dplyr::summarise(Count=n()) %>% arrange(-Count)
FTIC_202108_202105_CNTYTOP

ggplot(FTIC_202108_202105_Admitted_as_of_2_22_21, aes(HS_OFFER_GPA)) + geom_histogram(bins=20, alpha=0.5, fill="blue")
ggplot(FTIC_202108_202105_Admitted_as_of_2_22_21, aes(HS_FL_YRS)) + geom_histogram(bins=20, alpha=0.5, fill="blue")
ggplot(FTIC_202108_202105_Admitted_as_of_2_22_21, aes(HS_MAT_YRS)) + geom_histogram(bins=20, alpha=0.5, fill="blue")

# GPA by HS name

p1 <- ggplot(FTIC_202108_202105_Admitted_as_of_2_22_21, aes(CURRICULUM_DEPT, HS_OFFER_GPA))
p1 <- p1 + geom_boxplot(aes(group=CURRICULUM_DEPT,fill=factor(CURRICULUM_DEPT), alpha=0.4))
p2 <- p1 + scale_y_continuous(breaks = seq(min(2.5), max(5.2), by=0.1)) + theme_bw()
p2

# courses works
FTIC_202108_202105_Admitted_as_of_2_22_21$HS_MAT_YRS_index <- ifelse(FTIC_202108_202105_Admitted_as_of_2_22_21$HS_MAT_YRS<=4.0, "MATH<=4.0YRS","MATH>4.0YRS")
ggplot(FTIC_202108_202105_Admitted_as_of_2_22_21, aes(HS_OFFER_GPA)) + geom_histogram(aes(fill=HS_MAT_YRS_index),color="blue",binwidth=1) + theme_bw()



```

```{r eval=FALSE, include=FALSE}
library(dplyr)
library(read.dbc)
V2_APRMASTERDATASET <- read.csv("G:/My Drive/#1_Jay's Space/Application Data/V2_APRMASTERDATASET.csv", stringsAsFactors=TRUE) %>% 
  mutate(BESTTEST = pmax(ACT_PROPORTION, SAT_PROPORTION, na.rm = TRUE))
library(readr)
Top5OverallDWF_Program <- read_csv("G:/My Drive/#1_Jay's Space/Application Data/Top5OverallDWF_Program_imputed2.csv")

V2_APRMASTERDATASET <- merge(V2_APRMASTERDATASET, Top5OverallDWF_Program, by="ENTRY_PROGRAM", all.x = T)

colSums(is.na(V2_APRMASTERDATASET)) # 4 nas
aa <- V2_APRMASTERDATASET[which(is.na(V2_APRMASTERDATASET$Top5OverallDWF)),]
aa <- V2_APRMASTERDATASET[V2_APRMASTERDATASET$Deg_DepartmentCode=="PR",]
max(V2_APRMASTERDATASET$Top5OverallDWF, na.rm=T)
V2_APRMASTERDATASET[is.na(V2_APRMASTERDATASET$Top5OverallDWF), "Top5OverallDWF"] <- 0.2616279 # imputed program children & society DWF
V2_APRMASTERDATASET1 <- V2_APRMASTERDATASET %>%  mutate(codeHSGPA=ifelse(GPA_HIGHSCHOOL <3.1000,"GPA<3.10",
                          ifelse(GPA_HIGHSCHOOL <3.20,"GPA<3.20",
                                 ifelse(GPA_HIGHSCHOOL <3.30,"GPA<3.30",
                                        ifelse(GPA_HIGHSCHOOL <3.40,"GPA<3.40",
                                               ifelse(GPA_HIGHSCHOOL <3.50,"GPA<3.50",
                                                      ifelse(GPA_HIGHSCHOOL <3.60,"GPA<3.60",
                                                             ifelse(GPA_HIGHSCHOOL <3.70,"GPA<3.70",
                                                                    ifelse(GPA_HIGHSCHOOL <3.80,"GPA<3.80",
                                                                           ifelse(GPA_HIGHSCHOOL <3.90,"GPA<3.90",
                                                                                  ifelse(GPA_HIGHSCHOOL <4.00,"GPA<4.00","GPA>=4.00")))))))))))


write.csv(V2_APRMASTERDATASET1, "V3_APRMASTERDATASET.csv")

######################################################################
### need to combine with APR by the most failed course by program ####
######################################################################
```


```{r include=FALSE}
library(readr)
V3_APRMASTERDATASET <- read.csv("G:/My Drive/#1_Jay's Space/Application Data/V3_APRMASTERDATASET.csv", stringsAsFactors=TRUE)

# math course
all_1819_mathyrs <- read.csv("G:/My Drive/#1_Jay's Space/Application Data/all_1819_mathyrs.csv") %>% select(UNIV_ROW_ID, contains("yrs"), contains("APP"))
V4_APR_MASTER_mathyrs <- merge(V3_APRMASTERDATASET,all_1819_mathyrs, by="UNIV_ROW_ID", all.x=T )

V4_APR_MASTER_mathyrs$HS_MAT_YRS_index <- ifelse(V4_APR_MASTER_mathyrs$HS_MAT_YRS<=4.0, "MATH<=4.0YRS","MATH>4.0YRS")
V4_APR_MASTER_mathyrs$HS_MAT_YRS_index <- as.factor(V4_APR_MASTER_mathyrs$HS_MAT_YRS_index)

V3_APR_MASTER <- V4_APR_MASTER_mathyrs %>% mutate(Top5PassRate=(1-Top5OverallDWF)) %>% filter(COHORT_YEAR>=20182019 & COHORT_YEAR<=20192020) %>%  mutate(CourseWork=(HS_ENG_YRS+HS_MAT_YRS+HS_NS_YRS+HS_SS_YRS)/4) 
colSums(is.na(V3_APR_MASTER))

NA_95 <- V3_APR_MASTER[which(is.na(V3_APR_MASTER$CourseWork)),]
str(V3_APR_MASTER)
# imputed NA
library(randomForest)
imputed_df <- rfImpute(APR~ GPA_HIGHSCHOOL+APR_HSNAME+HS_ENG_YRS+HS_MAT_YRS+HS_SS_YRS+HS_NS_YRS+HS_FL_YRS+CourseWork, data = V3_APR_MASTER, iter=3)
# check multicollinearity
library(PerformanceAnalytics)
corcols <- V3_APR_MASTER %>% select(11:17)
chart.Correlation(imputed_df[,-1]) # suggested using CourseWork or math and ENG


xtabs(~V3_APR_MASTER$APR+V3_APR_MASTER$PELL_VERF)
V3_APR_MASTER$HS_ENG_YRS <- imputed_df$HS_ENG_YRS
V3_APR_MASTER$HS_MAT_YRS <- imputed_df$HS_MAT_YRS
V3_APR_MASTER$CourseWork <- imputed_df$CourseWork
V3_APR_MASTER$HS_MAT_YRS_index <-  ifelse(V3_APR_MASTER$HS_MAT_YRS<=4.0, "MATH<=4.0YRS","MATH>4.0YRS")
V3_APR_MASTER$HS_ENG_YRS_index <-  ifelse(V3_APR_MASTER$HS_ENG_YRS<=4.0, "ENG<=4.0YRS","ENG>4.0YRS")

colSums(is.na(V3_APR_MASTER))

V3_APR_MASTER_vari <- V3_APR_MASTER %>% select(APR,APP_NO,APR_PROGRAM,APR_HSNAME,GENDER,codeHSGPA,Top5PassRate,ETHNICITY,HOURS_BROUGHT_TO_UNIVERSITY,
                                               CourseWork,HS_ENG_YRS, HS_MAT_YRS,PELL_VERF,HS_MAT_YRS_index,HS_ENG_YRS_index,APP_NO) %>% na.omit()
xtabs(~V3_APR_MASTER_vari$APP_NO+ V3_APR_MASTER_vari$APR)
#write.csv(V3_APR_MASTER_vari, "APR_FTIC201819_DF.csv")
```

```{r include=FALSE}
library(readr)
Final_V3_APR_MASTER <- read.csv("Final_V3_APR_MASTER.csv",  stringsAsFactors=TRUE)
library(caret)
# Performs stratified random split of the data set
set.seed(123)
TrainingIndex <- createDataPartition(Final_V3_APR_MASTER$APR, p=0.7, list = FALSE)
TrainingSet <- Final_V3_APR_MASTER[TrainingIndex,] # Training Set
TestingSet <- Final_V3_APR_MASTER[-TrainingIndex,] # Test Set
# Build model using all factors and labels
set.seed(111)
admitted_glm<- glm(APR ~ ., data = TrainingSet, family="binomial")
summary(admitted_glm)
exp(coef(admitted_glm))
p_0 <- predict(admitted_glm, TrainingSet, type = 'response')
#misclassfication
tab_0 <- table(TrainingSet$APR, p_0>0.5)
(sum(diag(tab_0))/sum(tab_0))

#reduced model
set.seed(222)
admitted_glm_re <- glm(APR ~APR_PROGRAM+APR_HSNAME+GENDER+codeHSGPA+CourseWork+Top5PassRate*codeHSGPA+ETHNICITY+HS_MAT_YRS_index +HS_ENG_YRS_index, data = TrainingSet, family="binomial")
summary(admitted_glm_re)
#anova(admitted_glm, admitted_glm_re, test = "Chisq")
library(tidyverse)  # data manipulation and visualization
library(modelr)     # provides easy pipeline modeling functions
library(broom)      # helps to tidy up model outputs
P_value <- tidy(admitted_glm_re)
round(P_value$p.value, digits = 6)
#coefficient
coef1 <- exp(coef(admitted_glm_re))
coef_table <- knitr::kable(coef1)
coef_table
#confint(HSGPAunderlgm_re)
vari_imp <- knitr::kable(caret::varImp(admitted_glm_re))
vari_imp

#accuracy 
p_1 <- predict(admitted_glm_re, TrainingSet, type="response")
#P_a <- predict(admitted_glm_re, TrainingSet)
#confusionMatrix(P_a, as.factor(TrainingSet$APR))

#head(p_1,10)
#head(TrainingSet,10)
pred_1 <- ifelse(p_1>0.5, 1,0)
tab_1 <- table(Predicted=pred_1, Actural=TrainingSet$APR)
tab_1
round(sum(diag(tab_1))/sum(tab_1),4) #0.8275
#test data
p_2 <- predict(admitted_glm_re, TestingSet, type="response")
#head(p_2,10)
#head(TestingSet,10)
pred_2 <- ifelse(p_2>0.5, 1,0)
tab_2 <- table(Predicted=pred_2, Actural=TestingSet$APR)
tab_2
round(sum(diag(tab_2))/sum(tab_2),4) #0.8288
### goodness of fit
with(admitted_glm_re, pchisq(null.deviance - deviance, df.null-df.residual, lower.tail=F))#p-value==1.86439e-162

```









```{r eval=FALSE, include=FALSE}
## read datset
library(readr)
library(dplyr)
applicants2021_3466DF <- read.csv("G:/My Drive/#1_Jay's Space/AdmittedFTIC2021/applicants2021_3466DF.csv", stringsAsFactors=TRUE) 
xtabs(~applicants2021_3466DF$PELL_ELGBL +  applicants2021_3466DF$PELL_VERF)
all2021admittedFTIC_all <- applicants2021_3466DF %>% select(APR_PROGRAM, APR_HSNAME="APR_HSNAME1", GENDER, codeHSGPA, PELL_VERF, ETHNICITY,HOURS_BROUGHT_TO_UNIVERSITY)
hist(all2021admittedFTIC_all$HOURS_BROUGHT_TO_UNIVERSITY)
PREDICTION <- round(predict(admitted_glm_re, all2021admittedFTIC_all, type="response"),4)
hist(PREDICTION)

PREDICTION_DF <- cbind(applicants2021_3466DF, PREDICTION)
PREDICTION_DF$codeProb_APR <- ifelse(PREDICTION_DF$PREDICTION>=0.90,"low-risk",
                                     ifelse(PREDICTION_DF$PREDICTION>=0.60,"moderate-risk","high-risk"))




#write.csv(PREDICTION_DF,"v4all_vari_PREDICTION.csv")
write.csv(PREDICTION_DF,"v5all_vari_PREDICTION.csv")

xtabs(~PREDICTION_DF$codeProb_APR)
xtabs(~all_PREDICTION$TIER)
# save model
saveRDS(admitted_glm_re, "./final_admissionFTIC_model.rds")
### NO RUN ###
#load model
final_admission <- readRDS("./final_admissionFTIC_model.rds")
print(final_admission)
#prediction
prediction_admissions <- predict(final_admission, newadmitteddatahere)
```



## Summary

Total 3,558 FTIC admitted and it's necessary to predict who will meet APR or non-APR.
There were 3,466 applicants after removing missing values for HS GPA (92) were predicted.

* Factors considered:
    + APR by program
    + APR by high school name
    + GENDER (female / male)
    + HS GPA by tenths
    + Prior Hours
    + Ethnicity: white / non-White / Non-Resident Alien
    + Pell Verification: Yes / No
    
   
    
* Data for prediction modeling from:
    + FTIC 2017 to 2019
    + Total of 3,178 data point (removed NA)
    + Data partitioned 70% (train) and 30% (test)
    + Overall accuracy is 0.82
    
* Issues with new 2021 admission DATA:
    + If there is no HS name from the historical data, it is listed as "NOTAVAILABLE/FL"
    + Replacement values are based on APR by "not-available/non-FL" and by HS GPA 
    + If APR by HS name is 0%, it is replaced with the minimum APR by HS name, but no greater than APR by not-available/non-FL (0.74)
    + It was observed that some high HS GPA (>3.60) are high-risk as their APR by HS name was historically between 0 and 20%
    + For example, 2 FTIC came from Fort Myers HS and in the last five years all dropped out
    
    

## Coefficients (reference  is HSGPA < 3.10, Gender=Fenale)




## Cross-Validation

* Used Methods
  + Leave one out cross-validation (LOOCV)
  + K-fold cross-validation (cv)
  + Bootstrapping (boot)
  + Repeated K-fold cross-validation (repeatedcv)
  

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(Boruta)
library(mlbench)  
library(randomForest)
library(caret)

# using LOOCV RESAMPLING
model.loocv <- train(APR ~APR_PROGRAM+APR_HSNAME+GENDER+codeHSGPA+HOURS_BROUGHT_TO_UNIVERSITY*codeHSGPA+PELL_VERF*GENDER+ETHNICITY, data=TrainingSet, method="glm",family="binomial",trControl=trainControl(method = "LOOCV", number=5,classProbs = TRUE))

# using K-folder corss validation resampling
model.cv <- train(APR ~APR_PROGRAM+APR_HSNAME+GENDER+codeHSGPA+HOURS_BROUGHT_TO_UNIVERSITY*codeHSGPA+PELL_VERF*GENDER+ETHNICITY, data=TrainingSet, method="glm",family="binomial",trControl=trainControl(method = "cv", number=5,classProbs = TRUE))
# using bootstrapping resampling
model.boot <- train(APR ~APR_PROGRAM+APR_HSNAME+GENDER+codeHSGPA+HOURS_BROUGHT_TO_UNIVERSITY*codeHSGPA+PELL_VERF*GENDER+ETHNICITY, data=TrainingSet, method="glm",family="binomial",trControl=trainControl(method = "boot", number=5,classProbs = TRUE))
# using repeated k-fold cross validation resampling
model.repeatedcv <- train(APR ~APR_PROGRAM+APR_HSNAME+GENDER+codeHSGPA+HOURS_BROUGHT_TO_UNIVERSITY*codeHSGPA+PELL_VERF*GENDER+ETHNICITY, data=TrainingSet, method="glm",family="binomial",trControl=trainControl(method = "repeatedcv", number=5,classProbs = TRUE))
# prediction
ploocv <- predict(model.loocv, TrainingSet, "prob")
pcv <- predict(model.cv, TrainingSet, "prob")
pboot <- predict(model.boot, TrainingSet, "prob")
prepeatedcv <- predict(model.repeatedcv, TrainingSet, "prob")

#print(model.loocv)
#print(model.cv)
#print(model.boot)
#print(model.repeatedcv)
acc1 <- model.boot$results[,c(2:4)]
str(acc1)
# Models
models <- c("LOOCV","CV","BOOT","REPEATEDCV")
accuracy <- c(model.loocv$results[ ,2],
              model.cv$results[,2],
              model.boot$results[ ,2],
              model.repeatedcv$results[,2])

accy <- data.frame(models, accuracy)
write.csv(accy, "accuracy_application.csv")
```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
accuracy_application <- read.csv("G:/My Drive/#1_Jay's Space/AdmittedFTIC2021/accuracy_application.csv") %>% select(-1)
knitr::kable(accuracy_application)
```




## Prediction Index
  
  * Low-risk: Pr >= 0.90
  * Moderate-risk: 0.60 <= Pr < 0.90
  * High-risk: Pr < 0.60
  
## Table by risk

```{r echo=FALSE}
 knitr::kable(with(all_vari_PREDICTION, table(codeProb_APR)))
```

## Index Prediction by High School GPA

```{r echo=FALSE, message=FALSE, warning=FALSE}

library(highcharter)
library(dplyr)
table_APR <- all_vari_PREDICTION %>% group_by(codeProb_APR, codeHSGPA) %>% dplyr::summarise( Count=n())
hchart_index <- hchart(table_APR, "column", hcaes(x= codeHSGPA, y = Count, group= codeProb_APR)) 
hchart_index 
```






## Included  resutls table

```{r echo=FALSE}

library(readxl)
FTIC_202108_202105_Admitted_as_of_2_22_21 <- read_excel("G:/Shared drives/HMCSE-PAM Lab/DATA 202101/Admissions Data 2018-2020/FTIC 202108 202105 Admitted as of 2-22-21.xlsx") %>% select( UWFID,RACE, AA_ADDR1,AA_CITY,AA_ST,AA_ZIP,AA_CNTY, APP_NO,HS_MAT_YRS,HS_ENG_YRS)



library(DT)
 FOR_TABLE <- all_vari_PREDICTION %>% select(codeProb_APR,Prob_APR="PREDICTION",UWFID,HS_OFFER_GPA,codeHSGPA,Pred_Tier,College,Departments="CURRICULUM_DEPT",APR_PROGRAM="APR_PROGRAM",APR_HSNAME_imputated="APR_HSNAME1",APR_HSNAME, PROGRAM_DESC="PROGRAM_DESC",PROGRAM_CODE="PROGRAM_CODE",HS_NAME,GENDER="GENDER",contains("PELL"), BEST_SCORE ,HS_CORE_GPA ,HOURS_BROUGHT_TO_UNIVERSITY,ETHNICITY)

 FOR_TABLE1 <- merge(FOR_TABLE, FTIC_202108_202105_Admitted_as_of_2_22_21, by= "UWFID", all = T) %>% arrange(Prob_APR)
 
#write.csv(FOR_TABLE,"PREDICTION_results_better.csv")
 
results_table <- datatable(FOR_TABLE1,extensions = "Buttons",caption = "Probability of APR - 2021 FTIC(3,466)",
          filter = "top",
          options = list(dom="Blfrtip",buttons=c("copy","csv","excel","pdf","print"), lengthMenu=list(c(10,25,50,-1), c(10,25,50,"All")),pageLength=25))

results_table
```

## Pivot Table (Index Prediction vs. College)

```{r echo=FALSE, warning=FALSE}

library(rpivotTable)
For_pivot <- rpivotTable(FOR_TABLE, aggregatorName = "Count",
            rows = c("codeHSGPA", "codeProb_APR") ,
            
            cols = c("College"),
           
            subtotals = T, rendererName = "Table With Subtotal")

For_pivot
```

